---
title: "feature selection"
author: "ifanlyn@outlook.com"
date: "2021/7/29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Feature selection in species level

-   Step1: selected same trend features (SSTF), require more than 3/4 cohorts (at less 6 cohorts) perform the same trend. Multiple Median Fold Change (MMFC), compared each samples between 2 stages (Control vs CRC)

-   Step2: normalized by control median abundance (NCM), to correct for bias arising from uneven geographic sampling, we measured the relative abundance of each taxon by calculating the fraction of reads classified to each taxon and compared the raw distribution to the distribution of median control abundances within each cohort.

-   Step3: Comparison with Wilcoxon signed-rank test.

-   Step4: strict parameters, filtration the absolute MMFC \> 1

## Step 0: Loading packages and subroutines {.tabset}

### Package

```{r packages, message = F, warning = F}
require(DT) # for databale
require(reshape2) # for dcast
require(ggplot2)
require(ggthemes)
```

### Import table

```{r ImportTable }
ImportTable <- function(file, header = T, row.names = 1, sep = ',', check.names = FALSE, ...){
  data <- read.csv(file = file, header = header, row.names = row.names, sep = sep, check.names = check.names, ...)
  return(data)
}
```

### Show_table

```{r show_table}
Show_table <- function(df, rownames = T, filter="top", options = list(pageLength = 10, scrollX=T), ...){
  if (ncol(df) > 100) {
    df <- df[, 1:100]
    message('Due to the column dim is > 100, it only shows the top 100 columns')
  }
  datatable(df, rownames = rownames, 
            filter=filter, options = options, ... )

}
```

### MultMedFC

MultMedFC: Multiple Median Fold Change, compared each samples between 2 Stages (Control vs CRC)

```{r}
MultMedFC <- function(data1, data2, other_df = NULL, FeatureAsRow = T, instead_zero = 'rnorm'){
  if (FeatureAsRow) {
    data1 <- t(data1)
    data2 <- t(data2)
    if (! is.null(other_df)) {
      other_df <- t(other_df)
    }
  }
  if (! is.null(other_df)) {
    other_df <- as.data.frame(other_df)
  }
  data1 <- as.data.frame(data1)
  data2 <- as.data.frame(data2)
  if (! all(colnames(data1) == colnames(data2))) {
    message("rowname of data1 and data2 is not fullly same!! plz check it !!")
    return()
  }
  
  sf_df <- matrix(NA, ncol = 4); sf_df <- sf_df[-1, ]
  for (sf in colnames(data1)) {
    sf_FC <- NULL
    if (sum(data1[[sf]] == 0) == nrow(data1) & sum(data2[[sf]] == 0) == nrow(data2)) {
      sf_df <- rbind(sf_df, c(0, 0, 0, 1))
    } else {
      min_value <- min(c(data1[[sf]][data1[[sf]] != 0], data2[[sf]][data2[[sf]] != 0]))/10
      if(instead_zero == 'rnorm'){
        if (! is.null(other_df)) {
          other_df[[sf]][other_df[[sf]] == 0] <- rnorm(n = length(other_df[[sf]][other_df[[sf]] == 0]),
                                                       mean = min_value, sd = min_value/10)
        }
        set.seed(123)
        data1[[sf]][data1[[sf]] == 0] <- rnorm(n = length(data1[[sf]][data1[[sf]] == 0]), 
                                               mean = min_value, sd = min_value/10)
        data2[[sf]][data2[[sf]] == 0] <- rnorm(n = length(data2[[sf]][data2[[sf]] == 0]), 
                                               mean = min_value, sd = min_value/10)
      }else if(instead_zero == 'min'){
        if (! is.null(other_df)) {
          other_df[[sf]][other_df[[sf]] == 0] <- min_value
        }
        data1[[sf]][data1[[sf]] == 0] <- min_value
        data2[[sf]][data2[[sf]] == 0] <- min_value
      }else{
        message('instead_zero only accept the rnorm and min.')
        return()
      }
      modif_Data <- rbind(rbind(data1, data2), other_df)
      for (i1 in 1:nrow(data1)) {
        for (i2 in 1:nrow(data2)) {
          sf_FC <- c(sf_FC, data1[[sf]][i1]/data2[[sf]][i2])
        } 
      }
      sf_FC <- log2(sf_FC)
      sf_wil <- wilcox.test(sf_FC, conf.int = TRUE, mu = 0)
      sf_df <- rbind(sf_df, c(as.numeric(sf_wil$estimate), as.numeric(sf_wil$conf.int), as.numeric(sf_wil$p.value)))
    }
  }
  sf_df <- as.data.frame(sf_df); rownames(sf_df) <- colnames(data1); colnames(sf_df) <- c('Median', 'low-CI', 'high-CI', 'p-value')
  return(list(selFeatures_Data =sf_df, modif_Data = modif_Data))
}
```

### FileCreate

```{r}
FileCreate <- function(DirPath = "./",Prefix = "ExampleTest", Suffix = "pdf", version = "0.0.0"){
  date=as.character(Sys.Date())
  DirPath = gsub("/$","",DirPath)
  Suffix = gsub('^[.]',"",Suffix)
  if(! dir.exists(DirPath)){
    dir.create(DirPath,recursive =TRUE)
  }
  if (version == "0.0.0") {
    version=100
    while(TRUE){
      version_=paste(unlist(strsplit(as.character(version),"")),collapse=".")
      DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
      if(! file.exists(DirPathName)){
        return(DirPathName)
        break
      }
      version = version + 1
    }
  }else{
    DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
    if(! dir.exists(DirPathName)){
      return(DirPathName)
    }
    return(DirPathName)
  }
}
  
```

### matrix_wilcox

matrix_wilcox: subroutine to comparison each row in matrix

```{r subroutine: wilcoxon test from matrix}
matrix_wilcox <- function(data, group, ad_pvalue = 'BH'){
  data <- as.matrix(data)
  group1 <- names(group)[group == unique(group)[1]]
  group2 <- names(group)[group == unique(group)[2]]
  p_values <- list()
  for (i in 1:nrow(data)) {
    if (sum(data[i, ]) == 0) {
      p_values[[rownames(data)[i]]] = 1
    }else{
      p_values[[rownames(data)[i]]] = wilcox.test(data[i, group1], data[i, group2], 
                                                  alternative = "two.sided", exact = FALSE)$p.value
    }
  }
  p_values = data.frame(p_values = sapply(p_values, c))
  p_values$adj_pvalue = p.adjust(p_values$p_values, method = ad_pvalue, n = nrow(p_values))
  return(p_values)
}
```

## Step0: Import data

### Taxonomy name

```{r taxonomy name, warning = FALSE}
tax_name <- ImportTable(file = '../00.RawData/kraken2/2021-07-19-taxonomy_name-v1.0.csv', sep = ',')
tax_name <- tax_name[grep('k__Eukaryota', rownames(tax_name)), ]
tax_name$new_name <- gsub(pattern = 's__', replacement = '', x = tax_name$Specie) %>% 
  gsub('_', ' ', .) # due to Feature names of CreateSeuratObject cannot have underscores, replace with blanket space.
Show_table(tax_name)
```

### Taxonomy matrix

```{r taxonomy matrix}
relAbun_df <- ImportTable(file = '../05.Normalized/Rarefy_1329_fungi/2021-07-26-RelativeAbundance_matrix-v1.0.csv')
Show_table(relAbun_df)%>%
  formatSignif(columns = colnames(relAbun_df)[1:100],
               digits = 3, interval = 1)
```

### Meta information

```{r meta information}
meta_df <- ImportTable(file = '../00.RawData/metaData/2021-07-26-metaInfo-subgroup_1329-v6.0.csv')
Show_table(meta_df)
```

## Step 1: Selected same trend features (SSTF)

SSTF: selected same trend features, require more than 3/4 cohorts (at less 6 cohorts) perform the same trend.\
MultMedFC: Multiple Median Fold Change, compared each samples between 2 Stages (Control vs CRC).

-   Due to 0 couldn't be the denominator, the all the zero would instead by the value: `rnorm(n, min_value, min_sd)`,

    -   `n` means the counts of zero in the OTU matrix.

    -   `min_value` means the min value excluded the 0.

    -   `min_sd` equal `min_value /10`.

-   In order to reduce the bias of Fold change, the Fold Change was calculate by:

$$
Fold\ Change_i = median(\displaystyle \sum^{n_{crc}}_{j\ =\ 1}\displaystyle \sum^{n_{ctrl}}_{k\ =\ 1}{\frac{ReAbund_{j\ i}}{ReAbund_{k\ i}}})
$$

-   `i` means the fungi names.

-   `n_crc` or `n_ctrl` means the counts of CRC or control samples.

```{r, warning = FALSE}
calculateOrRead <- F

Trend_List <- list()
modify_List <- list()
if (calculateOrRead) {
  for (coh in unique(meta_df$Cohort)) {
    message(coh, ' begins at ', date())
    data1 <- relAbun_df[, rownames(meta_df)[meta_df$Cohort == coh & meta_df$Stage == 'CRC']]
    data2 <- relAbun_df[, rownames(meta_df)[meta_df$Cohort == coh & meta_df$Stage == 'CTRL']]
    other_df <- relAbun_df[, rownames(meta_df)[meta_df$Cohort == coh & meta_df$Stage == 'adenoma']]
    mmfc_res<- MultMedFC(data1 = data1, data2 = data2, other_df = other_df, FeatureAsRow = T)
    sub_trend <- mmfc_res$selFeatures_Data
    Trend_List[[coh]] <- sub_trend
    message(coh, ' has finished at ', date())
    median_file <- FileCreate(DirPath = '../07.FeatureSelection/01.SSTF/Cohort', Prefix = paste0('walsh-median-', coh), Suffix = 'csv')
    write.csv(x = Trend_List[[coh]], file = median_file)
    modify_csv <- FileCreate(DirPath = '../07.FeatureSelection/01.SSTF/Cohort', Prefix = paste0('modify_martix_norm-', coh), Suffix = 'csv')
    modify_List[[coh]] <- mmfc_res$modif_Data
    write.csv(x = modify_List[[coh]], file = modify_csv)
  }
  # Combine the cohorts together
  Trend_Comb <- Trend_List[[unique(meta_df$Cohort)[1]]]
  colnames(Trend_Comb) <- paste0(unique(meta_df$Cohort)[1], '-', colnames(Trend_Comb))
  modify_Comb <- modify_List[[unique(meta_df$Cohort)[1]]]
  for (coh in unique(meta_df$Cohort)[-1]) {
    sub_trend <- Trend_List[[coh]]
    colnames(sub_trend) <- paste0(coh, '-', colnames(sub_trend))
    Trend_Comb <- cbind(Trend_Comb, sub_trend)
    modify_Comb <- rbind(modify_Comb, modify_List[[coh]])
  }

  
  Trend_Comb_file <- FileCreate(DirPath = '../07.FeatureSelection/01.SSTF/',
                                Prefix = 'walsh-median-ALL', Suffix = 'csv')
  write.csv(x = Trend_Comb, file = Trend_Comb_file)
  modify_Comb_file <- FileCreate(DirPath = '../07.FeatureSelection/01.SSTF/',
                                 Prefix = 'modify_martix_norm-ALL', Suffix = 'csv')
  write.csv(x = modify_Comb, file = modify_Comb_file)
  
  Trend_Med_log2 <- Trend_Comb[,grep(pattern = 'Median', x = colnames(Trend_Comb))]
  Trend_Med_summary <- cbind(data.frame(N.count = rowSums(Trend_Med_log2 < 0), 
                                        P.count = rowSums(Trend_Med_log2 > 0),
                                        Missing = rowSums(Trend_Med_log2 == 0)),
                             Trend_Med_log2) %>% as.data.frame()
  median_file <- FileCreate(DirPath = '../07.FeatureSelection/01.SSTF',
                            Prefix = 'Summary-walsh-median', Suffix = 'csv')
  write.csv(x = Trend_Med_summary, file = median_file)
  
}else{
  for (coh in unique(meta_df$Cohort)) {
    Trend_List[[coh]] <- read.csv(paste0('../07.FeatureSelection/01.SSTF/Cohort/2021-07-29-walsh-median-', coh, '-v1.0.0.csv'), header = T, row.names = 1, check.names = F)
    modify_List[[coh]] <- read.csv(paste0('../07.FeatureSelection/01.SSTF/Cohort/2021-07-29-modify_martix_norm-', coh, '-v1.0.0.csv'), header = T, row.names = 1, check.names = F)
    
  }
  
  Trend_Comb <- read.csv('../07.FeatureSelection/01.SSTF/2021-07-29-walsh-median-ALL-v1.0.0.csv', header = T, row.names = 1, check.names = F)
  modify_Comb <- read.csv('../07.FeatureSelection/01.SSTF/2021-07-29-modify_martix_norm-ALL-v1.0.0.csv', header = T, row.names = 1, check.names = F)
  Trend_Med_summary <- read.csv('../07.FeatureSelection/01.SSTF/2021-07-29-Summary-walsh-median-v1.0.0.csv', header = T, row.names = 1, check.names = F)

}
Show_table(Trend_Med_summary) %>%
  formatSignif(columns = colnames(Trend_Med_summary[, c(4:ncol(Trend_Med_summary))]),
               digits = 3, interval = 1) %>%
  formatRound(columns = colnames(Trend_Med_summary[, c(1:3)]),
              digits = 0, interval = 1) %>%
  formatStyle(columns = colnames(Trend_Med_summary[, c(1:3)]),
              fontWeight = 'bold')
```

## Step 2: Multiple-Median Fold Change and Wilcoxon-rank test by Relative Median Abundance

Utilize Relative Median Abundance instead of raw or relative abundance. Only focus on the fungi with mean abundance \> 0.1% in each group

```{r}
calculate_or_not <- F
if (calculate_or_not) {
  # Relative Median Abundance
  comb_cohort <- matrix(NA, ncol = ncol(modify_Comb)); comb_cohort <- comb_cohort[-1, ]
  colnames(comb_cohort) <- colnames(modify_Comb)
  
  ctrl_modify_Comb <- modify_Comb[rownames(meta_df)[meta_df$Stage == 'CTRL'], ]
  crc_modify_Comb <- modify_Comb[rownames(meta_df)[meta_df$Stage == 'CRC'], ]
  
  #  mean abundance > 0.1%
  sel_idx <- colnames(ctrl_modify_Comb)[(colSums(ctrl_modify_Comb)/nrow(ctrl_modify_Comb) > 0.1) | (colSums(crc_modify_Comb)/nrow(crc_modify_Comb) > 0.1)]
  core_modify_Comb <- modify_Comb[, sel_idx]
  
  for (coh in unique(meta_df$Cohort)) {
    sub_ctrl <- core_modify_Comb[rownames(meta_df)[meta_df$Cohort == coh & meta_df$Stage == 'CTRL'], ]
    ctrl_median <- apply(sub_ctrl, 2, median)
    ctrl_median[ctrl_median==0] <- rep(10^-06, sum(ctrl_median==0))
    sub_cohort <- sweep(core_modify_Comb[rownames(meta_df)[meta_df$Cohort == coh], ], 2, ctrl_median, "/")
    comb_cohort <- rbind(comb_cohort, sub_cohort)
  }
  relMedian_csv <- FileCreate(DirPath = '../07.FeatureSelection/02.RelMedianAbundance', 
                              Prefix = 'normalized-0.1%-RelMedianAbundance-matrix', Suffix = 'csv')
  write.csv(x = comb_cohort, file = relMedian_csv)
  
  # multiple-Median Fold Change
  data1 <- comb_cohort[rownames(meta_df)[meta_df$Stage == 'CRC'], ] %>%
    t() %>% as.data.frame()
  data2 <- comb_cohort[rownames(meta_df)[meta_df$Stage == 'CTRL'], ]%>%
    t() %>% as.data.frame()
  all_mmfc_res<- MultMedFC(data1 = data1, data2 = data2, FeatureAsRow = T)
  
  all_trend <- all_mmfc_res$selFeatures_Data
  colnames(all_trend) <- paste0('all-', colnames(all_trend))
  median_file <- FileCreate(DirPath = '../07.FeatureSelection/01.SSTF/', Prefix = paste0('walsh-median-', 'all'), Suffix = 'csv')
  write.csv(x = all_trend, file = median_file)
  
  # wilcoxon-rank test 
  # mData <- as.data.frame(t(all_mmfc_res$modif_Data))
  mData <- as.data.frame(t(comb_cohort[rownames(meta_df)[meta_df$Stage != 'adenoma'], ]))
  groupList <- meta_df$Stage[meta_df$Stage %in% c('CTRL', 'CRC')]
  names(groupList) <- rownames(meta_df)[meta_df$Stage %in% c('CTRL', 'CRC')]
  wil_res <- matrix_wilcox(data = mData, group = groupList)
  wil_csv <- FileCreate(DirPath = '../07.FeatureSelection/01.SSTF/', Prefix = paste0('wilcoxon-', 'all'), Suffix = 'csv')
  write.csv(x = wil_res, file = wil_csv)
  
}else{
  comb_cohort <- ImportTable('../07.FeatureSelection/02.RelMedianAbundance/2021-07-30-normalized-0.1%-RelMedianAbundance-matrix-v1.0.0.csv')
  all_trend <- ImportTable('../07.FeatureSelection/01.SSTF/2021-07-30-walsh-median-all-v1.0.0.csv')
  wil_res <- ImportTable('../07.FeatureSelection/01.SSTF/2021-07-30-wilcoxon-all-v1.0.0.csv')
}

```

## Step 3: Intersection of 2 criterion

Only focus on 0.1% abundance relative Eukaryota.

Totally **42** fungi leave after twice filters, and **6** of them are also general Fold change difference in all samples (`|log2(FC)| > 0.5`).

```{r}
sel_idx_1 <- rownames(wil_res)[Trend_Med_summary[rownames(wil_res), ]$N.count >=6 | Trend_Med_summary[rownames(wil_res), ]$N.count <= 2]
length(sel_idx_1) # 112

sel_idx_2 <- rownames(wil_res)[wil_res$adj_pvalue < 0.05]
length(sel_idx_2) # 59

sel_idx <- intersect(sel_idx_1, sel_idx_2)
length(sel_idx) # 42

sel_trend <- all_trend[sel_idx, ]

sel3_trend <- sel_trend[abs(sel_trend$`all-Median`) > 0.5,  ] 
sel3_trend <- sel3_trend[order(abs(sel3_trend$`all-Median`), decreasing = T), ]

```

## Step 4: Core Eukaryota adjust p value

Plot the bar-plot and add the cutoff line with FDR \< 0.1 (main list) and FDR \< 0.01 (core list)

```{r}
# import the combine data 
comb_cohort <- ImportTable('../07.FeatureSelection/02.RelMedianAbundance/2021-07-30-normalized-0.1%-RelMedianAbundance-matrix-v1.0.0.csv')
wil_res <- ImportTable('../07.FeatureSelection/01.SSTF/2021-07-30-wilcoxon-all-v1.0.0.csv')
gg_wil_data <- wil_res
all_wil_data <- wil_res
gg_wil_data$EnrDep <- ifelse(all_trend[rownames(gg_wil_data),]$`all-Median` > 0, "Enriched", "Depleted")
# add log adjust p-value.
gg_wil_data <- gg_wil_data[gg_wil_data$adj_pvalue < 0.1, ]
nrow(gg_wil_data) # 74
gg_wil_data$`-log_q` <- -log10(gg_wil_data$adj_pvalue)
gg_wil_data_en <- gg_wil_data[gg_wil_data$EnrDep == "Enriched",]; gg_wil_data_en <- gg_wil_data_en[order(gg_wil_data_en$`-log_q`, decreasing = F), ]
gg_wil_data_de <- gg_wil_data[gg_wil_data$EnrDep != "Enriched",]; gg_wil_data_de <- gg_wil_data_de[order(gg_wil_data_de$`-log_q`, decreasing = T), ]

# gg_wil_data <- gg_wil_data[order(gg_wil_data$`-log_q`, decreasing = T), ]

gg_wil_data$name <- factor(rownames(gg_wil_data), c(rownames(gg_wil_data_de), rownames(gg_wil_data_en)))
sum(gg_wil_data$adj_pvalue < 0.01) # 33

message(gg_wil_data$`-log_q`[gg_wil_data$`-log_q` > 10])
gg_wil_data$`-log_q`[gg_wil_data$`-log_q` > 10] <- 10


main_list <- rownames(gg_wil_data) 
core_list <- rownames(gg_wil_data)[gg_wil_data$`-log_q` > 2]

g_bar <- ggplot(data = gg_wil_data, mapping = aes(x = name, y = `-log_q`))+
  geom_bar(stat="identity", alpha = ifelse(gg_wil_data$`-log_q` > 2, 0.9, 0.5), 
           # color = ifelse(gg_wil_data$EnrDep == "Enriched",  "#c00000", '#002060'),
           fill = ifelse(gg_wil_data$EnrDep == "Enriched",  "#c00000", '#002060'),
           color = 'black') +
  theme_tufte() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.line = element_line(size=1, colour = "black"))+ 
  scale_y_continuous(expand=c(0,0)) 

plot(g_bar)
barplot_pdf <- FileCreate(DirPath = '../07.FeatureSelection/03.Plot', 
                          Prefix = 'BarPlot-coreEukaryota', Suffix = 'pdf')
pdf(file = barplot_pdf, width = 20, height = 7)
plot(g_bar)
dev.off()

## all 

all_wil_data$`-log_q` <- -log10(all_wil_data$adj_pvalue)
all_wil_data <- all_wil_data[order(all_wil_data$`-log_q`, decreasing = T), ]
all_wil_data$name <- factor(rownames(all_wil_data), rownames(all_wil_data))
all_wil_data$`-log_q`[all_wil_data$`-log_q` > 10] <- 10

all_bar <- ggplot(data = all_wil_data, mapping = aes(x = name, y = `-log_q`))+
  geom_bar(stat="identity", fill = ifelse(all_wil_data$`-log_q` > 2, "snow4", 
                                          ifelse(all_wil_data$`-log_q` > 1,'snow3', 'snow2')), 
           # color = ifelse(gg_wil_data$`-log_q` > 2,  "snow4", 'snow1')) +
           color = 'white') +
  theme_tufte() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.line = element_line(size=1, colour = "black"))+ 
  scale_y_continuous(expand=c(0,0)) +
  geom_hline(yintercept = 2, color = 'red3', size = 0.8, linetype = "dashed") +
  geom_hline(yintercept = 1, color = 'indianred', size = 0.8, linetype = "dashed")

plot(all_bar)
all_barplot_pdf <- FileCreate(DirPath = '../07.FeatureSelection/03.Plot', 
                          Prefix = 'BarPlot-all-coreEukaryota', Suffix = 'pdf')
pdf(file = all_barplot_pdf, width = 20, height = 7)
plot(all_bar)
dev.off()

```

## Step 5: Multiple-Median Fold Change and Wilcoxon-rank test by Relative Median Abundance (plot figures)

```{r, fig.height=10, fig.width=12}

colRP3 <- c(colorRampPalette(colors = c("#004996","white"))(12),
           colorRampPalette(colors = c("white","white"))(6), 
           colorRampPalette(colors = c("white","#E20001"))(12))
bk <- seq(-1.5, 1.5, length.out = 30)

qbk <- seq(0, 5, length.out = 21)
qColPR <- c(rep('white', 4),  colorRampPalette(colors = c("white","black"))(17))
plot_list=list()

wil_matrix <- NULL
for (coh in unique(meta_df$Cohort)) {
  Trend_List[[coh]] <- read.csv(paste0('../07.FeatureSelection/01.SSTF/Cohort/2021-07-29-walsh-median-', coh, '-v1.0.0.csv'), header = T, row.names = 1, check.names = F)
  modify_List[[coh]] <- read.csv(paste0('../07.FeatureSelection/01.SSTF/Cohort/2021-07-29-modify_martix_norm-', coh, '-v1.0.0.csv'), header = T, row.names = 1, check.names = F)
  
  tmp <- Trend_List[[coh]]
  tmp1 <- tmp[main_list, "Median", drop = F] %>% t()
  p1=pheatmap(mat = tmp1, cluster_cols = F, cluster_rows = F,
              show_colnames = F, show_rownames  = F, legend = F,
              color = colRP3, breaks = bk, 
              cellheight = 20, cellwidth = 10)
  
  
  mod_df <- modify_List[[coh]] 
  sel_sample <- rownames(meta_df)[meta_df$Stage %in% c("CRC", "CTRL") &
                                    meta_df$Cohort == coh ]
  mod_df <- mod_df[sel_sample, main_list] %>% t()
  group_list <- meta_df[sel_sample, 'Stage']
  names(group_list) <- sel_sample
  
  sub_wil_res <- matrix_wilcox(data = mod_df, group = group_list)
  wil_csv <- FileCreate(DirPath = '../07.FeatureSelection/01.SSTF/Cohort/',
                     Prefix = paste0(coh, '-wilcox-pvalue_FDR'), Suffix = 'csv')
  sub_wil_res_ <- sub_wil_res 
  colnames(sub_wil_res_) <- paste0(coh, '_', colnames(sub_wil_res_))
  if (is.null(wil_matrix)) {
    wil_matrix <- sub_wil_res_
  }else{
    wil_matrix <- cbind(wil_matrix, sub_wil_res_)
  }
  write.csv(x = sub_wil_res, file = wil_csv)
  
  sub_wil_res$`-log_P` <- -log10(sub_wil_res$p_values)
  p_tmp <- sub_wil_res[main_list, "-log_P", drop = F] %>% t()
  
  p0=pheatmap(mat = p_tmp, cluster_cols = F, cluster_rows = F,
              show_colnames = F, show_rownames  = F, legend = F,
              color = qColPR, breaks = qbk, 
              cellheight = 10, cellwidth = 10)
  
  plot_list[[paste0(coh, "-p")]]=p0[[4]]
  plot_list[[paste0(coh, "-fc")]]=p1[[4]]
}
wil_csv_all <- FileCreate(DirPath = '../07.FeatureSelection/01.SSTF/',
                          Prefix = 'all-wilcox-pvalue_FDR', Suffix = 'csv')
write.csv(x = wil_matrix, file = wil_csv_all)
# tmp <- Trend_List$`2014_ZellerG`
# tmp1 <- tmp[main_list, "Median", drop = F] %>% t()
# 
# require(pheatmap)
# p1=pheatmap(mat = tmp1, cluster_cols = F, cluster_rows = F, show_colnames = F, show_rownames  = F, legend = F, color = colRP3, breaks = bk, cellheight = 20, cellwidth = 10)
# plot_list[['p1']]=p1[[4]]
# plot_list[['p2']]=p2[[4]]
# plot_list[['p3']]=p3[[4]]

require(gridExtra)
hm_pdf <- FileCreate(DirPath = '../07.FeatureSelection/03.Plot/',
                     Prefix = 'Core_Associations_Each_Studies', Suffix = 'pdf')
pdf(file = hm_pdf, width = 12, height = 10)
grid.arrange(grobs=plot_list, nrow=16)
dev.off()


p1=pheatmap(mat = tmp1, cluster_cols = F, cluster_rows = F,
            show_rownames  = F, legend = T,
            color = colRP3, breaks = bk, 
            cellheight = 120, cellwidth = 10)
p0=pheatmap(mat = p_tmp, cluster_cols = F, cluster_rows = F,
              show_colnames = F, show_rownames  = F, legend = T,
              color = qColPR, breaks = qbk, 
              cellheight = 120, cellwidth = 10)

l_plot_list <- list()
l_plot_list[[paste0(coh, "-p")]]=p0[[4]]
l_plot_list[[paste0(coh, "-fc")]]=p1[[4]]
hm_pdf <- FileCreate(DirPath = '../07.FeatureSelection/03.Plot/',
                     Prefix = 'Heatmap_legend', Suffix = 'pdf')
pdf(file = hm_pdf, width = 12, height = 30)
grid.arrange(grobs=l_plot_list, nrow=16)
dev.off()
```

## Step 6: All cohorts combined Fold Change

```{r}

fc_data <- all_trend[core_list, "all-Median", drop = F]
fc_data$name <- factor(core_list, core_list)
fc_data$absFC <- abs(fc_data$`all-Median`)

g_bar_fc <- ggplot(data = fc_data, mapping = aes(x = name, y = absFC))+
  geom_bar(stat="identity", fill = ifelse(fc_data$`all-Median` >= 0.5, "#E20001",
                                          ifelse(fc_data$`all-Median`>=0, '#F9D0D0',
                                                 ifelse(fc_data$`all-Median`>=-0.5,
                                                        '#B9CDE2', '#004996'))), 
           color = 'black') +
  theme_tufte() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.line = element_line(size=1, colour = "black"))+ 
  geom_hline(yintercept = 0.5, color = 'yellow', size = 1.3, linetype = "dashed") +
  scale_y_continuous(expand=c(0,0)) 

plot(g_bar_fc)
barplot_pdf <- FileCreate(DirPath = '../07.FeatureSelection/03.Plot', 
                          Prefix = 'BarPlot-abs_FC', Suffix = 'pdf')
pdf(file = barplot_pdf, width = 10, height = 5)
plot(g_bar_fc)
dev.off()


g_bar_fc2 <- ggplot(data = fc_data, mapping = aes(x = name, y = absFC))+
  geom_bar(stat="identity", fill = ifelse(fc_data$`all-Median` >= 1, "#E20001",
                                          ifelse(fc_data$`all-Median`>=0, '#F9D0D0',
                                                 ifelse(fc_data$`all-Median`>=-1,
                                                        '#B9CDE2', '#004996'))), 
           color = 'black') +
  theme_tufte() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.line = element_line(size=1, colour = "black"))+ 
  geom_hline(yintercept = 1, color = 'yellow', size = 1.3, linetype = "dashed") +
  scale_y_continuous(expand=c(0,0)) 

plot(g_bar_fc2)
barplot_pdf2 <- FileCreate(DirPath = '../07.FeatureSelection/03.Plot', 
                           Prefix = 'BarPlot-abs_FC_1', Suffix = 'pdf')
pdf(file = barplot_pdf2, width = 10, height = 5)
plot(g_bar_fc2)
dev.off()

```
