---
title: "Correlation in Eukaryota"
author: "ifanlyn@outlook.com"
date: "2021/8/4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages and subroutines {.tabset}

### Package

```{r packages, message = F, warning = F}
require(DT) # for databale
require(reshape2) # for dcast
require(ggplot2)
require(ggthemes)
require(corrplot)
```

### Import table

```{r ImportTable }
ImportTable <- function(file, header = T, row.names = 1, sep = ',', check.names = FALSE, ...){
  data <- read.csv(file = file, header = header, row.names = row.names, sep = sep, check.names = check.names, ...)
  return(data)
}
```

### Show_table

```{r show_table}
Show_table <- function(df, rownames = T, filter="top", options = list(pageLength = 10, scrollX=T), ...){
  if (ncol(df) > 100) {
    df <- df[, 1:100]
    message('Due to the column dim is > 100, it only shows the top 100 columns')
  }
  datatable(df, rownames = rownames, 
            filter=filter, options = options, ... )

}
```

### Spearman_corr

```{r}
Spearman_corr <- function(mat,rho = c(-0.2, 0.4), pvalue = 0.05, SampleAsColumn = T){
  translateNum <- function(data){
    data1 <- apply(data,2,as.numeric)
    rownames(data1) <- rownames(data)
    data1 <- as.data.frame(data1)
    return(data1)
  }
  if (!require(Hmisc)) {
    install.packages("Hmisc")
    library(Hmisc)
  }
  library(reshape2)
  # calculate the spearman
  if (SampleAsColumn) {
    mat_t <- t(translateNum(mat))
  }else{
    mat_t <- as.matrix(translateNum(mat))
  }
  
  mat_corr <- rcorr(mat_t, type = "spearman")
  # get the pvalue matrix and melt the matrix
  mat_p <- mat_corr[["P"]]
  sub_mat_p <- lower.tri(mat_p, diag = TRUE)
  mat_p[sub_mat_p] <- NA
  melt_p <- melt(mat_p, na.rm = TRUE)
  # get the rho matrix and melt the matrix
  mat_r <- mat_corr[["r"]]
  sub_mat_r <- lower.tri(mat_r, diag = TRUE)
  mat_r[sub_mat_r] <- NA
  melt_r <- melt(mat_r, na.rm = TRUE)
  # merge the melt-rho and melt-pvalue together
  melt_mat <- cbind(melt_r, melt_p[,3])
  colnames(melt_mat) <- c("Feature1", "Feature2", "rho", "pvalue")
  melt_mat_p <- melt_mat[which(melt_mat$pvalue < pvalue),]
  melt_mat_rho <- melt_mat[which(melt_mat$rho < rho[1] | melt_mat$rho > rho[2]),]
  melt_mat_p_rho <- melt_mat_rho[which(melt_mat_rho$pvalue < pvalue),]
  message("There are totally ", nrow(melt_mat), "-pair correlations\n",
          nrow(melt_mat_p), "-pair correlations are left after filtering the features whose pvalue is less than ", pvalue, "\n",
          nrow(melt_mat_rho), "-pair correlations are left after filtering the features whose rho is less than ", rho[1], " or large than ", rho[2], "\n",
          nrow(melt_mat_p_rho), "-pair correlations are left after above parameters.\n")
  rList <- list(mat_p = mat_corr[["P"]], mat_corr = mat_corr[["r"]], melt_mat = melt_mat, melt_filter = melt_mat_p_rho)
  return(rList)
}

```

### FileCreate

```{r}
FileCreate <- function(DirPath = "./",Prefix = "ExampleTest", Suffix = "pdf", version = "0.0.0"){
  date=as.character(Sys.Date())
  DirPath = gsub("/$","",DirPath)
  Suffix = gsub('^[.]',"",Suffix)
  if(! dir.exists(DirPath)){
    dir.create(DirPath,recursive =TRUE)
  }
  if (version == "0.0.0") {
    version=100
    while(TRUE){
      version_=paste(unlist(strsplit(as.character(version),"")),collapse=".")
      DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
      if(! file.exists(DirPathName)){
        return(DirPathName)
        break
      }
      version = version + 1
    }
  }else{
    DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
    if(! dir.exists(DirPathName)){
      return(DirPathName)
    }
    return(DirPathName)
  }
}
  
```

## Import data

### Import the modified data

Modified by three steps,

-   only leave the median abundance Eukaryota candidates( \>0.1%),

-   replace the zero by `rnorm(min/10, min/100)`,

-   and calculate the relative median abundance in each cohort

```{r}
mod_df <- ImportTable(file = '../07.FeatureSelection/02.RelMedianAbundance/2021-07-30-normalized-0.1%-RelMedianAbundance-matrix-v1.0.0.csv')
Show_table(mod_df)%>%
  formatSignif(columns = colnames(mod_df)[1:100],
               digits = 3, interval = 1)
```

### Import the meta-information

```{r}
meta_df <- ImportTable('../00.RawData/metaData/2021-08-03-metaInfo-subgroup_1625-v5.1.csv')

wil_res <- ImportTable('../07.FeatureSelection/01.SSTF/2021-07-30-wilcoxon-all-v1.0.0.csv')
wil_res <- wil_res[order(wil_res$adj_pvalue), ]
main_list <- rownames(wil_res)[wil_res$adj_pvalue < 0.05]   # 59
core_list <- rownames(wil_res)[wil_res$adj_pvalue < 0.01]   # 33

```

### Spearman correlation in all samples

```{r}

crc_spm_res <- Spearman_corr(mat = mod_df[rownames(meta_df)[meta_df$Stage == 'CRC'], core_list],
                             rho = c(-0.2, 0.4), pvalue = 0.05, SampleAsColumn = F)
crc_spm_res_rds <- FileCreate(DirPath = '../08.correlation/table/CRC',
                              Prefix = 'crc_spm_res',
                              Suffix = 'rds')
saveRDS(object = crc_spm_res, file = crc_spm_res_rds)

crc_spm_mat_p_csv <- FileCreate(DirPath = '../08.correlation/table/CRC',
                                Prefix = 'Matrix-pvalue-spearman-CRC',
                                Suffix = 'csv')
write.csv(x = crc_spm_res$mat_p, file = crc_spm_mat_p_csv)

crc_spm_mat_rho_csv <- FileCreate(DirPath = '../08.correlation/table/CRC',
                                Prefix = 'Matrix-rho-spearman-CRC',
                                Suffix = 'csv')
write.csv(x = crc_spm_res$mat_corr, file = crc_spm_mat_rho_csv)

crc_spm_melt_csv <- FileCreate(DirPath = '../08.correlation/table/CRC',
                                Prefix = 'Melt-all-spearman-CRC',
                                Suffix = 'csv')
write.csv(x = crc_spm_res$melt_mat, file = crc_spm_melt_csv)

crc_spm_selmelt_csv <- FileCreate(DirPath = '../08.correlation/table/CRC',
                                Prefix = 'Melt-filter-spearman-CRC',
                                Suffix = 'csv')
write.csv(x = crc_spm_res$melt_filter, file = crc_spm_selmelt_csv)


crc_spm_pval <- crc_spm_res$mat_p
crc_spm_corr <- crc_spm_res$mat_corr

crc_spm_corr[lower.tri(crc_spm_corr, diag = T)] <- NA
crc_spm_pval[lower.tri(crc_spm_pval, diag = T)] <- NA

crc_corr_melt <- melt(crc_spm_corr, na.rm = T)
crc_pval_melt <- melt(crc_spm_pval, na.rm = T)

crc_melt <- cbind(crc_corr_melt, pvalue = crc_pval_melt$value) %>% as.data.frame()
colnames(crc_melt) <- c('feature_1', 'feature_2', 'rho', 'pvalue')

crc_melt$rho <- ifelse(crc_melt$pvalue > 0.05, NA, crc_melt$rho)
crc_melt$feature_1 <- factor(crc_melt$feature_1, rev(core_list))
crc_melt$feature_2 <- factor(crc_melt$feature_2, rev(core_list))

crc_plot <- ggplot(crc_melt, aes(x = feature_2, y = feature_1, fill = rho)) +
  geom_tile(color="black", size=0.1) +
  scale_fill_distiller(palette = "Spectral", limits = c(-0.2, 0.5), na.value="grey80") +
  theme_minimal() +
  theme(panel.grid = element_blank(), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        axis.title = element_blank())

plot(crc_plot)
  
crc_pdf <- FileCreate(DirPath = '../08.correlation/', 
                      Prefix = 'triangle-corr-heatmap-CRC',
                      Suffix = 'pdf')
pdf(file = crc_pdf, width = 10, height = 10)
plot(crc_plot)
plot(crc_plot + theme(legend.position = 'none', 
                      axis.text= element_blank(),
                      axis.text.x = element_blank()))
dev.off()


ctrl_spm_res <- Spearman_corr(mat = mod_df[rownames(meta_df)[meta_df$Stage == 'CTRL'], core_list],
                             rho = c(-0.2, 0.4), pvalue = 0.05, SampleAsColumn = F)
ctrl_spm_res_rds <- FileCreate(DirPath = '../08.correlation/table/ctrl',
                              Prefix = 'ctrl_spm_res',
                              Suffix = 'rds')
saveRDS(object = ctrl_spm_res, file = ctrl_spm_res_rds)

ctrl_spm_mat_p_csv <- FileCreate(DirPath = '../08.correlation/table/ctrl',
                                Prefix = 'Matrix-pvalue-spearman-ctrl',
                                Suffix = 'csv')
write.csv(x = ctrl_spm_res$mat_p, file = ctrl_spm_mat_p_csv)

ctrl_spm_mat_rho_csv <- FileCreate(DirPath = '../08.correlation/table/ctrl',
                                Prefix = 'Matrix-rho-spearman-ctrl',
                                Suffix = 'csv')
write.csv(x = ctrl_spm_res$mat_corr, file = ctrl_spm_mat_rho_csv)

ctrl_spm_melt_csv <- FileCreate(DirPath = '../08.correlation/table/ctrl',
                                Prefix = 'Melt-all-spearman-ctrl',
                                Suffix = 'csv')
write.csv(x = ctrl_spm_res$melt_mat, file = ctrl_spm_melt_csv)

ctrl_spm_selmelt_csv <- FileCreate(DirPath = '../08.correlation/table/ctrl',
                                Prefix = 'Melt-filter-spearman-ctrl',
                                Suffix = 'csv')
write.csv(x = ctrl_spm_res$melt_filter, file = ctrl_spm_selmelt_csv)



ctrl_spm_pval <- ctrl_spm_res$mat_p
ctrl_spm_corr <- ctrl_spm_res$mat_corr

ctrl_spm_corr[lower.tri(ctrl_spm_corr, diag = T)] <- NA
ctrl_spm_pval[lower.tri(ctrl_spm_pval, diag = T)] <- NA

ctrl_corr_melt <- melt(ctrl_spm_corr, na.rm = T)
ctrl_pval_melt <- melt(ctrl_spm_pval, na.rm = T)

ctrl_melt <- cbind(ctrl_corr_melt, pvalue = ctrl_pval_melt$value) %>% as.data.frame()
colnames(ctrl_melt) <- c('feature_1', 'feature_2', 'rho', 'pvalue')

ctrl_melt$rho <- ifelse(ctrl_melt$pvalue > 0.05, NA, ctrl_melt$rho)
ctrl_melt$feature_1 <- factor(ctrl_melt$feature_1, rev(core_list))
ctrl_melt$feature_2 <- factor(ctrl_melt$feature_2, core_list)

ctrl_plot <- ggplot(ctrl_melt, aes(x = feature_2, y = feature_1, fill = rho)) +
  geom_tile(color="black", size=0.1) +
  scale_fill_distiller(palette = "Spectral", limits = c(-0.2, 0.5), na.value="grey80") +
  theme_minimal() +
  theme(panel.grid = element_blank(), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        axis.title = element_blank())

plot(ctrl_plot)
  
ctrl_pdf <- FileCreate(DirPath = '../08.correlation/', 
                      Prefix = 'triangle-corr-heatmap-CTRL',
                      Suffix = 'pdf')
pdf(file = ctrl_pdf, width = 10, height = 10)
plot(ctrl_plot)
plot(ctrl_plot + theme(legend.position = 'none', 
                       axis.text= element_blank(),
                      axis.text.x = element_blank()))
dev.off()


```
