---
title: "bacteria eukaryota correlation"
author: "ifanlyn@outlook.com"
date: "2021/8/6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages and subroutines {.tabset}

### Package

```{r packages, message = F, warning = F}
require(DT) # for databale
require(reshape2) # for dcast
require(ggplot2)
require(ggthemes)
require(corrplot)
require(DGCA)
require(ggthemes)  # for theme_calc()
```

### Import table

```{r ImportTable , message = F, warning = F}
ImportTable <- function(file, header = T, row.names = 1, sep = ',', check.names = FALSE, ...){
  data <- read.csv(file = file, header = header, row.names = row.names, sep = sep, check.names = check.names, ...)
  return(data)
}
```

### Show_table

```{r show_table, message = F, warning = F}
Show_table <- function(df, rownames = T, filter="top", options = list(pageLength = 10, scrollX=T), ...){
  if (ncol(df) > 100) {
    df <- df[, 1:100]
    message('Due to the column dim is > 100, it only shows the top 100 columns')
  }
  datatable(df, rownames = rownames, 
            filter=filter, options = options, ... )

}
```

### FileCreate

```{r , message = F, warning = F}
FileCreate <- function(DirPath = "./",Prefix = "ExampleTest", Suffix = "pdf", version = "0.0.0"){
  date=as.character(Sys.Date())
  DirPath = gsub("/$","",DirPath)
  Suffix = gsub('^[.]',"",Suffix)
  if(! dir.exists(DirPath)){
    dir.create(DirPath,recursive =TRUE)
  }
  if (version == "0.0.0") {
    version=100
    while(TRUE){
      version_=paste(unlist(strsplit(as.character(version),"")),collapse=".")
      DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
      if(! file.exists(DirPathName)){
        return(DirPathName)
        break
      }
      version = version + 1
    }
  }else{
    DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
    if(! dir.exists(DirPathName)){
      return(DirPathName)
    }
    return(DirPathName)
  }
}
  
```

## Import data {.tabset}

### metadata

```{r , message = F, warning = F}
meta_df <- ImportTable(file = '../00.RawData/metaData/2021-07-26-metaInfo-subgroup_1329-v6.0.csv')

```

### Import Eukaryota matrix

```{r , message = F, warning = F}

euk_otu <- ImportTable(file = '../07.FeatureSelection/02.RelMedianAbundance/2021-07-30-normalized-0.1%-RelMedianAbundance-matrix-v1.0.0.csv')
euk_sel <- ImportTable(file = '../07.FeatureSelection/01.SSTF/2021-07-30-wilcoxon-all-v1.0.0.csv')
euk_sel <- euk_sel[order(euk_sel$adj_pvalue),]
euk_core <- euk_otu[, rownames(euk_sel)[euk_sel$adj_pvalue < 0.01]]
euk_sel_list <- colnames(euk_core)

```

### Import Bacteria matrix

```{r , message = F, warning = F}

bac_otu <- ImportTable(file = '../09.Bacteria/02.RelMedianAbundance/2021-08-05-normalized-0.1%-RelMedianAbundance-matrix-v1.0.0.csv')
bac_sel <- ImportTable(file = '../09.Bacteria/2021-08-06-wilcoxon-all-v1.0.0.csv')
bac_sel <- bac_sel[order(bac_sel$adj_pvalue),]
bac_core_all <- bac_otu[, rownames(bac_sel)[bac_sel$adj_pvalue < 0.01 & bac_sel$AbsLog2FC > 0.5]]
bac_sel_list_all <- colnames(bac_core_all)
bac_sel_list <- bac_sel_list_all[!grepl('sp. ', bac_sel_list_all)]
bac_core <- bac_core_all[, bac_sel_list]
# bac_sel_list <- bac_sel_list_all[c(1:7, 9:10, 12:24, 27, 30, 32, 35, 38:42)]
```


### Set Parameters

```{r , message = F, warning = F}
param_list <- list(
  CTRL_CRC = list(
    filter_stage = 'adenoma',
    compared_stages = c('CTRL', 'CRC'),
    group_name = 'CTRL_CRC'
  ),
  adenoma_CRC = list(
    filter_stage = 'CTRL',
    compared_stages = c('adenoma', 'CRC'),
    group_name = 'adenoma_CRC'
  ),
  CTRL_adenoma = list(
    filter_stage = 'CRC',
    compared_stages = c( 'CTRL', 'adenoma'),
    group_name = 'CTRL_adenoma'
  )
)

```

## DGCA {.tabset}

The goal of DGCA is to calculate differential correlations across conditions. [link](https://github.com/andymckenzie/DGCA)

### pre-process DGCA import data formula

```{r , message = F, warning = F}
# pre-process the DGCA import data formula
## expression matrix
for (compared_name in names(param_list)) {
  filter_stage <- param_list[[compared_name]]$filter_stage
  compared_stages <- param_list[[compared_name]]$compared_stages
  group_name <- param_list[[compared_name]]$group_name
  
  comb_euk_bac <- cbind(bac_core, euk_core) %>% t()
  comb_euk_bac <- comb_euk_bac[, meta_df[colnames(comb_euk_bac), 'Stage'] != filter_stage]
  param_list[[compared_name]]$comb_euk_bac <- comb_euk_bac
  
  ## conditional matrix
  stage_design_mat <- lapply(meta_df[colnames(comb_euk_bac), 'Stage'], function(x){
    if (x == compared_stages[1]) {
      return(c(1, 0))
    }else{
      return(c(0, 1))
    }
  }) %>% as.data.frame() %>% t()
  colnames(stage_design_mat) <- compared_stages
  param_list[[compared_name]]$stage_design_mat <- stage_design_mat
}
```

### Comparing pairwise correlations across conditions

Runs the full discovery of differential correlation (ddcor) section for comparing pairwise correlations across conditions in the Differential Gene Correlation Analysis (DGCA) package.

```{r , message = F, warning = F}
for (compared_name in names(param_list)) {
  # DGCA calculate 
  set.seed(123)
  ddcor_res_mcrc <- ddcorAll(inputMat = param_list[[compared_name]]$comb_euk_bac, 
                             corrType = 'pearson', 
                             design = param_list[[compared_name]]$stage_design_mat,
                             compare = rev(param_list[[compared_name]]$compared_stages))
  param_list[[compared_name]]$ddcor_res_mcrc <- ddcor_res_mcrc
}

```

### Split Bacteria and Eukaryota

split the bacteria individual, eukaryota individual and bacteria & eukaryota

```{r , message = F, warning = F}
save_or_not <- T
for (compared_name in names(param_list)) {
  filter_stage <- param_list[[compared_name]]$filter_stage
  compared_stages <- param_list[[compared_name]]$compared_stages
  group_name <- param_list[[compared_name]]$group_name
  ddcor_res_mcrc <- param_list[[compared_name]]$ddcor_res_mcrc
  group_ddcor_index <- apply(ddcor_res_mcrc[,c(1,2)], 1, function(x){
    if (sum(x %in% colnames(euk_core)) == 1) {
      return('Intersect')
    }else if (sum(x %in% colnames(euk_core)) == 2){
      return('Eukaryota')
    }else{
      return('Bacteria')
    }
  })
  param_list[[compared_name]]$group_ddcor_index <- group_ddcor_index
  
  ## bacteria individual
  bac_dgca_res <- ddcor_res_mcrc[group_ddcor_index == 'Bacteria',]
  param_list[[compared_name]]$bac_dgca_res <- bac_dgca_res
  
  
  ## eukaryota individual
  euk_dgca_res <- ddcor_res_mcrc[group_ddcor_index == 'Eukaryota',]
  param_list[[compared_name]]$euk_dgca_res <- euk_dgca_res
  
  
  ## bacteria & eukaryota
  both_dgca_res <- ddcor_res_mcrc[group_ddcor_index == 'Intersect',]
  param_list[[compared_name]]$both_dgca_res <- both_dgca_res
  
  
  # all combine together
  all_dgca_summary <- ddcor_res_mcrc
  all_dgca_summary$type <- group_ddcor_index
  all_dgca_summary$abs_zScoreDiff <- abs(all_dgca_summary$zScoreDiff)
  param_list[[compared_name]]$all_dgca_summary <- all_dgca_summary
  
  if (save_or_not) {
    bac_dgca_csv <- FileCreate(DirPath = paste0('../08.correlation/DGCA/', group_name, '/table'), 
                               Prefix = 'melt-correlation-DGCA-Bacteria',
                               Suffix = 'csv')
    euk_dgca_csv <- FileCreate(DirPath = paste0('../08.correlation/DGCA/', group_name, '/table'), 
                               Prefix = 'melt-correlation-DGCA-Eukaryota',
                               Suffix = 'csv')
    both_dgca_csv <- FileCreate(DirPath = paste0('../08.correlation/DGCA/', group_name, '/table'), 
                                Prefix = 'melt-correlation-DGCA-Intersect',
                                Suffix = 'csv')
    all_dgca_csv <- FileCreate(DirPath = paste0('../08.correlation/DGCA/', group_name, '/table'), 
                               Prefix = 'melt-correlation-DGCA-ALL',
                               Suffix = 'csv')
    write.csv(x = bac_dgca_res, file = bac_dgca_csv, row.names = F)
    write.csv(x = euk_dgca_res, file = euk_dgca_csv, row.names = F)
    write.csv(x = both_dgca_res, file = both_dgca_csv, row.names = F)
    write.csv(x = all_dgca_summary, file = all_dgca_csv, row.names = F)
  }
}

```

### z-score distribution density plot
```{r , message = F, warning = F}
save_or_not <- T
for (compared_name in names(param_list)) {
  filter_stage <- param_list[[compared_name]]$filter_stage
  compared_stages <- param_list[[compared_name]]$compared_stages
  group_name <- param_list[[compared_name]]$group_name
  density_df <- param_list[[compared_name]]$all_dgca_summary
  
  den_fil_df <- density_df[density_df$empPVals < 0.05, ]
  param_list[[compared_name]]$sel_dgca_summary <- den_fil_df

  p <- ggplot(data=den_fil_df, aes(x=zScoreDiff, group=type, fill=type)) +
    geom_density(adjust=1.5, position="stack") +
    theme_calc() + scale_fill_calc()
  plot(p)


  p_s <- ggplot(data=den_fil_df, aes(x=zScoreDiff, group=type, fill=type)) +
    geom_density(adjust=1.5) +
    theme_calc() + scale_fill_calc()+
    facet_wrap(~type, ncol = 3, scale = 'free_y') +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      axis.ticks.x=element_blank())
  plot(p_s)
  if (save_or_not) {
    p_pdf <- FileCreate(DirPath = paste0('../08.correlation/DGCA/', group_name, '/fig'), 
                        Prefix = 'DensityPlot-z_score-distribution-all',
                        Suffix = 'pdf')
    ps_pdf <- FileCreate(DirPath = paste0('../08.correlation/DGCA/', group_name, '/fig'),
                        Prefix = 'DensityPlot-z_score-distribution-seperate',
                        Suffix = 'pdf')
    pdf(file = p_pdf, width = 15, height = 5)
    plot(p)
    dev.off()
    
    pdf(file = ps_pdf, width = 15, height = 3)
    plot(p_s)
    dev.off()
    sel_dgca_csv <- FileCreate(DirPath = paste0('../08.correlation/DGCA/', group_name, '/table'), 
                               Prefix = 'melt-correlation-DGCA-sel',
                               Suffix = 'csv')
    write.csv(x = den_fil_df, file = sel_dgca_csv, row.names = F)
    
  }
}



```

### z-score bar-plot subgroup

```{r , message = F, warning = F}
save_or_not <- T
for (compared_name in names(param_list)) {
  filter_stage <- param_list[[compared_name]]$filter_stage
  compared_stages <- param_list[[compared_name]]$compared_stages
  group_name <- param_list[[compared_name]]$group_name
  bar_df <- param_list[[compared_name]]$sel_dgca_summary 
    
  bar_df2 <- data.frame(Classes = factor(bar_df$Classes), 
                        type = factor(bar_df$type))
  bar_dcast = dcast(bar_df2, type ~ Classes, fun.aggregate = length)
  bar_melt = melt(bar_dcast, id.vars = "type", variable.name = 'Classes')
 
  g_bar <- ggplot(bar_melt, aes(x = Classes,y = value, fill = type)) + 
    geom_bar(stat = "identity", colour = "black", position = position_dodge(width = .8), width = 0.7) +
    geom_text(aes(label = value), position = position_dodge(width = .8), vjust = -0.5) +
    theme_calc() + scale_fill_calc() + theme(legend.position="none", axis.title = element_blank())
  plot(g_bar)
  
  if (save_or_not) {
    bar_melt_csv <- FileCreate(DirPath = paste0('../08.correlation/DGCA/', group_name, '/table'),
                            Prefix = 'DGCA-subgroup-table',
                            Suffix = 'csv')
    write.csv(x = bar_melt, file = bar_melt_csv, row.names = F)
    
    g_bar_pdf <- FileCreate(DirPath = paste0('../08.correlation/DGCA/', group_name, '/fig'),
                            Prefix = 'BarPlot-Class-table',
                            Suffix = 'pdf')
    pdf(file = g_bar_pdf, width = 15, height = 5)
    plot(g_bar)
    dev.off()

  }
  
} 
```



### Plot Eukaryotic triangle heat-map

```{r , message = F, warning = F}
save_or_not <- T
for (compared_name in names(param_list)) {
  filter_stage <- param_list[[compared_name]]$filter_stage
  compared_stages <- param_list[[compared_name]]$compared_stages
  group_name <- param_list[[compared_name]]$group_name
  melt_euk_df <- param_list[[compared_name]]$euk_dgca_res
  
  melt_euk_df$Gene1 <- factor(melt_euk_df$Gene1, rev(euk_sel_list))
  melt_euk_df$Gene2 <- factor(melt_euk_df$Gene2, rev(euk_sel_list))

  melt_euk_df[[paste0(compared_stages[2], '_cor')]][melt_euk_df[[paste0(compared_stages[2], '_pVal')]] >= 0.1] <- NA
  melt_euk_df[[paste0(compared_stages[1], '_cor')]][melt_euk_df[[paste0(compared_stages[1], '_pVal')]] >= 0.1] <- NA
  
  melt_euk_df[[paste0(compared_stages[2], '_cor')]][melt_euk_df[[paste0(compared_stages[2], '_cor')]] > 0.7] <- 0.7
  melt_euk_df[[paste0(compared_stages[2], '_cor')]][melt_euk_df[[paste0(compared_stages[2], '_cor')]] < -0.2] <- -0.2
  melt_euk_df[[paste0(compared_stages[1], '_cor')]][melt_euk_df[[paste0(compared_stages[1], '_cor')]] > 0.7] <- 0.7
  melt_euk_df[[paste0(compared_stages[2], '_cor')]][melt_euk_df[[paste0(compared_stages[1], '_cor')]] < -0.2] <- -0.2

  melt_euk_sel <- melt_euk_df[melt_euk_df$empPVals < 0.05,]
  param_list[[compared_name]]$melt_euk_sel <- melt_euk_sel
  
  g1_plot <- ggplot(melt_euk_df, aes(x = Gene2, y = Gene1, fill = .data[[paste0(compared_stages[1], '_cor')]])) +
    geom_tile(color="white", size=0.1) +
    scale_fill_distiller(palette = "Spectral", na.value="#7F7F7F", limits = c(-0.2, 0.7)) +
    theme_minimal() +
    theme(panel.grid = element_blank(), 
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
          axis.title = element_blank())
    # geom_point(data = melt_euk_sel, aes(x = Gene2, y = Gene1),
               # shape = 23, size = 4, fill = 'deeppink')
     # geom_text(label = ifelse(melt_euk_df$empPVals < 0.001, '***',
     #                          ifelse(melt_euk_df$empPVals < 0.01, "**",
     #                                 ifelse(melt_euk_df$empPVals < 0.05, "*", ""))),
     #           size=5, color = 'purple', fontface = "bold")
  plot(g1_plot)
  
  g2_plot <- ggplot(melt_euk_df, aes(x = Gene2, y = Gene1, fill = .data[[paste0(compared_stages[2], '_cor')]])) +
    geom_tile(color="white", size=0.1) +
    scale_fill_distiller(palette = "Spectral", na.value="#7F7F7F", limits = c(-0.2, 0.7)) +
    theme_minimal() +
    theme(panel.grid = element_blank(), 
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
          axis.title = element_blank())
    # geom_point(data = melt_euk_sel, aes(x = Gene2, y = Gene1),
               # shape = 23, size = 4, fill = 'deeppink')
     # geom_text(label = ifelse(melt_euk_df$empPVals < 0.001, '***',
     #                          ifelse(melt_euk_df$empPVals < 0.01, "**",
     #                                 ifelse(melt_euk_df$empPVals < 0.05, "*", ""))),
     #           size=5, color = 'purple', fontface = "bold")
  plot(g2_plot)
  
  param_list[[compared_name]][[paste0('triangle_HM_', compared_stages[1])]] <- g1_plot
  param_list[[compared_name]][[paste0('triangle_HM_', compared_stages[2])]] <- g2_plot
  
  if (save_or_not) {
    g1_pdf <- FileCreate(DirPath = paste0('../08.correlation/DGCA/', group_name, '/fig'),
                          Prefix = paste0('triangle-corr-heatmap-DGCA-', compared_stages[1]),
                          Suffix = 'pdf')
    pdf(file = g1_pdf, width = 10, height = 10)
    plot(g1_plot)
    plot(g1_plot + 
           theme(legend.position = 'none', 
                 axis.text= element_blank(),
                 axis.text.x = element_blank()))
    plot(g1_plot + 
           theme(legend.position = 'none', 
                 axis.text= element_blank(),
                 axis.text.x = element_blank())+
           geom_point(data = melt_euk_sel, aes(x = Gene2, y = Gene1),
                      shape = 11, size = 3, color = 'darkblue'))
    dev.off()
      
    g2_pdf <- FileCreate(DirPath = paste0('../08.correlation/DGCA/', group_name, '/fig'),
                         Prefix = paste0('triangle-corr-heatmap-DGCA-', compared_stages[2]),
                         Suffix = 'pdf')
    pdf(file = g2_pdf, width = 10, height = 10)
    plot(g2_plot)
    plot(g2_plot + 
           theme(legend.position = 'none', 
                 axis.text= element_blank(),
                 axis.text.x = element_blank()))
    plot(g2_plot + 
           theme(legend.position = 'none', 
                 axis.text= element_blank(),
                 axis.text.x = element_blank())+
           geom_point(data = melt_euk_sel, aes(x = Gene2, y = Gene1),
                      shape = 11, size = 3, color = 'darkblue'))
    dev.off()
  }
}

```

### Plot Eukaryotic & Bateria rectangle heat-map

```{r , message = F, warning = F}
save_or_not <- T
for (compared_name in names(param_list)) {
  filter_stage <- param_list[[compared_name]]$filter_stage
  compared_stages <- param_list[[compared_name]]$compared_stages
  group_name <- param_list[[compared_name]]$group_name
  melt_both_df <- param_list[[compared_name]]$both_dgca_res
  
  
  melt_both_df$Gene1 <- factor(melt_both_df$Gene1, rev(bac_sel_list))
  melt_both_df$Gene2 <- factor(melt_both_df$Gene2, rev(euk_sel_list))
  melt_both_df[[paste0(compared_stages[2], '_cor')]][melt_both_df[[paste0(compared_stages[2], '_pVal')]] >= 0.1] <- NA
  melt_both_df[[paste0(compared_stages[1], '_cor')]][melt_both_df[[paste0(compared_stages[1], '_pVal')]] >= 0.1] <- NA
  
  melt_both_df[[paste0(compared_stages[2], '_cor')]][melt_both_df[[paste0(compared_stages[2], '_cor')]] > 0.7] <- 0.7
  melt_both_df[[paste0(compared_stages[2], '_cor')]][melt_both_df[[paste0(compared_stages[2], '_cor')]] < -0.2] <- -0.2
  melt_both_df[[paste0(compared_stages[1], '_cor')]][melt_both_df[[paste0(compared_stages[1], '_cor')]] > 0.7] <- 0.7
  melt_both_df[[paste0(compared_stages[2], '_cor')]][melt_both_df[[paste0(compared_stages[1], '_cor')]] < -0.2] <- -0.2


  melt_both_sel <- melt_both_df[melt_both_df$empPVals < 0.05,]
  param_list[[compared_name]]$melt_both_sel <- melt_both_sel
  
  g1_plot <- ggplot(melt_both_df, aes(x = Gene1, y = Gene2, fill = .data[[paste0(compared_stages[1], '_cor')]])) +
    geom_tile(color="white", size=0.1) +
    scale_fill_distiller(palette = "Spectral", na.value="#7F7F7F", limits = c(-0.2, 0.7)) +
    theme_minimal() +
    theme(panel.grid = element_blank(), 
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
          axis.title = element_blank())
    # geom_point(data = melt_both_sel, aes(x = Gene1, y = Gene2), 
               # shape = 23, size = 4, fill = 'deeppink')
    # geom_text(label = ifelse(melt_both_df$empPVals < 0.001, '***',
    #                           ifelse(melt_both_df$empPVals < 0.01, "**",
    #                                  ifelse(melt_both_df$empPVals < 0.05, "*", ""))), 
    #            size=5, color = 'purple', fontface = "bold")
  plot(g1_plot)
  
  
  g2_plot <- ggplot(melt_both_df, aes(x = Gene1, y = Gene2, fill =  .data[[paste0(compared_stages[2], '_cor')]])) +
    geom_tile(color="white", size=0.1) +
    scale_fill_distiller(palette = "Spectral", na.value="#7F7F7F", limits = c(-0.2, 0.7)) +
    theme_minimal() +
    theme(panel.grid = element_blank(), 
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
          axis.title = element_blank())
    # geom_point(data = melt_both_sel, aes(x = Gene1, y = Gene2), 
               # shape = 23, size = 4, fill = 'deeppink')
     # geom_text(label = ifelse(melt_both_df$empPVals < 0.001, '***',
     #                          ifelse(melt_both_df$empPVals < 0.01, "**",
     #                                 ifelse(melt_both_df$empPVals < 0.05, "*", ""))), 
     #           size=5, color = 'purple', fontface = "bold")
  plot(g2_plot)
  
  param_list[[compared_name]][[paste0('rectangle_HM_', compared_stages[1])]] <- g1_plot
  param_list[[compared_name]][[paste0('rectangle_HM_', compared_stages[2])]] <- g2_plot

  if (save_or_not) {
    g1_pdf <- FileCreate(DirPath = paste0('../08.correlation/DGCA/', group_name, '/fig'),
                          Prefix = paste0('rectangle-corr-heatmap-DGCA-', compared_stages[1]),
                          Suffix = 'pdf')
    pdf(file = g1_pdf, width = 9.4, height = 10)
    plot(g1_plot)
    plot(g1_plot + 
           theme(legend.position = 'none', 
                 axis.text= element_blank(),
                 axis.text.x = element_blank()))
    plot(g1_plot + 
           theme(legend.position = 'none', 
                 axis.text= element_blank(),
                 axis.text.x = element_blank())+
           geom_point(data = melt_both_sel, aes(x = Gene1, y = Gene2),
                      shape = 11, size = 3, color = 'darkblue'))
    dev.off()
  
  
    
    g2_pdf <- FileCreate(DirPath = paste0('../08.correlation/DGCA/', group_name, '/fig'),
                          Prefix = paste0('rectangle-corr-heatmap-DGCA-', compared_stages[2]),
                          Suffix = 'pdf')
    pdf(file = g2_pdf, width = 9.4, height = 10)
    plot(g2_plot)
    plot(g2_plot + 
           theme(legend.position = 'none', 
                 axis.text= element_blank(),
                 axis.text.x = element_blank()))
    plot(g2_plot + 
           theme(legend.position = 'none', 
                 axis.text= element_blank(),
                 axis.text.x = element_blank())+
           geom_point(data = melt_both_sel, aes(x = Gene1, y = Gene2),
                      shape = 11, size = 3, color = 'darkblue'))
    dev.off()
  }
  
}

```



### Plot CRC heatmap

```{r}

tri_plot <- param_list[['CTRL_CRC']][['triangle_HM_CRC']]
rec_plot <- param_list[['CTRL_CRC']][['rectangle_HM_CRC']]

ctrl_crc_sel_both <- param_list$CTRL_CRC$melt_both_sel
aden_crc_sel_both <- param_list$adenoma_CRC$melt_both_sel

ctrl_crc_sel_euk <- param_list$CTRL_CRC$melt_euk_sel
aden_crc_sel_euk <- param_list$adenoma_CRC$melt_euk_sel


tri_pdf <- FileCreate(DirPath = '../08.correlation/DGCA/',
                        Prefix = 'triangle-corr-heatmap-DGCA-CRC-3group',
                        Suffix = 'pdf')
pdf(file = tri_pdf, width = 10, height = 10)
plot(tri_plot)
plot(tri_plot + 
       theme(legend.position = 'none', 
             axis.text= element_blank(),
             axis.text.x = element_blank()))
plot(tri_plot + 
       theme(legend.position = 'none',
             axis.text= element_blank(),
             axis.text.x = element_blank())+
       geom_point(data = ctrl_crc_sel_euk, aes(x = Gene2, y = Gene1),
                  shape = 23, size = 4, fill = 'deeppink')+
       geom_point(data = aden_crc_sel_euk, aes(x = Gene2, y = Gene1),
                  shape = 20, size = 3, color = 'navy'))

dev.off()

rev_pdf <- FileCreate(DirPath = '../08.correlation/DGCA/',
                        Prefix = 'rectangle-corr-heatmap-DGCA-CRC-3group',
                        Suffix = 'pdf')
pdf(file = rev_pdf, width = 9.4, height = 10)
plot(rec_plot)
plot(rec_plot + 
       theme(legend.position = 'none', 
             axis.text= element_blank(),
             axis.text.x = element_blank()))
plot(rec_plot + 
       theme(legend.position = 'none',
             axis.text= element_blank(),
             axis.text.x = element_blank())+
       geom_point(data = ctrl_crc_sel_both, aes(x = Gene1, y = Gene2),
                  shape = 23, size = 4, fill = 'deeppink')+
       geom_point(data = aden_crc_sel_both, aes(x = Gene1, y = Gene2),
                  shape = 20, size = 3, color = 'navy'))
dev.off()

```















