---
title: "Bacteria preprocess and feature selections"
author: "ifanlyn@outlook.com"
date: "2021/8/5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 0: Loading packages and subroutines {.tabset}

### Package

```{r packages, message = F, warning = F}
require(DT) # for databale
require(reshape2) # for dcast
require(ggplot2)
require(ggthemes)
require(ggrepel)
```

### Import table

```{r ImportTable }
ImportTable <- function(file, header = T, row.names = 1, sep = ',', check.names = FALSE, ...){
  data <- read.csv(file = file, header = header, row.names = row.names, sep = sep, check.names = check.names, ...)
  return(data)
}
```

### Show_table

```{r show_table}
Show_table <- function(df, rownames = T, filter="top", options = list(pageLength = 10, scrollX=T), ...){
  if (ncol(df) > 100) {
    df <- df[, 1:100]
    message('Due to the column dim is > 100, it only shows the top 100 columns')
  }
  datatable(df, rownames = rownames, 
            filter=filter, options = options, ... )

}
```

### MultMedFC

MultMedFC: Multiple Median Fold Change, compared each samples between 2 Stages (Control vs CRC)

```{r}
MultMedFC <- function(data1, data2, other_df = NULL, FeatureAsRow = T, instead_zero = 'rnorm'){
  if (FeatureAsRow) {
    data1 <- t(data1)
    data2 <- t(data2)
    if (! is.null(other_df)) {
      other_df <- t(other_df)
    }
  }
  if (! is.null(other_df)) {
    other_df <- as.data.frame(other_df)
  }
  data1 <- as.data.frame(data1)
  data2 <- as.data.frame(data2)
  if (! all(colnames(data1) == colnames(data2))) {
    message("rowname of data1 and data2 is not fullly same!! plz check it !!")
    return()
  }
  
  sf_df <- matrix(NA, ncol = 4); sf_df <- sf_df[-1, ]
  for (sf in colnames(data1)) {
    # message(date(), "\t", sf)
    sf_FC <- NULL
    if (sum(data1[[sf]] == 0) == nrow(data1) & sum(data2[[sf]] == 0) == nrow(data2)) {
    } else {
      min_value <- min(c(data1[[sf]][data1[[sf]] != 0], data2[[sf]][data2[[sf]] != 0]))/10
      if(instead_zero == 'rnorm'){
        if (! is.null(other_df)) {
          other_df[[sf]][other_df[[sf]] == 0] <- rnorm(n = length(other_df[[sf]][other_df[[sf]] == 0]),
                                                       mean = min_value, sd = min_value/10)
        }
        set.seed(123)
        data1[[sf]][data1[[sf]] == 0] <- rnorm(n = length(data1[[sf]][data1[[sf]] == 0]), 
                                               mean = min_value, sd = min_value/10)
        data2[[sf]][data2[[sf]] == 0] <- rnorm(n = length(data2[[sf]][data2[[sf]] == 0]), 
                                               mean = min_value, sd = min_value/10)
      }else if(instead_zero == 'min'){
        if (! is.null(other_df)) {
          other_df[[sf]][other_df[[sf]] == 0] <- min_value
        }
        data1[[sf]][data1[[sf]] == 0] <- min_value
        data2[[sf]][data2[[sf]] == 0] <- min_value
      }else{
        message('instead_zero only accept the rnorm and min.')
        return()
      }
      modif_Data <- rbind(rbind(data1, data2), other_df)
      # for (i1 in 1:nrow(data1)) {
      #   for (i2 in 1:nrow(data2)) {
      #     sf_FC <- c(sf_FC, data1[[sf]][i1]/data2[[sf]][i2])
      #   } 
      # }
      # sf_FC <- log2(sf_FC)
      # sf_wil <- wilcox.test(sf_FC, conf.int = TRUE, mu = 0)
      # sf_df <- rbind(sf_df, c(as.numeric(sf_wil$estimate), as.numeric(sf_wil$conf.int), as.numeric(sf_wil$p.value)))
    }
  }
  # sf_df <- as.data.frame(sf_df); rownames(sf_df) <- colnames(data1); colnames(sf_df) <- c('Median', 'low-CI', 'high-CI', 'p-value')
  # return(list(selFeatures_Data =sf_df, modif_Data = modif_Data))
  return(modif_Data)
}
```

### FileCreate

```{r}
FileCreate <- function(DirPath = "./",Prefix = "ExampleTest", Suffix = "pdf", version = "0.0.0"){
  date=as.character(Sys.Date())
  DirPath = gsub("/$","",DirPath)
  Suffix = gsub('^[.]',"",Suffix)
  if(! dir.exists(DirPath)){
    dir.create(DirPath,recursive =TRUE)
  }
  if (version == "0.0.0") {
    version=100
    while(TRUE){
      version_=paste(unlist(strsplit(as.character(version),"")),collapse=".")
      DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
      if(! file.exists(DirPathName)){
        return(DirPathName)
        break
      }
      version = version + 1
    }
  }else{
    DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
    if(! dir.exists(DirPathName)){
      return(DirPathName)
    }
    return(DirPathName)
  }
}
  
```

### matrix_wilcox

matrix_wilcox: subroutine to comparison each row in matrix

```{r subroutine: wilcoxon test from matrix}
matrix_wilcox <- function(data, group, ad_pvalue = 'BH'){
  data <- as.matrix(data)
  group1 <- names(group)[group == unique(group)[1]]
  group2 <- names(group)[group == unique(group)[2]]
  p_values <- list()
  for (i in 1:nrow(data)) {
    if (sum(data[i, ]) == 0) {
      p_values[[rownames(data)[i]]] = 1
    }else{
      p_values[[rownames(data)[i]]] = wilcox.test(data[i, group1], data[i, group2], 
                                                  alternative = "two.sided", exact = FALSE)$p.value
    }
  }
  p_values = data.frame(p_values = sapply(p_values, c))
  p_values$adj_pvalue = p.adjust(p_values$p_values, method = ad_pvalue, n = nrow(p_values))
  return(p_values)
}
```

## Step0: Import data

### Taxonomy name

```{r taxonomy name, warning = FALSE}
tax_name <- ImportTable(file = '../00.RawData/kraken2/2021-07-19-taxonomy_name-v1.0.csv', sep = ',')
tax_name <- tax_name[grep('k__Bacteria', rownames(tax_name)), ]
tax_name$new_name <- gsub(pattern = 's__', replacement = '', x = tax_name$Specie) %>% 
  gsub('_', ' ', .) # due to Feature names of CreateSeuratObject cannot have underscores, replace with blanket space.
Show_table(tax_name)
```

### Meta information

```{r meta information}
meta_df <- ImportTable(file = '../00.RawData/metaData/2021-07-26-metaInfo-subgroup_1329-v6.0.csv')
Show_table(meta_df)
```

### Taxonomy matrix

```{r taxonomy matrix}
relAbun_df <- ImportTable(file = '../02.Filter_samples/Bacteria_RelAbun_1625/2021-08-03-Selected-Bac_RelAbun-Species-v1.0.csv')
relAbun_df <- relAbun_df[, rownames(meta_df)]

Show_table(relAbun_df)%>%
  formatSignif(columns = colnames(relAbun_df)[1:100],
               digits = 3, interval = 1)
```

## Step 1: Selected same trend features (SSTF)

SSTF: selected same trend features, require more than 3/4 cohorts (at less 6 cohorts) perform the same trend.\
MultMedFC: Multiple Median Fold Change, compared each samples between 2 Stages (Control vs CRC).

-   Due to 0 couldn't be the denominator, the all the zero would instead by the value: `rnorm(n, min_value, min_sd)`,

    -   `n` means the counts of zero in the OTU matrix.

    -   `min_value` means the min value excluded the 0.

    -   `min_sd` equal `min_value /10`.

-   In order to reduce the bias of Fold change, the Fold Change was calculate by:

$$
Fold\ Change_i = median(\displaystyle \sum^{n_{crc}}_{j\ =\ 1}\displaystyle \sum^{n_{ctrl}}_{k\ =\ 1}{\frac{ReAbund_{j\ i}}{ReAbund_{k\ i}}})
$$

-   `i` means the fungi names.

-   `n_crc` or `n_ctrl` means the counts of CRC or control samples.

```{r, warning = FALSE}
calculateOrRead <- T

ctrl_relAbun_df <- relAbun_df[, rownames(meta_df)[meta_df$Stage == 'CTRL']]
crc_relAbun_df <- relAbun_df[, rownames(meta_df)[meta_df$Stage == 'CRC']]


sel_idx <- rownames(ctrl_relAbun_df)[(rowSums(ctrl_relAbun_df)/ncol(ctrl_relAbun_df) > 0.01) | (rowSums(crc_relAbun_df)/ncol(crc_relAbun_df) > 0.01)]
length(sel_idx) # 273
core_relAbun_df <- relAbun_df[sel_idx, ]

modify_Comb <- matrix(NA, ncol = length(sel_idx)); modify_Comb <- modify_Comb[-1, ]
modify_List <- list()
if (calculateOrRead) {
  for (coh in unique(meta_df$Cohort)) {
    message(coh, ' begins at ', date())
    data1 <- core_relAbun_df[, rownames(meta_df)[meta_df$Cohort == coh & meta_df$Stage == 'CRC']]
    data2 <- core_relAbun_df[, rownames(meta_df)[meta_df$Cohort == coh & meta_df$Stage == 'CTRL']]
    other_df <- core_relAbun_df[, rownames(meta_df)[meta_df$Cohort == coh & meta_df$Stage == 'adenoma']]
    mod_df <- MultMedFC(data1 = data1, data2 = data2, other_df = other_df, FeatureAsRow = T)
    message(coh, ' has finished at ', date())
    modify_csv <- FileCreate(DirPath = '../09.Bacteria/01.modified_data/Cohort', Prefix = paste0('modify_martix_norm-', coh), Suffix = 'csv')
    modify_List[[coh]] <- mod_df
    write.csv(x = modify_List[[coh]], file = modify_csv)
    modify_Comb <- rbind(modify_Comb, mod_df)
    if (any(is.na(mod_df))) {
      message(coh)
    }
  }
  
  modify_comb_csv <- FileCreate(DirPath = '../09.Bacteria/01.modified_data/', Prefix = paste0('modify_martix_norm-', 'ALL'), Suffix = 'csv')
  write.csv(x = modify_Comb, file = modify_comb_csv)
}else{
  for (coh in unique(meta_df$Cohort)) {
    modify_List[[coh]] <- read.csv(paste0('../09.Bacteria/01.modified_data/Cohort/2021-08-05-modify_martix_norm-', coh, '-v1.0.0.csv'), header = T, row.names = 1, check.names = F)
    
  }
  
  modify_Comb <- read.csv('../09.Bacteria/01.modified_data/2021-08-05-modify_martix_norm-ALL-v1.0.0.csv', header = T, row.names = 1, check.names = F)

}
Show_table(modify_Comb) %>%
  formatSignif(columns = colnames(modify_Comb)[1:100],
               digits = 3, interval = 1)
```

## Step 2: Multiple-Median Fold Change and Wilcoxon-rank test by Relative Median Abundance

Utilize Relative Median Abundance instead of raw or relative abundance. Only focus on the fungi with mean abundance \> 0.1% in each group

```{r}
calculate_or_not <- F
if (calculate_or_not) {
  # Relative Median Abundance
  comb_cohort <- matrix(NA, ncol = ncol(modify_Comb)); comb_cohort <- comb_cohort[-1, ]
  colnames(comb_cohort) <- colnames(modify_Comb)
  
  ctrl_modify_Comb <- modify_Comb[rownames(meta_df)[meta_df$Stage == 'CTRL'], ]
  crc_modify_Comb <- modify_Comb[rownames(meta_df)[meta_df$Stage == 'CRC'], ]
  
  #  mean abundance > 0.01%

  core_modify_Comb <- modify_Comb
  
  for (coh in unique(meta_df$Cohort)) {
    sub_ctrl <- core_modify_Comb[rownames(meta_df)[meta_df$Cohort == coh & meta_df$Stage == 'CTRL'], ]
    ctrl_median <- apply(sub_ctrl, 2, median)
    ctrl_median[ctrl_median==0] <- rep(10^-06, sum(ctrl_median==0))
    sub_cohort <- sweep(core_modify_Comb[rownames(meta_df)[meta_df$Cohort == coh], ], 2, ctrl_median, "/")
    comb_cohort <- rbind(comb_cohort, sub_cohort)
  }
  relMedian_csv <- FileCreate(DirPath = '../09.Bacteria/02.RelMedianAbundance', 
                              Prefix = 'normalized-0.1%-RelMedianAbundance-matrix', Suffix = 'csv')
  write.csv(x = comb_cohort, file = relMedian_csv)
  
  # multiple-Median Fold Change
  data1 <- comb_cohort[rownames(meta_df)[meta_df$Stage == 'CRC'], ] %>%
    t() %>% as.data.frame()
  data2 <- comb_cohort[rownames(meta_df)[meta_df$Stage == 'CTRL'], ]%>%
    t() %>% as.data.frame()
  all_mmfc_res<- MultMedFC(data1 = data1, data2 = data2, FeatureAsRow = T)

  
  data1_m <- all_mmfc_res[rownames(meta_df)[meta_df$Stage == 'CRC'], ] %>%
    t() %>% as.data.frame()
  data2_m <- all_mmfc_res[rownames(meta_df)[meta_df$Stage == 'CTRL'], ] %>%
    t() %>% as.data.frame()
  
  median_list <- NULL
  for (sf in rownames(data1)) {
    dd_tmp <- t(1/data2_m[sf, , drop = F]) %*% as.matrix(data1_m[sf, , drop = F])
    median_list[sf] <- median(dd_tmp)
  }
  
 
  # wilcoxon-rank test 
  mData <- as.data.frame(t(all_mmfc_res))
  groupList <- meta_df$Stage[meta_df$Stage %in% c('CTRL', 'CRC')]
  names(groupList) <- rownames(meta_df)[meta_df$Stage %in% c('CTRL', 'CRC')]
  wil_res <- matrix_wilcox(data = mData, group = groupList)
  wil_res <- cbind(wil_res, mFC = median_list[rownames(wil_res)])
  wil_res <- cbind(wil_res, AbsLog2FC = abs(log2(wil_res$mFC)))
  wil_res <- wil_res[order(wil_res$AbsLog2FC, decreasing = T), ]
  wil_csv <- FileCreate(DirPath = '../09.Bacteria/', Prefix = paste0('wilcoxon-', 'all'), Suffix = 'csv')
  write.csv(x = wil_res, file = wil_csv)
  
}else{
  comb_cohort <- ImportTable('../09.Bacteria/02.RelMedianAbundance/2021-08-05-normalized-0.1%-RelMedianAbundance-matrix-v1.0.0.csv')
  wil_res <- ImportTable('../09.Bacteria/2021-08-06-wilcoxon-all-v1.0.0.csv')
}

```

## Step 3: Volcano plot for bacteria candidate

```{r}
# volcano_df <- wil_res
# volcano_df$`-log10_Q` <- -log10(volcano_df$adj_pvalue)
# volcano_df$log2FC <- log2(volcano_df$mFC)
# g <- ggplot(volcano_df, aes(x = log2FC, y = `-log10_Q`)) +
#   geom_point(alpha = ifelse(volcano_df$`-log10_Q` > 2 & volcano_df$log2FC > 0.5, 0.7,
#                             ifelse(volcano_df$`-log10_Q` > 2 & volcano_df$log2FC < -0.5, 0.7, 0.5)),
#              color = ifelse(volcano_df$`-log10_Q` > 2 & volcano_df$log2FC > 0.5, "#E20001",
#                             ifelse(volcano_df$`-log10_Q` > 2 & volcano_df$log2FC < -0.5, "#004996","grey"))) + 
#   geom_vline(xintercept = c(-0.5, 0.5)) +
#   geom_hline(yintercept = 2)+
#   theme_bw()
# 
# volcano_plot  <- FileCreate(DirPath = '../09.Bacteria/', Prefix = paste0('volcanoPlot-', 'all'), Suffix = 'pdf')
# pdf(file = volcano_plot, width = 5, height = 5)
# plot(g)
# dev.off()



volcano_df <- wil_res
volcano_df$`-log10_Q` <- -log10(volcano_df$adj_pvalue)
volcano_df$log2FC <- log2(volcano_df$mFC)
volcano_df$name <- rownames(volcano_df)
volcano_df2 <- volcano_df[volcano_df$`-log10_Q` >= 2 & abs(volcano_df$log2FC) >= 0.75, ]
g <- ggplot(volcano_df, aes(x = log2FC, y = `-log10_Q`)) +
  geom_point(alpha = ifelse(volcano_df$`-log10_Q` > 2 & volcano_df$log2FC > 0.5, 0.7,
                            ifelse(volcano_df$`-log10_Q` > 2 & volcano_df$log2FC < -0.5, 0.7, 0.5)),
             color = ifelse(volcano_df$`-log10_Q` > 2 & volcano_df$log2FC > 0.5, "#E20001",
                            ifelse(volcano_df$`-log10_Q` > 2 & volcano_df$log2FC < -0.5, "#004996",
                                   ifelse(volcano_df$`-log10_Q` < 2 & abs(volcano_df$log2FC) < 0.5 ,"grey", "darkgreen")))) + 
  geom_vline(xintercept = c(-0.5, 0.5)) +
  geom_hline(yintercept = 2)+
  geom_text_repel(data = volcano_df2,
                   mapping = aes(label = name, x = log2FC, y = `-log10_Q`),
                   color = ifelse(volcano_df2$`-log10_Q` > 2 &
                                    volcano_df2$log2FC > 0.75, "#E20001",
                            ifelse(volcano_df2$`-log10_Q` > 2 & 
                                     volcano_df2$log2FC < -0.75, "#004996",
                                   ifelse(volcano_df2$`-log10_Q` < 2 &
                                            abs(volcano_df2$log2FC) < 0.75 ,"grey",
                                          "darkgreen"))),
                  size = ifelse((abs(volcano_df2$log2FC) < 0.7 | volcano_df2$`-log10_Q` < 2), 3, 5),
                  max.overlaps = getOption("ggrepel.max.overlaps", default = 20),) + 
  theme_bw()

volcano_plot  <- FileCreate(DirPath = '../09.Bacteria/', Prefix = paste0('volcanoPlot-', 'all'), Suffix = 'pdf')
pdf(file = volcano_plot, width = 10, height = 10)
plot(g)
dev.off()

```

```{=html}
<!-- ## Step 3: Intersection of 2 criterion

Only focus on 0.1% abundance relative Eukaryota.

Totally **42** fungi leave after twice filters, and **6** of them are also general Fold change difference in all samples (`|log2(FC)| > 0.5`).

```{r}
sel_idx_1 <- rownames(wil_res)[Trend_Med_summary[rownames(wil_res), ]$N.count >=6 | Trend_Med_summary[rownames(wil_res), ]$N.count <= 2]
length(sel_idx_1) # 112

sel_idx_2 <- rownames(wil_res)[wil_res$adj_pvalue < 0.05]
length(sel_idx_2) # 59

sel_idx <- intersect(sel_idx_1, sel_idx_2)
length(sel_idx) # 42

sel_trend <- all_trend[sel_idx, ]

sel3_trend <- sel_trend[abs(sel_trend$`all-Median`) > 0.5,  ] 
sel3_trend <- sel3_trend[order(abs(sel3_trend$`all-Median`), decreasing = T), ]

```
-->
```
