---
title: "selected feature analysis"
author: "ifanlyn@outlook.com"
date: "2021/8/12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages and subroutines {.tabset}

### Package

```{r packages, message = F, warning = F}
require(DT) # for databale
require(reshape2) # for dcast
require(ggplot2)
require(ggthemes)
require(ggpubr) # for stat_compare_means
```

### Import table

```{r ImportTable }
ImportTable <- function(file, header = T, row.names = 1, sep = ',', check.names = FALSE, ...){
  data <- read.csv(file = file, header = header, row.names = row.names, sep = sep, check.names = check.names, ...)
  return(data)
}
```

### Show_table

```{r show_table}
Show_table <- function(df, rownames = T, filter="top", options = list(pageLength = 10, scrollX=T), ...){
  if (ncol(df) > 100) {
    df <- df[, 1:100]
    message('Due to the column dim is > 100, it only shows the top 100 columns')
  }
  datatable(df, rownames = rownames, 
            filter=filter, options = options, ... )

}
```


### FileCreate

```{r}
FileCreate <- function(DirPath = "./",Prefix = "ExampleTest", Suffix = "pdf", version = "0.0.0"){
  date=as.character(Sys.Date())
  DirPath = gsub("/$","",DirPath)
  Suffix = gsub('^[.]',"",Suffix)
  if(! dir.exists(DirPath)){
    dir.create(DirPath,recursive =TRUE)
  }
  if (version == "0.0.0") {
    version=100
    while(TRUE){
      version_=paste(unlist(strsplit(as.character(version),"")),collapse=".")
      DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
      if(! file.exists(DirPathName)){
        return(DirPathName)
        break
      }
      version = version + 1
    }
  }else{
    DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
    if(! dir.exists(DirPathName)){
      return(DirPathName)
    }
    return(DirPathName)
  }
}
  
```


## Import data {.tabset}

### Custom feature

```{r}
# custom_feature <- c('Aspergillus rambellii', 'Moniliophthora perniciosa', 
#                     'Trichophyton mentagrophytes', 'Aspergillus kawachii', 
#                     'Aspergillus ochraceoroseus')
custom_feature <- c('Saitoella complicata', 'Cryptococcus neoformans', 
                    "Fusarium pseudograminearum", "Lentinula edodes",
                    "Pseudogymnoascus sp. VKM F-3808", "Trichosporon asahii")


```



### Import modified-relative abundance matrix (remove zero)
```{r}
otu_mod_df <- ImportTable(file = '../07.FeatureSelection/01.SSTF/2021-07-29-modify_martix_norm-ALL-v1.0.0.csv')

Show_table(otu_mod_df)%>%
  formatSignif(columns = colnames(otu_mod_df)[1:100],
               digits = 3, interval = 1)

```

### Import median-relative abundance matrix
```{r}
otu_med_df <- ImportTable(file = '../07.FeatureSelection/01.SSTF/2021-07-29-modify_martix_norm-ALL-v1.0.0.csv')

Show_table(otu_med_df)%>%
  formatSignif(columns = colnames(otu_med_df)[1:100],
               digits = 3, interval = 1)

```



### Import meta data 
```{r}
meta_df2 <- ImportTable(file = '../00.RawData/metaData/2021-07-26-metaInfo-subgroup_1329-v6.0.csv')
# meta_df <- meta_df2
# meta_df$Stage <- factor(meta_df$Stage, c('CTRL', 'adenoma', 'CRC'))
meta_df <- meta_df2[meta_df2$Stage != 'CRC', ]
meta_df$Stage <- factor(meta_df$Stage, c('CTRL', 'adenoma'))
Show_table(meta_df)

```

### Import Bacteria median modified matrix

```{r}
bac_med_df <- ImportTable(file = '../09.Bacteria/01.modified_data/2021-08-05-modify_martix_norm-ALL-v1.0.0.csv')

comb_med_df <- cbind(otu_med_df, bac_med_df[rownames(otu_med_df),])

```

## Barplot in each cohort {.tabset}


### raw relative abundance matrix
```{r raw relative abundance, message = F, warning = F}
# combine_df <- as.data.frame(cbind(meta_df, otu_mod_df[, custom_feature]))
combine_df <- as.data.frame(cbind(meta_df, otu_mod_df[rownames(meta_df), custom_feature]))
combine_df <- combine_df[combine_df$Cohort %in% c("2014_ZellerG", "2015_FengQ", "2019_Thom", "2019_Yachida", "2021_JunY_2"), ]
g_list <- list()
boxplot_pdf <- FileCreate(DirPath = '../09.euk-ctrl_ade//SelFeatureAnalysi', 
                          Prefix = 'Boxplot-relMedAbun_norm',
                          Suffix = 'pdf')
pdf(file = boxplot_pdf, width = 8, height = 3)
for (cf in custom_feature) {
  g_bar <- ggplot(data = combine_df, mapping = aes(x = Stage, y = .data[[cf]])) +
    geom_violin(aes(fill = Stage), scale = 'width', trim = F)+
    geom_boxplot(outlier.colour = NA, width = 0.2)+
    facet_wrap(~Cohort, nrow = 1) + 
    theme_calc()+
    scale_fill_manual(values = c(CTRL = '#006f3c',
                                  adenoma = '#f9a73e'))+
    scale_y_log10()+
    stat_compare_means(comparisons = list(c('CTRL', 'adenoma')))+
     theme(axis.text.x=element_text(angle=90, hjust=1))
  g_list[[cf]] <- g_bar
  plot(g_bar)
}
dev.off()

for (cf in custom_feature) {
  plot(g_list[[cf]])
}

combine_df <- as.data.frame(cbind(meta_df, otu_mod_df[rownames(meta_df), custom_feature]))

boxplot_pdf2 <- FileCreate(DirPath = '../09.euk-ctrl_ade/SelFeatureAnalysi', 
                           Prefix = 'Boxplot-relMedAbun_norm_combine',
                           Suffix = 'pdf')
pdf(file = boxplot_pdf2, width = 4, height = 3)
for (cf in custom_feature) {
  g_bar <- ggplot(data = combine_df, mapping = aes(x = Stage, y = .data[[cf]])) +
    geom_violin(aes(fill = Stage), scale = 'width', trim = F)+
    geom_boxplot(outlier.colour = NA, width = 0.2)+
    theme_calc()+
    scale_fill_manual(values = c(CTRL = '#006f3c',
                                  adenoma = '#f9a73e'))+
    scale_y_log10()+
    stat_compare_means(comparisons = list(c('CTRL', 'adenoma')))+
     theme(axis.text.x=element_text(angle=90, hjust=1))
  g_list[[cf]] <- g_bar
  plot(g_bar)
}
dev.off()

for (cf in custom_feature) {
  plot(g_list[[cf]])
}


# combine_df <- as.data.frame(cbind(meta_df2, otu_mod_df[rownames(meta_df2), overlap_feature]))
# combine_df$Stage <- factor(x = combine_df$Stage, levels = c("CTRL", "adenoma", "CRC"))
# 
# boxplot_pdf2 <- FileCreate(DirPath = '../07.FeatureSelection/', 
#                            Prefix = 'Overlap-Boxplot-relMedAbun_norm_combine',
#                            Suffix = 'pdf')
# pdf(file = boxplot_pdf2, width = 6, height = 5)
# for (cf in overlap_feature) {
#   g_bar <- ggplot(data = combine_df, mapping = aes(x = Stage, y = .data[[cf]])) +
#     geom_violin(aes(fill = Stage), scale = 'width', trim = F)+
#     geom_boxplot(outlier.colour = NA, width = 0.2)+
#     theme_calc()+
#     scale_fill_manual(values = c(CTRL = '#006f3c',
#                                   adenoma = '#f9a73e', 
#                                   CRC = '#bf212f'))+
#     scale_y_log10()+
#     stat_compare_means(comparisons = list(c('CTRL', 'CRC'), c('adenoma', 'CRC')))+
#      theme(axis.text.x=element_text(angle=90, hjust=1))
#   g_list[[cf]] <- g_bar
#   plot(g_bar)
# }
# dev.off()
```


### zero modified matrix
```{r zero modified matrix, message = F, warning = F}
# combine_df <- as.data.frame(cbind(meta_df, otu_mod_df[, custom_feature]))
combine_df <- as.data.frame(cbind(meta_df, otu_mod_df[rownames(meta_df), custom_feature]))
g_list <- list()
boxplot_pdf <- FileCreate(DirPath = '../10.SelFeatureAnalysi', 
                          Prefix = 'Boxplot-modified_zero_norm',
                          Suffix = 'pdf')
pdf(file = boxplot_pdf, width = 13, height = 5)
for (cf in custom_feature) {
  g_bar <- ggplot(data = combine_df, mapping = aes(x = Stage, y = .data[[cf]])) +
    geom_violin(aes(fill = Stage), scale = 'width', trim = F)+
    geom_boxplot(outlier.colour = NA, width = 0.2)+
    # geom_boxplot(aes(col = Stage), outlier.colour = NA)+
    facet_wrap(~Cohort, nrow = 1) + 
    theme_calc()+
    scale_fill_manual(values = c(CTRL = '#006f3c',
    # scale_color_manual(values = c(CTRL = '#006f3c', 
                                  adenoma = '#f9a73e', 
                                  CRC = '#bf212f'))+
    scale_y_log10()+
    # scale_y_log10(limits = c(min(combine_df[[cf]])/10, max(combine_df[[cf]])*500))+
    stat_compare_means(comparisons = list(c('CTRL', 'CRC')
                                        # , c('adenoma', 'CRC')
                                          ))+
     theme(axis.text.x=element_text(angle=90, hjust=1))
  plot(g_bar)
  g_list[[cf]] <- g_bar
}
dev.off()

for (cf in custom_feature) {
  plot(g_list[[cf]])
}
```

### median modified matrix
```{r median modified matrix, message = F, warning = F}
# combine_df <- as.data.frame(cbind(meta_df, otu_mod_df[, custom_feature]))
combine_df <- as.data.frame(cbind(meta_df, otu_med_df[rownames(meta_df), custom_feature]))
g_list <- list()
boxplot_pdf <- FileCreate(DirPath = '../10.SelFeatureAnalysi', 
                          Prefix = 'Boxplot-median_norm',
                          Suffix = 'pdf')
pdf(file = boxplot_pdf, width = 13, height = 5)
for (cf in custom_feature) {
  g_bar <- ggplot(data = combine_df, mapping = aes(x = Stage, y = .data[[cf]])) +
    geom_violin(aes(fill = Stage), scale = 'width', trim = F)+
    geom_boxplot(outlier.colour = NA, width = 0.2)+
    # geom_boxplot(aes(col = Stage), outlier.colour = NA)+
    facet_wrap(~Cohort, nrow = 1) + 
    theme_calc()+
    scale_fill_manual(values = c(CTRL = '#006f3c',
    # scale_color_manual(values = c(CTRL = '#006f3c', 
                                  adenoma = '#f9a73e', 
                                  CRC = '#bf212f'))+
    scale_y_log10()+
    # scale_y_log10(limits = c(min(combine_df[[cf]])/10, max(combine_df[[cf]])*500))+
    stat_compare_means(comparisons = list(c('CTRL', 'CRC')
                                        # , c('adenoma', 'CRC')
                                          ))+
     theme(axis.text.x=element_text(angle=90, hjust=1))
  plot(g_bar)
  g_list[[cf]] <- g_bar
}
dev.off()

for (cf in custom_feature) {
  plot(g_list[[cf]])
}
```


## Selected the median value as cutoff

```{r}
i = 1

tmp_df <- as.data.frame(cbind(meta_df, otu_med_df[rownames(meta_df), custom_feature]))

lower_limit <- min(tmp_df[[custom_feature[i]]])
upper_limit <- max(tmp_df[[custom_feature[i]]])
ctrl_density <- density(subset(tmp_df, Stage == 'CTRL')[[custom_feature[i]]],
                        from = lower_limit, to = upper_limit, n = 2^15)
crc_density <- density(subset(tmp_df, Stage == 'CRC')[[custom_feature[i]]],
                        from = lower_limit, to = upper_limit, n = 2^15)

density_difference <- crc_density$y - ctrl_density$y
intersection_point <- crc_density$x[which(density_difference > 0) + 1][1]

g <- ggplot(tmp_df, aes(x = .data[[custom_feature[i]]])) +
  # geom_histogram(bins = 100)+
  scale_x_log10() +
  # xlim(0,10)+
  geom_density(color = 'grey', linetype = 2)+
  geom_density(aes(color = Stage))+
  # geom_line(aes(x = crc_density$x, y = crc_density$y), color = '#bf212f')+
  # geom_line(aes(x = ctrl_density$x, y = ctrl_density$y), color = '#006f3c')+
  geom_vline(xintercept = intersection_point[1],
             linetype = 4) +
  theme_calc() +
  scale_color_manual(values = c(CTRL = '#006f3c', CRC = '#bf212f'))+
  geom_text(x = log10(intersection_point[1]*3), y = 0.62, 
            label = paste0('x = ', signif(intersection_point,4)))

plot(g)

density_pdf <- FileCreate(DirPath = '../10.SelFeatureAnalysi/density',
                          Prefix = paste0('Density-',
                                          strsplit(custom_feature[i], " ") %>% 
                                            unlist() %>% paste0(., collapse = "_")),
                          Suffix = 'pdf')
pdf(file = density_pdf, width = 7, height = 3)
plot(g)
dev.off()


AR_high <- rownames(tmp_df)[tmp_df$`Aspergillus rambellii` >= intersection_point]
AR_low <- rownames(tmp_df)[tmp_df$`Aspergillus rambellii` < intersection_point]

AR_high_df <- comb_med_df[AR_high, ]

AR_high_ctrl_df <- AR_high_df[meta_df[AR_high, ]$Stage == "CTRL", ]
AR_high_crc_df <- AR_high_df[meta_df[AR_high, ]$Stage == "CRC", ]


AR_low_df <- comb_med_df[AR_low, ]

AR_low_ctrl_df <- AR_low_df[meta_df[AR_low, ]$Stage == "CTRL", ]
AR_low_crc_df <- AR_low_df[meta_df[AR_low, ]$Stage == "CRC", ]



cor.test(comb_med_df[rownames(tmp_df), ]$`Aspergillus rambellii`, 
         comb_med_df[rownames(tmp_df), ]$`Fusobacterium nucleatum`, 
         method = 'spearman', exact = FALSE)


cor.test(AR_high_df$`Aspergillus rambellii`, AR_high_df$`Fusobacterium nucleatum`, 
         method = 'spearman', exact = FALSE)
cor.test(AR_low_df$`Aspergillus rambellii`, AR_low_df$`Fusobacterium nucleatum`,
         method = 'spearman', exact = FALSE)

cor.test(AR_high_ctrl_df$`Aspergillus rambellii`, AR_high_ctrl_df$`Fusobacterium nucleatum`, 
         method = 'spearman', exact = FALSE)
cor.test(AR_high_crc_df$`Aspergillus rambellii`, AR_high_crc_df$`Fusobacterium nucleatum`,
         method = 'spearman', exact = FALSE)

cor.test(AR_low_ctrl_df$`Aspergillus rambellii`, AR_low_ctrl_df$`Fusobacterium nucleatum`, 
         method = 'spearman', exact = FALSE)
cor.test(AR_low_crc_df$`Aspergillus rambellii`, AR_low_crc_df$`Fusobacterium nucleatum`,
         method = 'spearman', exact = FALSE)


```

















