---
title: "RandomForest"
author: "ifanlyn@outlook.com"
date: "2021/9/28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages and subroutines {.tabset}

### Package

```{r packages, message = F, warning = F}
require(DT) # for databale
require(randomForest) # randomForest
require(pROC)
require(ROCR)
require(mlbench)
require(caret)
require(ComplexHeatmap)
require(circlize)
```


### Import table

```{r ImportTable , message = F, warning = F}
ImportTable <- function(file, header = T, row.names = 1, sep = ',', check.names = FALSE, ...){
  data <- read.csv(file = file, header = header, row.names = row.names, sep = sep, check.names = check.names, ...)
  return(data)
}
```

### Show_table

```{r show_table, message = F, warning = F}
Show_table <- function(df, rownames = T, filter="top", options = list(pageLength = 10, scrollX=T), ...){
  if (ncol(df) > 100) {
    df <- df[, 1:100]
    message('Due to the column dim is > 100, it only shows the top 100 columns')
  }
  datatable(df, rownames = rownames, 
            filter=filter, options = options, ... )

}
```

### FileCreate

```{r , message = F, warning = F}
FileCreate <- function(DirPath = "./",Prefix = "ExampleTest", Suffix = "pdf", version = "0.0.0"){
  date=as.character(Sys.Date())
  DirPath = gsub("/$","",DirPath)
  Suffix = gsub('^[.]',"",Suffix)
  if(! dir.exists(DirPath)){
    dir.create(DirPath,recursive =TRUE)
  }
  if (version == "0.0.0") {
    version=100
    while(TRUE){
      version_=paste(unlist(strsplit(as.character(version),"")),collapse=".")
      DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
      if(! file.exists(DirPathName)){
        return(DirPathName)
        break
      }
      version = version + 1
    }
  }else{
    DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
    if(! dir.exists(DirPathName)){
      return(DirPathName)
    }
    return(DirPathName)
  }
}
  
```

## Import data {.tabset}

### metadata

```{r , message = F, warning = F}
meta_df <- ImportTable(file = '../00.RawData/metaData/2021-07-26-metaInfo-subgroup_1329-v6.0.csv')

```

### Import Eukaryota matrix

```{r , message = F, warning = F}

euk_otu <- ImportTable(file = '../07.FeatureSelection/02.RelMedianAbundance/2021-07-30-normalized-0.1%-RelMedianAbundance-matrix-v1.0.0.csv')
euk_sel <- ImportTable(file = '../07.FeatureSelection/01.SSTF/2021-07-30-wilcoxon-all-v1.0.0.csv')
euk_sel <- euk_sel[order(euk_sel$adj_pvalue),]
euk_core <- euk_otu[, rownames(euk_sel)[euk_sel$adj_pvalue < 0.01]]
euk_sel_list <- colnames(euk_core)

```

### Import Bacteria matrix

```{r , message = F, warning = F}

bac_otu <- ImportTable(file = '../09.Bacteria/02.RelMedianAbundance/2021-08-05-normalized-0.1%-RelMedianAbundance-matrix-v1.0.0.csv')
bac_sel <- ImportTable(file = '../09.Bacteria/2021-08-06-wilcoxon-all-v1.0.0.csv')
bac_sel <- bac_sel[order(bac_sel$adj_pvalue),]
bac_core_all <- bac_otu[, rownames(bac_sel)[bac_sel$adj_pvalue < 0.01 & bac_sel$AbsLog2FC > 0.5]]
bac_sel_list_all <- colnames(bac_core_all)
bac_sel_list <- bac_sel_list_all[!grepl('sp. ', bac_sel_list_all)]
bac_core <- bac_core_all[, bac_sel_list]
```

### Combine the core sets and stages

```{r , message = F, warning = F}
comb_core <- as.data.frame(cbind(Stage = meta_df$Stage, cbind(euk_core, bac_core)))
colnames(comb_core) <- gsub("\\W+", '_', colnames(comb_core))
colnames(comb_core) <- gsub("^_", '', colnames(comb_core))
cn_combin_core <- comb_core[comb_core$Stage!= 'adenoma', ]
cn_combin_core$Stage <- factor(x = cn_combin_core$Stage, levels = c('CRC', 'CTRL'))
cn_meta_df <- meta_df[meta_df$Stage != 'adenoma', ]
head(sapply(cn_combin_core, class))
```


## Random Forest LOSO

random forest, leave one set out.



### Rank Features By Importance

[link](https://stats.stackexchange.com/questions/443234/random-forest-in-r-how-to-perform-feature-extraction-and-reach-the-best-accura)
```{r, message = F, warning = F}
require(dplyr)
require(caret)
require(randomForest)
require(eegkit)
require(rlist)
require(edfReader)
require(eegUtils)
require(e1071)
require(ggplot2)
# sub_type <- 'fungi'
for (sub_type in c('fungi', 'bac', 'all')) {
  if (sub_type == 'fungi') {
    n_feature <- 33
    myData <- cn_combin_core[, 1:(1 + n_feature)]
  }else if (sub_type == 'bac') {
    n_feature <- 31
    myData <- cn_combin_core[, c(1, 35: (34 + n_feature))]
  }else{
    n_feature <- 64
    myData <- cn_combin_core
  }
  
  important_score <- 50
  set.seed(123)
  ###############################Random Forest
  Importance_list <- matrix(data = NA, nrow = 1, ncol = n_feature)
  Importance_list <- Importance_list[-1, ]
  for (c in cohort_lst){
    Importance_Table <- matrix(data = NA, nrow = 1, ncol = n_feature )
    Importance_Table <- Importance_Table[-1, ]
    for (t in 1:10) {
      seed <- sample.int(10)
      set.seed(seed)  
      seeds <- vector(mode = "list", length = 50)
      for(i in 1:50){
        seeds[[i]] <- sample.int(1000, 12)
      }
      
      ## For the last model:
      seeds[[50]] <- sample.int(1000, 1)
    
      myData$Stage <- as.factor(myData$Stage)
      
      training <- myData[rownames(cn_meta_df)[cn_meta_df$Cohort == c],]
      testing <- myData[rownames(cn_meta_df)[cn_meta_df$Cohort != c], ] 
      
      train_control <- trainControl(method="cv", number=10, verboseIter = TRUE, seeds = seeds) 
      
      model <- train(training[,2:ncol(training)],
                     as.factor(training[,1]), 
                     method = "rf", type="classification",
                     metric= "Accuracy", maximize= TRUE, 
                     trControl = train_control, importance = TRUE) 
      
      
      Adding_columns <- NULL
      varImp2 <- varImp(model, scale = TRUE)
      Adding_columns <- t(varImp2$importance)
      rownames(Adding_columns) <- paste0(rownames(Adding_columns),".", t)
      Importance_Table <- rbind(Importance_Table, Adding_columns[1, ,drop = F])
    }
    Importance_Table_csv <- FileCreate(DirPath = paste0('../11.Classifier/RF/FeatureSel_Important/', sub_type), 
                                       Prefix = paste0('Importance_Table_', c), Suffix = 'csv')
    Importance_Table_Mean <- t(apply(Importance_Table, MARGIN = 2, function(x) mean(x, na.rm=TRUE)))
    rownames(Importance_Table_Mean) <- paste0(c, "_Mean")
    Importance_list <- rbind(Importance_list, Importance_Table_Mean)
    Importance_Table_2 <- as.data.frame(Importance_Table)
    Importance_Table_2 <- rbind(Mean=Importance_Table_Mean, Importance_Table_2)
    write.csv(x = Importance_Table_2, file = Importance_Table_csv)
  
    Importance_Table_Filter <- as.data.frame(Importance_Table_Mean)
    Importance_Table_Filter2 <- Importance_Table_Filter[,Importance_Table_Filter< important_score]
    Importance_Table_Filter3 <- colnames(Importance_Table_Filter2) 
    Excluding_Channels <- names(myData) %in% Importance_Table_Filter3
    myData_filter <- myData[!Excluding_Channels]
    Importance_filter_csv <- FileCreate(DirPath = paste0('../11.Classifier/RF/FeatureSel_Important/', sub_type), 
                                       Prefix = paste0('Importance_feature_', c, "_", important_score), Suffix = 'csv')
    write.csv(x = Importance_Table_Filter[,Importance_Table_Filter>= important_score], file = Importance_filter_csv)

    
  }
  
  Importance_df2 <- as.data.frame(t(Importance_list))
  for (c in cohort_lst) {
    Importance_df2[[paste0(c, "_Rank")]] <- rank(-Importance_df2[[paste0(c, "_Mean")]])
  }
  rank_list <- grep('_Rank', colnames(Importance_df2))
  Importance_df2$Rank3 <- apply(Importance_df2[, rank_list] , 1, function(x){ifelse(any(x <= 3), 'Sel', "Ign")})
  Importance_df2$Rank5 <- apply(Importance_df2[, rank_list] , 1, function(x){ifelse(any(x <= 5), 'Sel', "Ign")})
  Importance_df2$Rank10 <- apply(Importance_df2[, rank_list] , 1, function(x){ifelse(any(x <= 10), 'Sel', "Ign")})
  
  Importance_df2_csv <-  FileCreate(DirPath = paste0('../11.Classifier/RF/FeatureSel_Important/', sub_type), 
                                     Prefix = 'Combine_importantScore_rank_all', Suffix = 'csv')
  write.csv(Importance_df2, Importance_df2_csv)
  
  
  

    
  
    # Importance_Table_Filter <- as.data.frame(Importance_Table_Mean)
    # Importance_Table_Filter2 <- Importance_Table_Filter[,Importance_Table_Filter< important_score]
    # Importance_Table_Filter3 <- colnames(Importance_Table_Filter2) 
    # Excluding_Channels <- names(myData) %in% Importance_Table_Filter3
    # myData_filter <- myData[!Excluding_Channels]
    # myData_filter <- as.data.frame(t(myData_filter))
    # Importance_filter_csv <- FileCreate(DirPath = paste0('../11.Classifier/RF/FeatureSel_Important/'), 
    #                                    Prefix = paste0('Importance_feature_all_', important_score), Suffix = 'csv')
    # write.csv(x = Importance_Table_Filter[,Importance_Table_Filter>= important_score], file = Importance_filter_csv)

    
  # }
  

}

###############################Random Forest
# for (sub_type in c("all", "bac", "fungi")) {

for (sub_type in c("fungi")) {
  if (sub_type == 'all') {
    tmp_column = 1:65
    n_feature = 64
  }else if(sub_type == 'bac'){
    tmp_column = c(1, 35:65)
    n_feature = 31
  }else{
    tmp_column = 1:34
    n_feature = 33
  }
  Importance_list <- matrix(data = NA, nrow = 1, ncol = n_feature)
  Importance_list <- Importance_list[-1, ]
  
  set.seed(123)
  Importance_Table <- matrix(data = NA, nrow = 1, ncol = n_feature )
  Importance_Table <- Importance_Table[-1, ]
  for (t in 1:10) {
    seed <- sample.int(10)
    set.seed(seed)  
    seeds <- vector(mode = "list", length = 50)
    for(i in 1:50){
      seeds[[i]] <- sample.int(1000, 12)
    }
    
    ## For the last model:
    seeds[[50]] <- sample.int(1000, 1)
  
    myData$Stage <- as.factor(myData$Stage)
    
    training <- myData[, tmp_column]
    
    train_control <- trainControl(method="cv", number=10, verboseIter = TRUE, seeds = seeds) 
    
    model <- train(training[,2:ncol(training)],
                   as.factor(training[,1]), 
                   method = "rf", type="classification",
                   metric= "Accuracy", maximize= TRUE, 
                   trControl = train_control, importance = TRUE) 
    
    
    Adding_columns <- NULL
    varImp2 <- varImp(model, scale = TRUE)
    Adding_columns <- t(varImp2$importance)
    rownames(Adding_columns) <- paste0(rownames(Adding_columns),".", t)
    Importance_Table <- rbind(Importance_Table, Adding_columns[1, ,drop = F])
  }
  Importance_Table_csv <- FileCreate(DirPath = paste0('../11.Classifier/RF/FeatureSel_Important/'), 
                                     Prefix = paste0('Importance_Table_',sub_type), Suffix = 'csv')
  Importance_Table_Mean <- t(apply(Importance_Table, MARGIN = 2, function(x) mean(x, na.rm=TRUE)))
  rownames(Importance_Table_Mean) <- "all_Mean"
  Importance_list <- rbind(Importance_list, Importance_Table_Mean)
  Importance_Table_2 <- as.data.frame(Importance_Table)
  Importance_Table_2 <- rbind(Mean=Importance_Table_Mean, Importance_Table_2)
  Importance_Table_3 <- as.data.frame(t(Importance_Table_2))
  Importance_Table_3$Rank <- rank(-Importance_Table_3$Mean)
  write.csv(x = Importance_Table_3, file = Importance_Table_csv)
}
```

### Train model
```{r, message = F, warning = F}
for (sub_type in c('fungi', 'bac', 'all')) {
# for (sub_type in c('bac', 'all')) {
  Importance_df2 <- ImportTable(file = paste0('../11.Classifier/RF/FeatureSel_Important/',
                                              sub_type, '/2021-09-29-Combine_importantScore_rank_all-v1.0.0.csv'), 
                                sep = ',')
  sel_all_list <- rownames(Importance_df2)[Importance_df2$Rank3 == 'Sel']
  sel_imp_df <- cn_combin_core[, c('Stage', sel_all_list)]
  
  auc_df <- as.data.frame(matrix(data = NA, nrow = 100, ncol = 9))
  rownames(auc_df) <- paste0("Seed_", 1:100)
  colnames(auc_df) <- c(cohort_lst, "Mean")
  auc_mean <- 0
  auc_seed <- 1
  for (k in c(1:100)) {
    auc_tmp <- NULL
    for (c in cohort_lst) {
      train_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort != c], ]
      valda_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort == c], ]
      set.seed(k)
      # set.seed(123)
      rf <- randomForest(Stage ~ ., data = train_set)
      pred <- predict(rf, newdata=valda_set[-1], type = 'prob')[,2]
      pred_result <- prediction(pred, valda_set[,1])
      auc_value <- round(performance(pred_result, measure = "auc")@y.values[[1]]*100, 2)
      # message(c, "\t", auc_value, "%")
      auc_tmp <- c(auc_tmp, auc_value)
    }
    auc_df[paste0("Seed_", k),] <- c(auc_tmp, mean(auc_tmp))
    if (mean(auc_tmp) > auc_mean) {
      message(sub_type, "\t", k, "\t", round(mean(auc_tmp), 2), "%")
    }
    auc_seed <- ifelse(mean(auc_tmp) > auc_mean, k, auc_seed)
    auc_mean <- ifelse(mean(auc_tmp) > auc_mean, mean(auc_tmp), auc_mean)
    
  }
  auc_df_csv <- FileCreate(DirPath = paste0('../11.Classifier/RF/AUC_res/'), 
                           Prefix = paste0('AUC_matrix-', sub_type), Suffix = 'csv')
  write.csv(auc_df, auc_df_csv)
}

```




```{r, message = F, warning = F}
auc_comp <- matrix(data = NA, nrow = 1, ncol = 2)
auc_comp <- auc_comp[-1, ]
for (sub_type in c('fungi', 'bac', 'all')) {
  auc_df <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-',
                                              sub_type, '-v1.0.0.csv'))
  auc_comp <- rbind(auc_comp, cbind(auc_df, Group = rep(sub_type, 100)))
}

wilcox.test(Mean ~ Group, data = auc_comp, subset = Group %in% c('all', 'bac'))
wilcox.test(Mean ~ Group, data = auc_comp, subset = Group %in% c('all', 'fungi'))
wilcox.test(Mean ~ Group, data = auc_comp, subset = Group %in% c('fungi', 'bac'))
```

## Plot Heatmap

```{r, message = F, warning = F}
Importance_df2 <- ImportTable(file = '../11.Classifier/RF/FeatureSel_Important/all/2021-09-29-Combine_importantScore_rank_all-v1.0.0.csv')
subset_imp <- Importance_df2[Importance_df2$Rank3 == 'Sel', ]
subset_imp2 <- subset_imp[, grep('_Rank', colnames(subset_imp))]

fungi_median <- ImportTable(file = '../07.FeatureSelection/01.SSTF/2021-07-30-walsh-median-all-v1.0.0.csv')
rownames(fungi_median) <- gsub("\\W+", '_', rownames(fungi_median))
rownames(fungi_median) <- gsub("^_", '', rownames(fungi_median))
fungi_median$EnrDep <- ifelse(fungi_median$`all-Median` >0, 'Enriched', 'Depleted')
fungi_median$Kingdoms <- 'fungi'

bac_median <- ImportTable(file = '../09.Bacteria/2021-08-06-wilcoxon-all-v1.0.0.csv')
rownames(bac_median) <- gsub("\\W+", '_', rownames(bac_median))
rownames(bac_median) <- gsub("^_", '', rownames(bac_median))
bac_median$EnrDep <- ifelse(bac_median$mFC >1, 'Enriched', 'Depleted')
bac_median$Kingdoms <- 'bac'

combine_EnrDep <- rbind(fungi_median[, c(5, 6), drop = F], bac_median[, c(5, 6), drop = F])
combine_EnrDep <- combine_EnrDep[rownames(subset_imp), , drop = F]

comb_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-all-v1.0.0.csv'))
bac_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-bac-v1.0.0.csv'))

summary_AUC <- rbind(Comb_AUC = comb_AUC[which.max(comb_AUC$Mean),], Bac_AUC = bac_AUC[which.max(bac_AUC$Mean),])
summary_AUC <- rbind(summary_AUC, Ratio = (summary_AUC[1,]/summary_AUC[2,]-1)*100)

all_Feature_imp <- ImportTable(file ="../11.Classifier/RF/FeatureSel_Important/2021-09-30-Importance_Table_all-v1.0.0.csv")
all_Feature_imp1 <- all_Feature_imp[rownames(subset_imp), "Rank", drop = F]
all_Feature_imp2 <- cbind(all_Feature_imp1, combine_EnrDep)
all_Feature_imp2 <- all_Feature_imp2[order(all_Feature_imp2$Rank,all_Feature_imp2$EnrDep), ]


## plot enriched sub
enrich_mt <- subset_imp2[rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched'],]
enrich_col <- colorRamp2(c(0, 30), c('#bc0025', '#ffffcc'))

enrich_all_list <- all_Feature_imp2$Rank[all_Feature_imp2$EnrDep == 'Enriched']
names(enrich_all_list) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched']

enrich_all_list_label <- as.character(enrich_all_list)
names(enrich_all_list_label) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched']

anno_col_enrich <- enrich_col(enrich_all_list)
names(anno_col_enrich) <- enrich_all_list
enrich_label_col <- lapply(enrich_all_list, function(x){ifelse(x<=3, "white", "black")}) %>% unlist

enrich_rowname_col <- lapply(all_Feature_imp2$Kingdoms[all_Feature_imp2$EnrDep == 'Enriched'],
                             function(x){ifelse(x == 'fungi', "#ff410d", "#014586")})%>% unlist

enrich_row_anno <- rowAnnotation(all = anno_simple(enrich_all_list, 
                                                   pch = enrich_all_list_label,
                                                   col = anno_col_enrich,
                                                   gp = gpar(col = "black", lwd = 2),
                                                   simple_anno_size = unit(1, "cm"),
                                                   pt_gp = gpar(fontsize = 12, fontface = 'bold',
                                                                col = enrich_label_col),
                                                   pt_size = unit(5, "mm")), annotation_label = c(""))
h1 <- Heatmap(enrich_mt, col = enrich_col, name = "CRC-enriched\nSpecies",
              cluster_columns = F, cluster_rows = F, border = T, 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(enrich_mt[i, j], x, y, gp = gpar(fontsize = 12, fontface = 'bold', 
                                    col = ifelse(enrich_mt[i, j]<=3,'white', 'black')))},
              width = ncol(enrich_mt)*unit(10, "mm"), 
              height = nrow(enrich_mt)*unit(10, "mm"),
              right_annotation = enrich_row_anno,
              row_names_gp = gpar(col = enrich_rowname_col, fontface = 4))

h1




## plot depleted sub

deplet_mt <- subset_imp2[rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted'],]
deplet_col <- colorRamp2(c(0, 30), c('#243493', '#ffffcc'))

deplet_all_list <- all_Feature_imp2$Rank[all_Feature_imp2$EnrDep == 'Depleted']
names(deplet_all_list) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted']

deplet_all_list_label <- as.character(deplet_all_list)
names(deplet_all_list_label) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted']

anno_col_deplet <- deplet_col(deplet_all_list)
names(anno_col_deplet) <- deplet_all_list
deplet_label_col <- lapply(deplet_all_list, function(x){ifelse(x<=3, "white", "black")}) %>% unlist

deplet_rowname_col <- lapply(all_Feature_imp2$Kingdoms[all_Feature_imp2$EnrDep == 'Depleted'],
                             function(x){ifelse(x == 'fungi', "#ff410d", "#014586")})%>% unlist

deplet_row_anno <- rowAnnotation(all = anno_simple(deplet_all_list, 
                                                   pch = deplet_all_list_label,
                                                   col = anno_col_deplet,
                                                   gp = gpar(col = "black", lwd = 2),
                                                   simple_anno_size = unit(1, "cm"),
                                                   pt_gp = gpar(fontsize = 12, fontface = 'bold',
                                                                col = deplet_label_col),
                                                   pt_size = unit(5, "mm")))

h2 <- Heatmap(deplet_mt, col = deplet_col, name = "CRC-depleted\nSpecies",
              cluster_columns = F, cluster_rows = F, border = T, 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(deplet_mt[i, j], x, y, 
                          gp = gpar(fontsize = 12, fontface = 'bold', 
                                    col = ifelse(deplet_mt[i, j]<=3,'white', 'black')))},
              width = ncol(deplet_mt)*unit(10, "mm"), 
              height = nrow(deplet_mt)*unit(10, "mm"), 
              right_annotation = deplet_row_anno,
              row_names_gp = gpar(col = deplet_rowname_col, fontface = 4))
h2

AUC_col <- colorRamp2(c(65, 70,75,80,85,90), c('snow', 'khaki1', 'yellow', 'red', 'red3', 'red4'))

h3 <- Heatmap(summary_AUC[1, c(1:8), drop =F],
              cluster_columns = F, col = AUC_col, 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(round(summary_AUC[1, j],2), x, y, 
                          gp = gpar(fontsize = 8, fontface = 'bold', 
                                    col = ifelse(summary_AUC[1, j] > 80, "white",'black')))},
              height = unit(10, "mm"))

h4 <- Heatmap(summary_AUC[2, c(1:8), drop =F],
              cluster_columns = F, col = AUC_col, 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(round(summary_AUC[2, j],2), x, y, 
                          gp = gpar(fontsize = 8, fontface = 'bold', 
                                    col = ifelse(summary_AUC[2, j] > 80, "white",'black')))},
              height = unit(10, "mm"))


Ratio_col <- colorRamp2(c(-1, 0, 5, 10), c('darkblue', 'white', 'violetred', '#810f7b'))

h5 <- Heatmap(summary_AUC[3, c(1:8), drop =F],
              cluster_columns = F, col = Ratio_col, 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(round(abs(summary_AUC[3, j]),2), x, y, 
                          gp = gpar(fontsize = 8, fontface = 'bold', 
                                    col = ifelse(abs(summary_AUC[3, j]) > 5, "white",'black')))},
              height = unit(10, "mm"))

ht_list <- h1 %v% h2 %v% h3 %v% h4 %v% h5

ht_pdf <- FileCreate(DirPath = '../11.Classifier/RF/Plot', Prefix = 'heatmap_RF_AUC', Suffix = 'pdf')

pdf(file = ht_pdf, width = 15, height = 15)
draw(ht_list)
dev.off()

```




## Plot AUC

```{r, message = F, warning = F}

comb_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-all-v1.0.0.csv'))
bac_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-bac-v1.0.0.csv'))
fungi_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-fungi-v1.0.0.csv'))

best_seed_list <- list()
best_seed_list[['all']] <- which.max(comb_AUC$Mean)
best_seed_list[['bac']] <- which.max(bac_AUC$Mean)
best_seed_list[['fungi']] <- which.max(fungi_AUC$Mean)


roc_comp_df <- matrix(data = NA, nrow = 1, ncol = 2)
roc_comp_df <- roc_comp_df[-1,]
for (c in cohort_lst) {
  roc_list <- list()
  for (sub_type in c('fungi', 'bac', 'all')) {
    Importance_df2 <- ImportTable(file = paste0('../11.Classifier/RF/FeatureSel_Important/',
                                                sub_type, '/2021-09-29-Combine_importantScore_rank_all-v1.0.0.csv'), 
                                  sep = ',')
    sel_all_list <- rownames(Importance_df2)[Importance_df2$Rank3 == 'Sel']
    sel_imp_df <- cn_combin_core[, c('Stage', sel_all_list)]
    train_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort != c], ]
    valda_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort == c], ]
    set.seed(best_seed_list[[sub_type]])
    # set.seed(123)
    rf <- randomForest(Stage ~ ., data = train_set)
    pred <- predict(rf, newdata=valda_set[-1], type = 'prob')[,2]
    pred_result <- prediction(pred, valda_set[,1])
    auc_value <- round(performance(pred_result, measure = "auc")@y.values[[1]]*100, 2)
    
    roc_list[[sub_type]] <- roc(valda_set[,1], pred, legacy.axes = T, percent = T)
  }
  roc_pdf <- FileCreate(DirPath = "../11.Classifier/RF/Comparison/", Prefix = paste0('ROC-', c), Suffix = 'pdf')
  pdf(file = roc_pdf)
  par(pty = "s")
  plot(roc_list$all, col = '#ffd320', lwd = 4, legacy.axes = T, 
       xlab = "False Postive Percentage", ylab = "True Postive Percentage")
  # lines(roc_list$fungi, col = '#ff420e', lwd = 2, lty = 1)
  lines(roc_list$bac, col = '#004586', lwd = 2, lty = 1)
  dev.off()
  roc_comp <- roc.test(roc_list$all, roc_list$bac)
  roc_comp_df <- rbind(roc_comp_df, c(roc_comp$p.value, roc_comp$statistic))
  
}
roc_comp_df <- as.data.frame(roc_comp_df)
rownames(roc_comp_df) <- cohort_lst
colnames(roc_comp_df) <- c("p-value", "Z")

roc_comp_csv <- FileCreate(DirPath = "../11.Classifier/RF/Comparison", Prefix = "ROC_Comparison", Suffix = 'csv')
write.csv(roc_comp_df, roc_comp_csv)

#ff420e   fungi
#004586   bac
#ffd320   combine


```


## Supplementary Heatmap

### Bac Heatmap

```{r, message = F, warning = F}
Importance_df2 <- ImportTable(file = '../11.Classifier/RF/FeatureSel_Important/bac/2021-09-29-Combine_importantScore_rank_all-v1.0.0.csv')
subset_imp <- Importance_df2[Importance_df2$Rank3 == 'Sel', ]
subset_imp2 <- subset_imp[, grep('_Rank', colnames(subset_imp))]


bac_median <- ImportTable(file = '../09.Bacteria/2021-08-06-wilcoxon-all-v1.0.0.csv')
rownames(bac_median) <- gsub("\\W+", '_', rownames(bac_median))
rownames(bac_median) <- gsub("^_", '', rownames(bac_median))
bac_median$EnrDep <- ifelse(bac_median$mFC >1, 'Enriched', 'Depleted')
bac_median$Kingdoms <- 'bac'

subset_imp2 <- subset_imp2[rownames(subset_imp2) %in% rownames(bac_median), ]


combine_EnrDep <- bac_median[, c(5, 6), drop = F]
combine_EnrDep <- combine_EnrDep[rownames(subset_imp2), , drop = F]

comb_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-all-v1.0.0.csv'))
bac_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-bac-v1.0.0.csv'))

summary_AUC <- rbind(Comb_AUC = comb_AUC[which.max(comb_AUC$Mean),], Bac_AUC = bac_AUC[which.max(bac_AUC$Mean),])
summary_AUC <- rbind(summary_AUC, Ratio = (summary_AUC[1,]/summary_AUC[2,]-1)*100)

all_Feature_imp <- ImportTable(file ="../11.Classifier/RF/FeatureSel_Important/2021-09-30-Importance_Table_bac-v1.0.0.csv")
all_Feature_imp1 <- all_Feature_imp[rownames(subset_imp2), "Rank", drop = F]
all_Feature_imp2 <- cbind(all_Feature_imp1, combine_EnrDep)
all_Feature_imp2 <- all_Feature_imp2[order(all_Feature_imp2$Rank,all_Feature_imp2$EnrDep), ]


## plot enriched sub
enrich_mt <- subset_imp2[rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched'],]
enrich_col <- colorRamp2(c(0, 30), c('#bc0025', '#ffffcc'))

enrich_all_list <- all_Feature_imp2$Rank[all_Feature_imp2$EnrDep == 'Enriched']
names(enrich_all_list) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched']

enrich_all_list_label <- as.character(enrich_all_list)
names(enrich_all_list_label) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched']

anno_col_enrich <- enrich_col(enrich_all_list)
names(anno_col_enrich) <- enrich_all_list
enrich_label_col <- lapply(enrich_all_list, function(x){ifelse(x<=3, "white", "black")}) %>% unlist

enrich_rowname_col <- lapply(all_Feature_imp2$Kingdoms[all_Feature_imp2$EnrDep == 'Enriched'],
                             function(x){ifelse(x == 'fungi', "#ff410d", "#014586")})%>% unlist

enrich_row_anno <- rowAnnotation(all = anno_simple(enrich_all_list, 
                                                   pch = enrich_all_list_label,
                                                   col = anno_col_enrich,
                                                   gp = gpar(col = "black", lwd = 2),
                                                   simple_anno_size = unit(1, "cm"),
                                                   pt_gp = gpar(fontsize = 12, fontface = 'bold',
                                                                col = enrich_label_col),
                                                   pt_size = unit(5, "mm")), annotation_label = c(""))

colnames(enrich_mt) <- c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", "2019_ThomasA", "2019_WirbelJ", "2019_YachidaS", "2017_JunY", "2021_JunY")
h1 <- Heatmap(enrich_mt, col = enrich_col, name = "CRC-enriched\nSpecies",
              cluster_columns = F, cluster_rows = F, border = T, 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(enrich_mt[i, j], x, y, gp = gpar(fontsize = 12, fontface = 'bold', 
                                    col = ifelse(enrich_mt[i, j]<=3,'white', 'black')))},
              width = ncol(enrich_mt)*unit(10, "mm"), 
              height = nrow(enrich_mt)*unit(10, "mm"),
              right_annotation = enrich_row_anno,
              row_names_gp = gpar(col = enrich_rowname_col, fontface = 4))

h1




## plot depleted sub

deplet_mt <- subset_imp2[rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted'],]
deplet_col <- colorRamp2(c(0, 30), c('#243493', '#ffffcc'))

deplet_all_list <- all_Feature_imp2$Rank[all_Feature_imp2$EnrDep == 'Depleted']
names(deplet_all_list) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted']

deplet_all_list_label <- as.character(deplet_all_list)
names(deplet_all_list_label) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted']

anno_col_deplet <- deplet_col(deplet_all_list)
names(anno_col_deplet) <- deplet_all_list
deplet_label_col <- lapply(deplet_all_list, function(x){ifelse(x<=3, "white", "black")}) %>% unlist

deplet_rowname_col <- lapply(all_Feature_imp2$Kingdoms[all_Feature_imp2$EnrDep == 'Depleted'],
                             function(x){ifelse(x == 'fungi', "#ff410d", "#014586")})%>% unlist

deplet_row_anno <- rowAnnotation(all = anno_simple(deplet_all_list, 
                                                   pch = deplet_all_list_label,
                                                   col = anno_col_deplet,
                                                   gp = gpar(col = "black", lwd = 2),
                                                   simple_anno_size = unit(1, "cm"),
                                                   pt_gp = gpar(fontsize = 12, fontface = 'bold',
                                                                col = deplet_label_col),
                                                   pt_size = unit(5, "mm")))

colnames(deplet_mt) <- c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", "2019_ThomasA", "2019_WirbelJ", "2019_YachidaS", "2017_JunY", "2021_JunY")
h2 <- Heatmap(deplet_mt, col = deplet_col, name = "CRC-depleted\nSpecies",
              cluster_columns = F, cluster_rows = F, border = T, 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(deplet_mt[i, j], x, y, 
                          gp = gpar(fontsize = 12, fontface = 'bold', 
                                    col = ifelse(deplet_mt[i, j]<=3,'white', 'black')))},
              width = ncol(deplet_mt)*unit(10, "mm"), 
              height = nrow(deplet_mt)*unit(10, "mm"), 
              right_annotation = deplet_row_anno,
              row_names_gp = gpar(col = deplet_rowname_col, fontface = 4))
h2

ht_list <- h1 %v% h2 

ht_pdf <- FileCreate(DirPath = '../11.Classifier/RF/Plot', Prefix = 'supplementary-Bac-heatmap_RF_AUC', Suffix = 'pdf')

pdf(file = ht_pdf, width = 15, height = 15)
draw(ht_list)
dev.off()

```

### Fungi Heatmap

```{r, message = F, warning = F}
Importance_df2 <- ImportTable(file = '../11.Classifier/RF/FeatureSel_Important/fungi/2021-09-29-Combine_importantScore_rank_all-v1.0.0.csv')
subset_imp <- Importance_df2[Importance_df2$Rank3 == 'Sel', ]
subset_imp2 <- subset_imp[, grep('_Rank', colnames(subset_imp))]


fungi_median <- ImportTable(file = '../07.FeatureSelection/01.SSTF/2021-07-30-walsh-median-all-v1.0.0.csv')
rownames(fungi_median) <- gsub("\\W+", '_', rownames(fungi_median))
rownames(fungi_median) <- gsub("^_", '', rownames(fungi_median))
fungi_median$EnrDep <- ifelse(fungi_median$`all-Median` >0, 'Enriched', 'Depleted')
fungi_median$Kingdoms <- 'fungi'

subset_imp2 <- subset_imp2[(rownames(subset_imp2) %in% rownames(fungi_median)), ]


combine_EnrDep <- fungi_median[, c(5, 6), drop = F]
combine_EnrDep <- combine_EnrDep[rownames(subset_imp2), , drop = F]

comb_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-all-v1.0.0.csv'))
bac_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-bac-v1.0.0.csv'))

summary_AUC <- rbind(Comb_AUC = comb_AUC[which.max(comb_AUC$Mean),], Bac_AUC = bac_AUC[which.max(bac_AUC$Mean),])
summary_AUC <- rbind(summary_AUC, Ratio = (summary_AUC[1,]/summary_AUC[2,]-1)*100)

all_Feature_imp <- ImportTable(file ="../11.Classifier/RF/FeatureSel_Important/2021-09-30-Importance_Table_fungi-v1.0.0.csv")
all_Feature_imp1 <- all_Feature_imp[rownames(subset_imp2), "Rank", drop = F]
all_Feature_imp2 <- cbind(all_Feature_imp1, combine_EnrDep)
all_Feature_imp2 <- all_Feature_imp2[order(all_Feature_imp2$Rank,all_Feature_imp2$EnrDep), ]


## plot enriched sub
enrich_mt <- subset_imp2[rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched'],]
enrich_col <- colorRamp2(c(0, 30), c('#bc0025', '#ffffcc'))

enrich_all_list <- all_Feature_imp2$Rank[all_Feature_imp2$EnrDep == 'Enriched']
names(enrich_all_list) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched']

enrich_all_list_label <- as.character(enrich_all_list)
names(enrich_all_list_label) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched']

anno_col_enrich <- enrich_col(enrich_all_list)
names(anno_col_enrich) <- enrich_all_list
enrich_label_col <- lapply(enrich_all_list, function(x){ifelse(x<=3, "white", "black")}) %>% unlist

enrich_rowname_col <- lapply(all_Feature_imp2$Kingdoms[all_Feature_imp2$EnrDep == 'Enriched'],
                             function(x){ifelse(x == 'fungi', "#ff410d", "#014586")})%>% unlist

enrich_row_anno <- rowAnnotation(all = anno_simple(enrich_all_list, 
                                                   pch = enrich_all_list_label,
                                                   col = anno_col_enrich,
                                                   gp = gpar(col = "black", lwd = 2),
                                                   simple_anno_size = unit(1, "cm"),
                                                   pt_gp = gpar(fontsize = 12, fontface = 'bold',
                                                                col = enrich_label_col),
                                                   pt_size = unit(5, "mm")), annotation_label = c(""))

colnames(enrich_mt) <- c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", "2019_ThomasA", "2019_WirbelJ", "2019_YachidaS", "2017_JunY", "2021_JunY")
h1 <- Heatmap(enrich_mt, col = enrich_col, name = "CRC-enriched\nSpecies",
              cluster_columns = F, cluster_rows = F, border = T, 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(enrich_mt[i, j], x, y, gp = gpar(fontsize = 12, fontface = 'bold', 
                                    col = ifelse(enrich_mt[i, j]<=3,'white', 'black')))},
              width = ncol(enrich_mt)*unit(10, "mm"), 
              height = nrow(enrich_mt)*unit(10, "mm"),
              right_annotation = enrich_row_anno,
              row_names_gp = gpar(col = enrich_rowname_col, fontface = 4))

h1




## plot depleted sub

deplet_mt <- subset_imp2[rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted'],]
deplet_col <- colorRamp2(c(0, 30), c('#243493', '#ffffcc'))

deplet_all_list <- all_Feature_imp2$Rank[all_Feature_imp2$EnrDep == 'Depleted']
names(deplet_all_list) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted']

deplet_all_list_label <- as.character(deplet_all_list)
names(deplet_all_list_label) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted']

anno_col_deplet <- deplet_col(deplet_all_list)
names(anno_col_deplet) <- deplet_all_list
deplet_label_col <- lapply(deplet_all_list, function(x){ifelse(x<=3, "white", "black")}) %>% unlist

deplet_rowname_col <- lapply(all_Feature_imp2$Kingdoms[all_Feature_imp2$EnrDep == 'Depleted'],
                             function(x){ifelse(x == 'fungi', "#ff410d", "#014586")})%>% unlist

deplet_row_anno <- rowAnnotation(all = anno_simple(deplet_all_list, 
                                                   pch = deplet_all_list_label,
                                                   col = anno_col_deplet,
                                                   gp = gpar(col = "black", lwd = 2),
                                                   simple_anno_size = unit(1, "cm"),
                                                   pt_gp = gpar(fontsize = 12, fontface = 'bold',
                                                                col = deplet_label_col),
                                                   pt_size = unit(5, "mm")))

colnames(deplet_mt) <- c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", "2019_ThomasA", "2019_WirbelJ", "2019_YachidaS", "2017_JunY", "2021_JunY")
h2 <- Heatmap(deplet_mt, col = deplet_col, name = "CRC-depleted\nSpecies",
              cluster_columns = F, cluster_rows = F, border = T, 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(deplet_mt[i, j], x, y, 
                          gp = gpar(fontsize = 12, fontface = 'bold', 
                                    col = ifelse(deplet_mt[i, j]<=3,'white', 'black')))},
              width = ncol(deplet_mt)*unit(10, "mm"), 
              height = nrow(deplet_mt)*unit(10, "mm"), 
              right_annotation = deplet_row_anno,
              row_names_gp = gpar(col = deplet_rowname_col, fontface = 4))
h2

ht_list <- h1 %v% h2 

ht_pdf <- FileCreate(DirPath = '../11.Classifier/RF/Plot', Prefix = 'supplementary-Fungi-heatmap_RF_AUC', Suffix = 'pdf')

pdf(file = ht_pdf, width = 15, height = 15)
draw(ht_list)
dev.off()

```








## Plot single feature AUC

```{r, message = F, warning = F}

mydata <- cn_combin_core
cohort_lst <- unique(cn_meta_df$Cohort)

auc_sub_list <- list()
for (sub_cohort in cohort_lst) {
  roc_feature_df <- matrix(data = NA,
                           nrow = 100,
                           ncol = ncol(mydata)-1) %>% 
    as.data.frame()
  rownames(roc_feature_df) <- paste0('Seed_', 1:100)
  colnames(roc_feature_df) <- colnames(mydata)[2:ncol(mydata)]
  for (sel_seed in 1:100) {
    for (single_feature in colnames(mydata)[2:ncol(mydata)]) {
      train_set <- mydata[rownames(cn_meta_df)[cn_meta_df$Cohort != sub_cohort], ]
      valda_set <- mydata[rownames(cn_meta_df)[cn_meta_df$Cohort == sub_cohort], ]
      set.seed(sel_seed)
      rf <- randomForest(as.formula(paste0("Stage ~ ", single_feature)), data = train_set)
      pred <- predict(rf, newdata=valda_set[, single_feature, drop = F], type = 'prob')[,2]
      pred_result <- prediction(pred, valda_set[,1])
      auc_value <- round(performance(pred_result, measure = "auc")@y.values[[1]]*100, 2)
      
      roc_feature_df[paste0('Seed_', sel_seed), single_feature] <- auc_value
      
    }
  }
  roc_csv <- FileCreate(DirPath = "../11.Classifier/RF/single_AUC", Suffix = "csv",
                        Prefix = paste0("SingleFeature_AUC_Performance-", sub_cohort))
  write.csv(x = roc_feature_df, file = roc_csv)
  auc_sub_list[[sub_cohort]] <- roc_feature_df
}



best_seed_df <- matrix(NA, nrow = 1, ncol = 4)
best_seed_df <- best_seed_df[-1, ]
for (sub_cohort in cohort_lst) {
  tmp_df <- auc_sub_list[[sub_cohort]]
  for (single_feature in colnames(mydata)[2:ncol(mydata)]) {
    best_seed_df <- rbind(best_seed_df, 
                          c(sub_cohort, single_feature,
                            which.max(tmp_df[, single_feature]),
                            max(tmp_df[, single_feature])))
  }
}

best_seed_df <- as.data.frame(best_seed_df)
colnames(best_seed_df) <- c("Cohort", "Feature", "Seed", "AUC")
best_seed_df$MeanAUC <- apply(best_seed_df, 1, function(x){
  round(mean(as.numeric(best_seed_df$AUC[best_seed_df$Feature == x[2]])),2)
})

best_seed_csv <- FileCreate(DirPath = "../11.Classifier/RF/single_AUC",
                            Prefix = "Summary-Best_Seed_AUC", Suffix = "csv")

write.csv(x = best_seed_df, file = best_seed_csv)

# Mean AUC > 60%
####     here
tmp_df2 <- best_seed_df[best_seed_df$Cohort == cohort_lst[8],]

head(tmp_df2[order(tmp_df2$AUC, decreasing = T),], 10)



#ff420e   fungi
#004586   bac
#ffd320   combine


```









