---
title: "RandomForest"
author: "ifanlyn@outlook.com"
date: "2021/9/28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages and subroutines {.tabset}

### Package

```{r packages, message = F, warning = F}
if (!require(DT)) install.packages('DT'); require(DT) # for databale
if (!require(randomForest)) install.packages("randomForest"); require(randomForest) # randomForest
if (!require(pROC)) install.packages("pROC"); require(pROC)
if (!require(ROCR)) install.packages("ROCR"); require(ROCR)
if (!require(mlbench)) install.packages("mlbench"); require(mlbench)
if (!require(caret)) install.packages('caret', dependencies = TRUE); require(caret)
if (!require(ComplexHeatmap)) install.packages("ComplexHeatmap"); require(ComplexHeatmap)
if (!require(circlize)) install.packages("circlize"); require(circlize)
if (!require(reshape2)) install.packages("reshape2"); require(reshape2) # dcast
require(ochRe) # for scale_fill_ochre
require(ggrepel) # geom_text_repel
# random forest 
require(dplyr)
require(randomForest)
require(eegkit)
require(rlist)
require(edfReader)
require(eegUtils)
require(e1071)
require(ggplot2)
```


### Import table

```{r ImportTable , message = F, warning = F}
ImportTable <- function(file, header = T, row.names = 1, sep = ',', check.names = FALSE, ...){
  data <- read.csv(file = file, header = header, row.names = row.names, sep = sep, check.names = check.names, ...)
  return(data)
}
```

### Show_table

```{r show_table, message = F, warning = F}
Show_table <- function(df, rownames = T, filter="top", options = list(pageLength = 10, scrollX=T), ...){
  if (ncol(df) > 100) {
    df <- df[, 1:100]
    message('Due to the column dim is > 100, it only shows the top 100 columns')
  }
  datatable(df, rownames = rownames, 
            filter=filter, options = options, ... )

}
```

### FileCreate

```{r , message = F, warning = F}
FileCreate <- function(DirPath = "./",Prefix = "ExampleTest", Suffix = "pdf", version = "0.0"){
  date=as.character(Sys.Date())
  DirPath = gsub("/$","",DirPath)
  Suffix = gsub('^[.]',"",Suffix)
  if(! dir.exists(DirPath)){
    dir.create(DirPath,recursive =TRUE)
  }
  if (version == "0.0") {
    version=10
    while(TRUE){
      version_=paste(unlist(strsplit(as.character(version),"")),collapse=".")
      DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
      if(! file.exists(DirPathName)){
        return(DirPathName)
        break
      }
      version = version + 1
    }
  }else{
    DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
    if(! dir.exists(DirPathName)){
      return(DirPathName)
    }
    return(DirPathName)
  }
}
  
```

## Import data {.tabset}

### metadata

```{r , message = F, warning = F}
meta_df <- ImportTable(file = '../00.RawData/metaData/2021-07-26-metaInfo-subgroup_1329-v6.0.csv')

```

### Import Eukaryota matrix

```{r , message = F, warning = F}

euk_otu <- ImportTable(file = '../07.FeatureSelection/02.RelMedianAbundance/2021-07-30-normalized-0.1%-RelMedianAbundance-matrix-v1.0.0.csv')
euk_sel <- ImportTable(file = '../07.FeatureSelection/01.SSTF/2021-07-30-wilcoxon-all-v1.0.0.csv')
euk_sel <- euk_sel[order(euk_sel$adj_pvalue),]
euk_core <- euk_otu[, rownames(euk_sel)[euk_sel$adj_pvalue < 0.01]]
euk_sel_list <- colnames(euk_core)


```

### Import Bacteria matrix

```{r , message = F, warning = F}

bac_otu <- ImportTable(file = '../09.Bacteria/02.RelMedianAbundance/2021-08-05-normalized-0.1%-RelMedianAbundance-matrix-v1.0.0.csv')
bac_sel <- ImportTable(file = '../09.Bacteria/2021-08-06-wilcoxon-all-v1.0.0.csv')
bac_sel <- bac_sel[order(bac_sel$adj_pvalue),]
bac_core_all <- bac_otu[, rownames(bac_sel)[bac_sel$adj_pvalue < 0.01 & bac_sel$AbsLog2FC > 0.5]]
bac_sel_list_all <- colnames(bac_core_all)
bac_sel_list <- bac_sel_list_all[!grepl('sp. ', bac_sel_list_all)]
bac_core <- bac_core_all[, bac_sel_list]
```

### Combine the core sets and stages

```{r , message = F, warning = F}
comb_df <- as.data.frame(cbind(Stage = meta_df$Stage, cbind(euk_otu, bac_otu)))
colnames(comb_df) <- gsub("\\W+", '_', colnames(comb_df))
colnames(comb_df) <- gsub("^_", '', colnames(comb_df))
cn_combin_df <- comb_df[comb_df$Stage!= 'adenoma', ]
cn_combin_df$Stage <- factor(x = cn_combin_df$Stage, levels = c('CRC', 'CTRL'))
cn_meta_df <- meta_df[meta_df$Stage != 'adenoma', ]

comb_core <- as.data.frame(cbind(Stage = meta_df$Stage, cbind(euk_core, bac_core)))
colnames(comb_core) <- gsub("\\W+", '_', colnames(comb_core))
colnames(comb_core) <- gsub("^_", '', colnames(comb_core))
cn_combin_core <- comb_core[comb_core$Stage!= 'adenoma', ]
cn_combin_core$Stage <- factor(x = cn_combin_core$Stage, levels = c('CRC', 'CTRL'))


comb_euk <- as.data.frame(cbind(Stage = meta_df$Stage, euk_otu))
colnames(comb_euk) <- gsub("\\W+", '_', colnames(comb_euk))
colnames(comb_euk) <- gsub("^_", '', colnames(comb_euk))
cn_combin_euk <- comb_euk[comb_euk$Stage!= 'adenoma', ]
cn_combin_euk$Stage <- factor(x = cn_combin_euk$Stage, levels = c('CRC', 'CTRL'))


comb_bac <- as.data.frame(cbind(Stage = meta_df$Stage, bac_otu))
colnames(comb_bac) <- gsub("\\W+", '_', colnames(comb_bac))
colnames(comb_bac) <- gsub("^_", '', colnames(comb_bac))
cn_combin_bac <- comb_bac[comb_bac$Stage!= 'adenoma', ]
cn_combin_bac$Stage <- factor(x = cn_combin_bac$Stage, levels = c('CRC', 'CTRL'))

```

## Single Fungi Classifiers

### Calculate the model (best seed)
* it will cost a lot of time
```{r , message = F, warning = F}
mydata <- cn_combin_euk
cohort_lst <- unique(cn_meta_df$Cohort)

auc_sub_list <- list()
calculate_or_not <- F
# calculate_or_not <- T
seed_num <- 100
if (calculate_or_not) {
  My_function <- function(seed, train_set, valda_set){
    set.seed(seed)
    auc_list <- NULL
    for(single_feature in colnames(mydata)[2:ncol(mydata)]){
        rf <- randomForest::randomForest(as.formula(paste0("Stage ~ ", single_feature)), data = train_set)
        pred <- predict(rf, newdata=valda_set[, single_feature, drop = F], type = 'prob')[,2]
        pred_result <- ROCR::prediction(pred, valda_set[,1])
        auc_list <- c(auc_list, round(ROCR::performance(pred_result, measure = "auc")@y.values[[1]]*100, 2))
    }
    return(matrix(data = auc_list, nrow = 1, dimnames = list(c(paste0('Seed_', sel_seed)), 
                                                             colnames(mydata)[2:ncol(mydata)])))
  }
  for (sub_cohort in cohort_lst) {    
    train_set <- mydata[rownames(cn_meta_df)[cn_meta_df$Cohort != sub_cohort], ]
    valda_set <- mydata[rownames(cn_meta_df)[cn_meta_df$Cohort == sub_cohort], ]
    cores=detectCores()
    cl <- makeCluster(cores[1]-1) #not to overload your computer
    registerDoParallel(cl)
    # parallel
    auc_res <- foreach(sel_seed = 1:seed_num, .combine=rbind) %dopar%{
        tmp_res <- My_function(seed = sel_seed, train_set = train_set, valda_set = valda_set)
    }
    stopCluster(cl) #stop cluster   
    roc_csv <- FileCreate(DirPath = "../11.Classifier/RF/single_AUC", Suffix = "csv",
                          Prefix = paste0("SingleFeature_AUC_Performance-", sub_cohort))
    write.csv(x = auc_res, file = roc_csv)
    auc_sub_list[[sub_cohort]] <- auc_res
  }
}else{
  for (sub_cohort in cohort_lst) {
    auc_sub_list[[sub_cohort]] <- ImportTable(file = paste0( '../11.Classifier/RF/single_AUC/2022-01-18-SingleFeature_AUC_Performance-',sub_cohort,'-v1.0.csv'))
  }
}



```

### select the best seed for each cohort
```{r , message = F, warning = F}
best_seed_df <- matrix(NA, nrow = 1, ncol = 4)
best_seed_df <- best_seed_df[-1, ]
for (sub_cohort in cohort_lst) {
  tmp_df <- auc_sub_list[[sub_cohort]]
  for (single_feature in colnames(mydata)[2:ncol(mydata)]) {
    best_seed_df <- rbind(best_seed_df, 
                          c(sub_cohort, single_feature,
                            which.max(tmp_df[, single_feature]),
                            max(tmp_df[, single_feature])))
  }
}

best_seed_df <- as.data.frame(best_seed_df)
colnames(best_seed_df) <- c("Cohort", "Feature", "Seed", "AUC")
best_seed_df$MeanAUC <- apply(best_seed_df, 1, function(x){
  round(mean(as.numeric(best_seed_df$AUC[best_seed_df$Feature == x[2]])),2)
})

best_seed_df$Kingdoms <- rep('fungi', nrow(best_seed_df))

# save 
best_seed_csv <- FileCreate(DirPath = "../11.Classifier/RF/single_AUC",
                            Prefix = "Summary-Best_Seed_AUC", Suffix = "csv")

write.csv(x = best_seed_df, file = best_seed_csv)

```

### calculate the best model

```{r , message = F, warning = F}
all_euk_list <- colnames(cn_combin_euk)[2: ncol(cn_combin_euk)]
prj_list <- list()
auc_df <- matrix(data = NA, nrow = 0, ncol = 3)
for (c in cohort_lst) {
  
  for (sel_single_feature in all_euk_list) {
    sel_imp_df <- cn_combin_euk[, c('Stage', sel_single_feature)]
    train_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort != c], ]
    valda_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort == c], ]
    set.seed(best_seed_df$Seed[best_seed_df$Cohort == c & best_seed_df$Feature == sel_single_feature])
    # set.seed(123)
    rf <- randomForest(Stage ~ ., data = train_set)
    pred <- predict(rf, newdata=valda_set[-1], type = 'prob')[,2]
    pred_result <- prediction(pred, valda_set[,1])
    auc_value <- round(performance(pred_result, measure = "auc")@y.values[[1]]*100, 2)
    prj_list[[c]][[sel_single_feature]][['rf']] <- rf
    prj_list[[c]][[sel_single_feature]][['roc']] <- roc(valda_set[,1], pred, legacy.axes = T, percent = T)
    auc_df <- rbind(auc_df, c(cohort = c, fungi = sel_single_feature, auc = auc_value))
  }
}
auc_df <- as.data.frame(auc_df)
auc_df$auc <- as.numeric(auc_df$auc)
```

### Plot the mean of AUC

```{r , message = F, warning = F}
# mean of AUC
# mean
mean_auc_df <- aggregate(auc ~ fungi, auc_df, mean)
mean_auc_df <- mean_auc_df[order(mean_auc_df$auc, decreasing = T), ]
rownames(mean_auc_df) <- mean_auc_df$fungi
comb_auc_df <- cbind(auc_df, mean_auc = mean_auc_df[auc_df$fungi, 'auc'])
mean_auc_df2 <- mean_auc_df
mean_auc_df2$max <- unlist(lapply(mean_auc_df2$fungi, function(x){
  max(auc_df[auc_df$fungi == x, ]$auc)
}))
  
my_color <- c(ochre_palettes$namatjira_qual)
names(my_color) <-  c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", 
                      "2019_ThomasAM", "2019_WirbelJ", "2019_YachidaS", 
                      "2017_YuJ", "2021_YuJ")

comb_auc_df$cohort <- ifelse(comb_auc_df$cohort == "2014_ZellerG", "2014_ZellerG", 
                        ifelse(comb_auc_df$cohort == "2015_FengQ", "2015_FengQ", 
                               ifelse(comb_auc_df$cohort == "2016_VogtmannE", "2016_VogtmannE", 
                                      ifelse(comb_auc_df$cohort == "2019_Thom", "2019_ThomasAM", 
                                             ifelse(comb_auc_df$cohort == "2019_WirbelJ", "2019_WirbelJ", 
                                                    ifelse(comb_auc_df$cohort == "2019_Yachida", "2019_YachidaS", 
                                                           ifelse(comb_auc_df$cohort == "2021_JunY_1", "2017_YuJ","2021_YuJ")))))))

euk_sel_mod_list <- gsub("\\W+", '_', euk_sel_list)
euk_sel_mod_list <- gsub("^_", '', euk_sel_mod_list)

mean_auc_df3 <- mean_auc_df2[mean_auc_df2$fungi %in% euk_sel_mod_list,]


comb_auc_df$label <- ifelse(comb_auc_df$fungi %in% euk_sel_mod_list, 'core', 'other')
g_point <- ggplot() + 
  geom_point(data = comb_auc_df, 
             shape = 20, 
             aes(x = auc, y = mean_auc,
                 size = label, col = cohort, 
                 alpha = label)) +
  scale_color_manual(values = my_color)+
  scale_alpha_manual(values = c(core = 1, other = 0.2)) +
  scale_size_manual(values = c(core = 3, other = 2)) +
  geom_text_repel(data = mean_auc_df3[mean_auc_df3$auc > 55,], aes(y = auc, x= max,label = fungi),
                  hjust = -0.1, direction = 'x') +
  theme_bw()+
  # geom_hline(data = mean_auc_df3, aes(yintercept = auc), linetype = 2)
  geom_line(data = comb_auc_df[comb_auc_df$fungi %in% euk_sel_mod_list & comb_auc_df$mean_auc > 55, ],
            aes(x = auc, y = mean_auc, group = fungi), linetype = 2)
#ff420e   fungi
#004586   bac
#ffd320   combine

scatter_pdf <- FileCreate(DirPath = '../11.Classifier/RF/single_scatter', Prefix = 'AUC_scatter_plot_mean', Suffix = 'pdf')

pdf(file = scatter_pdf, width = 7, height = 7)
plot(g_point + theme(legend.position = "none"))
plot(g_point)
dev.off()
```

### Plot the median of AUC

```{r , message = F, warning = F}

# median AUC

# median
median_auc_df <- aggregate(auc ~ fungi, auc_df, median)
median_auc_df <- median_auc_df[order(median_auc_df$auc, decreasing = T), ]
rownames(median_auc_df) <- median_auc_df$fungi
comb_auc_df_ <- cbind(auc_df, median_auc = median_auc_df[auc_df$fungi, 'auc'])
median_auc_df2 <- median_auc_df
median_auc_df2$max <- unlist(lapply(median_auc_df2$fungi, function(x){
  max(auc_df[auc_df$fungi == x, ]$auc)
}))
  
my_color <- c(ochre_palettes$namatjira_qual)
names(my_color) <-  c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", 
                      "2019_ThomasAM", "2019_WirbelJ", "2019_YachidaS", 
                      "2017_YuJ", "2021_YuJ")

comb_auc_df_$cohort <- ifelse(comb_auc_df_$cohort == "2014_ZellerG", "2014_ZellerG", 
                        ifelse(comb_auc_df_$cohort == "2015_FengQ", "2015_FengQ", 
                               ifelse(comb_auc_df_$cohort == "2016_VogtmannE", "2016_VogtmannE", 
                                      ifelse(comb_auc_df_$cohort == "2019_Thom", "2019_ThomasAM", 
                                             ifelse(comb_auc_df_$cohort == "2019_WirbelJ", "2019_WirbelJ", 
                                                    ifelse(comb_auc_df_$cohort == "2019_Yachida", "2019_YachidaS", 
                                                           ifelse(comb_auc_df_$cohort == "2021_JunY_1", "2017_YuJ","2021_YuJ")))))))

euk_sel_mod_list <- gsub("\\W+", '_', euk_sel_list)
euk_sel_mod_list <- gsub("^_", '', euk_sel_mod_list)

median_auc_df3 <- median_auc_df2[median_auc_df2$fungi %in% euk_sel_mod_list,]


comb_auc_df_$label <- ifelse(comb_auc_df_$fungi %in% euk_sel_mod_list, 'core', 'other')
g_point_median <- ggplot() + 
  geom_point(data = comb_auc_df_, 
             shape = 20, 
             aes(x = auc, y = median_auc,
                 size = label, col = cohort, 
                 alpha = label)) +
  scale_color_manual(values = my_color)+
  scale_alpha_manual(values = c(core = 1, other = 0.2)) +
  scale_size_manual(values = c(core = 3, other = 2)) +
  geom_text(data = median_auc_df3[median_auc_df3$auc > 55,], aes(y = auc, x= max,label = fungi),
                  hjust = -0.1) +
  theme_bw()+
  # geom_hline(data = median_auc_df3, aes(yintercept = auc), linetype = 2)
  geom_line(data = comb_auc_df_[comb_auc_df_$fungi %in% euk_sel_mod_list & comb_auc_df_$median_auc > 55, ],
            aes(x = auc, y = median_auc, group = fungi), linetype = 2)
#ff420e   fungi
#004586   bac
#ffd320   combine

scatter_pdf <- FileCreate(DirPath = '../11.Classifier/RF/single_scatter', Prefix = 'AUC_scatter_plot_median', Suffix = 'pdf')

pdf(file = scatter_pdf, width = 7, height = 7)
plot(g_point_median + theme(legend.position = "none"))
plot(g_point_median)
dev.off()

```

### Draw the selected median AUC in ROC Plot 

```{r, message =F, warning = F}
my_color <- c(ochre_palettes$namatjira_qual)
names(my_color) <-  c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", 
                      "2019_Thom", "2019_WirbelJ", "2019_Yachida", 
                      "2021_JunY_1", "2021_JunY_2")
median_sel_fungi <- unique(comb_auc_df_$fungi[comb_auc_df_$fungi %in% euk_sel_mod_list & comb_auc_df_$median_auc > 55])

auc_df <- matrix(data = NA, nrow = 0, ncol = 3)
for (sel_single_feature in median_sel_fungi) {
    roc_list <- list()
    for (c in cohort_lst) {  
        sel_imp_df <- cn_combin_euk[, c('Stage', sel_single_feature)]
        train_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort != c], ]
        valda_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort == c], ]
        set.seed(best_seed_df$Seed[best_seed_df$Cohort == c & best_seed_df$Feature == sel_single_feature])
        # set.seed(123)
        rf <- randomForest(Stage ~ ., data = train_set)
        pred <- predict(rf, newdata=valda_set[-1], type = 'prob')[,2]
        pred_result <- prediction(pred, valda_set[,1])
        auc_value <- round(performance(pred_result, measure = "auc")@y.values[[1]]*100, 2)
        roc_list[[c]] <- roc(valda_set[,1], pred, legacy.axes = T, percent = T)
    }
    roc_pdf <- FileCreate(DirPath = "../11.Classifier/RF/single_Comparison/", 
                          Prefix = paste0('ROC-', sel_single_feature), 
                          Suffix = 'pdf')
    pdf(file = roc_pdf)
    par(pty = "s")
    plot(roc_list[[cohort_lst[1]]], col = my_color[cohort_lst[1]], lwd = 2, legacy.axes = T, 
         xlab = "False Postive Percentage", ylab = "True Postive Percentage")
    for(c in cohort_lst[2:length(cohort_lst)]){
        lines(roc_list[[c]], col = my_color[c], lwd = 2)
    }
    dev.off()   
}

```



## Single Bacteria Classifiers 

### Calculate the model (best seed)
* it will cost a lot of time
```{r , message = F, warning = F}
mydata <- cn_combin_bac
cohort_lst <- unique(cn_meta_df$Cohort)

auc_sub_list <- list()
calculate_or_not <- F
# calculate_or_not <- T
seed_num <- 100
if (calculate_or_not) {
  My_function <- function(seed, train_set, valda_set){
    set.seed(seed)
    auc_list <- NULL
    for(single_feature in colnames(mydata)[2:ncol(mydata)]){
        rf <- randomForest::randomForest(as.formula(paste0("Stage ~ ", single_feature)), data = train_set)
        pred <- predict(rf, newdata=valda_set[, single_feature, drop = F], type = 'prob')[,2]
        pred_result <- ROCR::prediction(pred, valda_set[,1])
        auc_list <- c(auc_list, round(ROCR::performance(pred_result, measure = "auc")@y.values[[1]]*100, 2))
    }
    return(matrix(data = auc_list, nrow = 1, dimnames = list(c(paste0('Seed_', sel_seed)), 
                                                             colnames(mydata)[2:ncol(mydata)])))
  }
  for (sub_cohort in cohort_lst) {    
    train_set <- mydata[rownames(cn_meta_df)[cn_meta_df$Cohort != sub_cohort], ]
    valda_set <- mydata[rownames(cn_meta_df)[cn_meta_df$Cohort == sub_cohort], ]
    cores=detectCores()
    cl <- makeCluster(cores[1]-1) #not to overload your computer
    registerDoParallel(cl)
    # parallel
    auc_res <- foreach(sel_seed = 1:seed_num, .combine=rbind) %dopar%{
        tmp_res <- My_function(seed = sel_seed, train_set = train_set, valda_set = valda_set)
    }
    stopCluster(cl) #stop cluster   
    roc_csv <- FileCreate(DirPath = "../11.Classifier/RF/single_AUC_bac", Suffix = "csv",
                          Prefix = paste0("SingleFeature_AUC_Performance-", sub_cohort))
    write.csv(x = auc_res, file = roc_csv)
    auc_sub_list[[sub_cohort]] <- auc_res
  }
}else{
  for (sub_cohort in cohort_lst) {
    auc_sub_list[[sub_cohort]] <- ImportTable(file = paste0( '../11.Classifier/RF/single_AUC_bac/2022-01-19-SingleFeature_AUC_Performance-',sub_cohort,'-v1.0.csv'))
  }
}



```

### select the best seed for each cohort
```{r , message = F, warning = F}
best_seed_df <- matrix(NA, nrow = 1, ncol = 4)
best_seed_df <- best_seed_df[-1, ]
for (sub_cohort in cohort_lst) {
  tmp_df <- auc_sub_list[[sub_cohort]]
  for (single_feature in colnames(mydata)[2:ncol(mydata)]) {
    best_seed_df <- rbind(best_seed_df, 
                          c(sub_cohort, single_feature,
                            which.max(tmp_df[, single_feature]),
                            max(tmp_df[, single_feature])))
  }
}

best_seed_df <- as.data.frame(best_seed_df)
colnames(best_seed_df) <- c("Cohort", "Feature", "Seed", "AUC")
best_seed_df$MeanAUC <- apply(best_seed_df, 1, function(x){
  round(mean(as.numeric(best_seed_df$AUC[best_seed_df$Feature == x[2]])),2)
})

best_seed_df$Kingdoms <- rep('fungi', nrow(best_seed_df))

# save 
best_seed_csv <- FileCreate(DirPath = "../11.Classifier/RF/single_AUC_bac",
                            Prefix = "Summary-Best_Seed_AUC", Suffix = "csv")

write.csv(x = best_seed_df, file = best_seed_csv)

```

### calculate the best model

```{r , message = F, warning = F}
all_bac_list <- colnames(cn_combin_bac)[2: ncol(cn_combin_bac)]
prj_list <- list()
auc_df <- matrix(data = NA, nrow = 0, ncol = 3)
for (c in cohort_lst) {
  
  for (sel_single_feature in all_bac_list) {
    sel_imp_df <- cn_combin_euk[, c('Stage', sel_single_feature)]
    train_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort != c], ]
    valda_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort == c], ]
    set.seed(best_seed_df$Seed[best_seed_df$Cohort == c & best_seed_df$Feature == sel_single_feature])
    # set.seed(123)
    rf <- randomForest(Stage ~ ., data = train_set)
    pred <- predict(rf, newdata=valda_set[-1], type = 'prob')[,2]
    pred_result <- prediction(pred, valda_set[,1])
    auc_value <- round(performance(pred_result, measure = "auc")@y.values[[1]]*100, 2)
    prj_list[[c]][[sel_single_feature]][['rf']] <- rf
    prj_list[[c]][[sel_single_feature]][['roc']] <- roc(valda_set[,1], pred, legacy.axes = T, percent = T)
    auc_df <- rbind(auc_df, c(cohort = c, bac = sel_single_feature, auc = auc_value))
  }
}
auc_df <- as.data.frame(auc_df)
auc_df$auc <- as.numeric(auc_df$auc)
```
<!--
### Plot the mean of AUC

```{r , message = F, warning = F}
# mean of AUC
# mean
mean_auc_df <- aggregate(auc ~ fungi, auc_df, mean)
mean_auc_df <- mean_auc_df[order(mean_auc_df$auc, decreasing = T), ]
rownames(mean_auc_df) <- mean_auc_df$fungi
comb_auc_df <- cbind(auc_df, mean_auc = mean_auc_df[auc_df$fungi, 'auc'])
mean_auc_df2 <- mean_auc_df
mean_auc_df2$max <- unlist(lapply(mean_auc_df2$fungi, function(x){
  max(auc_df[auc_df$fungi == x, ]$auc)
}))
  
my_color <- c(ochre_palettes$namatjira_qual)
names(my_color) <-  c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", 
                      "2019_ThomasAM", "2019_WirbelJ", "2019_YachidaS", 
                      "2017_YuJ", "2021_YuJ")

comb_auc_df$cohort <- ifelse(comb_auc_df$cohort == "2014_ZellerG", "2014_ZellerG", 
                        ifelse(comb_auc_df$cohort == "2015_FengQ", "2015_FengQ", 
                               ifelse(comb_auc_df$cohort == "2016_VogtmannE", "2016_VogtmannE", 
                                      ifelse(comb_auc_df$cohort == "2019_Thom", "2019_ThomasAM", 
                                             ifelse(comb_auc_df$cohort == "2019_WirbelJ", "2019_WirbelJ", 
                                                    ifelse(comb_auc_df$cohort == "2019_Yachida", "2019_YachidaS", 
                                                           ifelse(comb_auc_df$cohort == "2021_JunY_1", "2017_YuJ","2021_YuJ")))))))

euk_sel_mod_list <- gsub("\\W+", '_', euk_sel_list)
euk_sel_mod_list <- gsub("^_", '', euk_sel_mod_list)

mean_auc_df3 <- mean_auc_df2[mean_auc_df2$fungi %in% euk_sel_mod_list,]


comb_auc_df$label <- ifelse(comb_auc_df$fungi %in% euk_sel_mod_list, 'core', 'other')
g_point <- ggplot() + 
  geom_point(data = comb_auc_df, 
             shape = 20, 
             aes(x = auc, y = mean_auc,
                 size = label, col = cohort, 
                 alpha = label)) +
  scale_color_manual(values = my_color)+
  scale_alpha_manual(values = c(core = 1, other = 0.2)) +
  scale_size_manual(values = c(core = 3, other = 2)) +
  geom_text_repel(data = mean_auc_df3[mean_auc_df3$auc > 55,], aes(y = auc, x= max,label = fungi),
                  hjust = -0.1, direction = 'x') +
  theme_bw()+
  # geom_hline(data = mean_auc_df3, aes(yintercept = auc), linetype = 2)
  geom_line(data = comb_auc_df[comb_auc_df$fungi %in% euk_sel_mod_list & comb_auc_df$mean_auc > 55, ],
            aes(x = auc, y = mean_auc, group = fungi), linetype = 2)
#ff420e   fungi
#004586   bac
#ffd320   combine

scatter_pdf <- FileCreate(DirPath = '../11.Classifier/RF/single_scatter', Prefix = 'AUC_scatter_plot_mean', Suffix = 'pdf')

pdf(file = scatter_pdf, width = 7, height = 7)
plot(g_point + theme(legend.position = "none"))
plot(g_point)
dev.off()
```
-->
### Plot the median of AUC

```{r , message = F, warning = F}

# median AUC

# median
median_auc_df <- aggregate(auc ~ fungi, auc_df, median)
median_auc_df <- median_auc_df[order(median_auc_df$auc, decreasing = T), ]
rownames(median_auc_df) <- median_auc_df$fungi
comb_auc_df_ <- cbind(auc_df, median_auc = median_auc_df[auc_df$fungi, 'auc'])
median_auc_df2 <- median_auc_df
median_auc_df2$max <- unlist(lapply(median_auc_df2$fungi, function(x){
  max(auc_df[auc_df$fungi == x, ]$auc)
}))
  
my_color <- c(ochre_palettes$namatjira_qual)
names(my_color) <-  c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", 
                      "2019_ThomasAM", "2019_WirbelJ", "2019_YachidaS", 
                      "2017_YuJ", "2021_YuJ")

comb_auc_df_$cohort <- ifelse(comb_auc_df_$cohort == "2014_ZellerG", "2014_ZellerG", 
                        ifelse(comb_auc_df_$cohort == "2015_FengQ", "2015_FengQ", 
                               ifelse(comb_auc_df_$cohort == "2016_VogtmannE", "2016_VogtmannE", 
                                      ifelse(comb_auc_df_$cohort == "2019_Thom", "2019_ThomasAM", 
                                             ifelse(comb_auc_df_$cohort == "2019_WirbelJ", "2019_WirbelJ", 
                                                    ifelse(comb_auc_df_$cohort == "2019_Yachida", "2019_YachidaS", 
                                                           ifelse(comb_auc_df_$cohort == "2021_JunY_1", "2017_YuJ","2021_YuJ")))))))

euk_sel_mod_list <- gsub("\\W+", '_', euk_sel_list)
euk_sel_mod_list <- gsub("^_", '', euk_sel_mod_list)

median_auc_df3 <- median_auc_df2[median_auc_df2$fungi %in% euk_sel_mod_list,]


comb_auc_df_$label <- ifelse(comb_auc_df_$fungi %in% euk_sel_mod_list, 'core', 'other')
g_point_median <- ggplot() + 
  geom_point(data = comb_auc_df_, 
             shape = 20, 
             aes(x = auc, y = median_auc,
                 size = label, col = cohort, 
                 alpha = label)) +
  scale_color_manual(values = my_color)+
  scale_alpha_manual(values = c(core = 1, other = 0.2)) +
  scale_size_manual(values = c(core = 3, other = 2)) +
  geom_text(data = median_auc_df3[median_auc_df3$auc > 55,], aes(y = auc, x= max,label = fungi),
                  hjust = -0.1) +
  theme_bw()+
  # geom_hline(data = median_auc_df3, aes(yintercept = auc), linetype = 2)
  geom_line(data = comb_auc_df_[comb_auc_df_$fungi %in% euk_sel_mod_list & comb_auc_df_$median_auc > 55, ],
            aes(x = auc, y = median_auc, group = fungi), linetype = 2)
#ff420e   fungi
#004586   bac
#ffd320   combine

scatter_pdf <- FileCreate(DirPath = '../11.Classifier/RF/single_scatter', Prefix = 'AUC_scatter_plot_median', Suffix = 'pdf')

pdf(file = scatter_pdf, width = 7, height = 7)
plot(g_point_median + theme(legend.position = "none"))
plot(g_point_median)
dev.off()

```

### Draw the selected median AUC in ROC Plot 

```{r, message =F, warning = F}
my_color <- c(ochre_palettes$namatjira_qual)
names(my_color) <-  c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", 
                      "2019_Thom", "2019_WirbelJ", "2019_Yachida", 
                      "2021_JunY_1", "2021_JunY_2")
median_sel_fungi <- unique(comb_auc_df_$fungi[comb_auc_df_$fungi %in% euk_sel_mod_list & comb_auc_df_$median_auc > 55])

auc_df <- matrix(data = NA, nrow = 0, ncol = 3)
for (sel_single_feature in median_sel_fungi) {
    roc_list <- list()
    for (c in cohort_lst) {  
        sel_imp_df <- cn_combin_euk[, c('Stage', sel_single_feature)]
        train_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort != c], ]
        valda_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort == c], ]
        set.seed(best_seed_df$Seed[best_seed_df$Cohort == c & best_seed_df$Feature == sel_single_feature])
        # set.seed(123)
        rf <- randomForest(Stage ~ ., data = train_set)
        pred <- predict(rf, newdata=valda_set[-1], type = 'prob')[,2]
        pred_result <- prediction(pred, valda_set[,1])
        auc_value <- round(performance(pred_result, measure = "auc")@y.values[[1]]*100, 2)
        roc_list[[c]] <- roc(valda_set[,1], pred, legacy.axes = T, percent = T)
    }
    roc_pdf <- FileCreate(DirPath = "../11.Classifier/RF/single_Comparison/", 
                          Prefix = paste0('ROC-', sel_single_feature), 
                          Suffix = 'pdf')
    pdf(file = roc_pdf)
    par(pty = "s")
    plot(roc_list[[cohort_lst[1]]], col = my_color[cohort_lst[1]], lwd = 2, legacy.axes = T, 
         xlab = "False Postive Percentage", ylab = "True Postive Percentage")
    for(c in cohort_lst[2:length(cohort_lst)]){
        lines(roc_list[[c]], col = my_color[c], lwd = 2)
    }
    dev.off()   
}

```

## Random Forest 

random forest, leave one set out.


## Rank Features LOSO By Importance

[link](https://stats.stackexchange.com/questions/443234/random-forest-in-r-how-to-perform-feature-extraction-and-reach-the-best-accura)

### Individual study
```{r, message = F, warning = F}

for (sub_type in c('fungi', 'bac', 'all')) {
    if (sub_type == 'fungi') {
        n_feature <- ncol(euk_otu)
        myData <- cn_combin_df[, 1:(1 + n_feature)]
    }else if (sub_type == 'bac') {
        n_feature <- ncol(bac_otu)
        myData <- cn_combin_df[, c(1, (ncol(euk_otu) + 2): ncol(cn_combin_df))]
    }else{
        n_feature <- ncol(cn_combin_df)-1
        myData <- cn_combin_df
    }

    important_score <- 50
    set.seed(123)
    ###############################Random Forest
    Importance_list <- matrix(data = NA, nrow = 0, ncol = n_feature)
    for (c in cohort_lst){
        message(sub_type, " with ", c, " start at ", date())
        # parallel
        cores=detectCores()
        cl <- makeCluster(cores[1]-1) # not to overload your computer
        registerDoParallel(cl)
        Importance_Table <- foreach (t=1:10, .combine=rbind) %dopar% {
            seed <- sample.int(10)
            set.seed(seed)  
            seeds <- vector(mode = "list", length = 50)
            for(i in 1:50){
                seeds[[i]] <- sample.int(1000, 12)
            }

            ## For the last model:
            seeds[[50]] <- sample.int(1000, 1)

            myData$Stage <- as.factor(myData$Stage)

            training <- myData[rownames(cn_meta_df)[cn_meta_df$Cohort == c],]
            testing <- myData[rownames(cn_meta_df)[cn_meta_df$Cohort != c], ] 

            train_control <- caret::trainControl(method="cv", number=10, verboseIter = TRUE, seeds = seeds) 

            model <- caret::train(training[,2:ncol(training)],
                                  as.factor(training[,1]), 
                                  method = "rf", type="classification",
                                  metric= "Accuracy", maximize= TRUE, 
                                  trControl = train_control, importance = TRUE) 


            Adding_columns <- NULL
            varImp2 <- caret::varImp(model, scale = TRUE)
            Adding_columns <- t(varImp2$importance)
            rownames(Adding_columns) <- paste0(rownames(Adding_columns),".", t)
            tmp <- Adding_columns[1, ,drop = F] 
        }
        stopCluster(cl) #stop cluster   
        Importance_Table_csv <- FileCreate(DirPath = paste0('../11.Classifier/RF/FeatureSel_Important/', sub_type), 
                                           Prefix = paste0('Importance_Table_', c), Suffix = 'csv')
        Importance_Table_Mean <- t(apply(Importance_Table, MARGIN = 2, function(x) mean(x, na.rm=TRUE)))
        rownames(Importance_Table_Mean) <- paste0(c, "_Mean")
        Importance_list <- rbind(Importance_list, Importance_Table_Mean)
        Importance_Table_2 <- as.data.frame(Importance_Table)
        Importance_Table_2 <- rbind(Mean=Importance_Table_Mean, Importance_Table_2)
        write.csv(x = Importance_Table_2, file = Importance_Table_csv)

        Importance_Table_Filter <- as.data.frame(Importance_Table_Mean)
        Importance_Table_Filter2 <- Importance_Table_Filter[,Importance_Table_Filter< important_score]
        Importance_Table_Filter3 <- colnames(Importance_Table_Filter2) 
        Excluding_Channels <- names(myData) %in% Importance_Table_Filter3
        myData_filter <- myData[!Excluding_Channels]
        Importance_filter_csv <- FileCreate(DirPath = paste0('../11.Classifier/RF/FeatureSel_Important/', sub_type), 
                                           Prefix = paste0('Importance_feature_', c, "_", important_score), Suffix = 'csv')
        write.csv(x = Importance_Table_Filter[,Importance_Table_Filter>= important_score], file = Importance_filter_csv)
    }

    Importance_df2 <- as.data.frame(t(Importance_list))
    for (c in cohort_lst) {
        Importance_df2[[paste0(c, "_Rank")]] <- rank(-Importance_df2[[paste0(c, "_Mean")]])
    }
    rank_list <- grep('_Rank', colnames(Importance_df2))
    Importance_df2$Rank3 <- apply(Importance_df2[, rank_list] , 1,
                                  function(x){ifelse(any(x <= 3), 'Sel', "Ign")})
    Importance_df2$Rank5 <- apply(Importance_df2[, rank_list] , 1, 
                                  function(x){ifelse(any(x <= 5), 'Sel', "Ign")})
    Importance_df2$Rank10 <- apply(Importance_df2[, rank_list] , 1, 
                                   function(x){ifelse(any(x <= 10), 'Sel', "Ign")})

    Importance_df2_csv <-  FileCreate(DirPath = paste0('../11.Classifier/RF/FeatureSel_Important/', sub_type), 
                                     Prefix = 'Combine_importantScore_rank_all', Suffix = 'csv')
    write.csv(Importance_df2, Importance_df2_csv)

}

```

### Combined all cohorts
```{r, message = F, warning = F}
###############################Random Forest (merge all cohorts as train)
for (sub_type in c("all", "bac", "fungi")) {
  if (sub_type == 'fungi') {
      n_feature <- ncol(euk_otu)
      myData <- cn_combin_df[, 1:(1 + n_feature)]
  }else if (sub_type == 'bac') {
      n_feature <- ncol(bac_otu)
      myData <- cn_combin_df[, c(1, (ncol(euk_otu) + 2): ncol(cn_combin_df))]
  }else{
      n_feature <- ncol(cn_combin_df)-1
      myData <- cn_combin_df
  }
  Importance_list <- matrix(data = NA, nrow = 0, ncol = n_feature)
  
  set.seed(123)
  cores=detectCores()
  cl <- makeCluster(cores[1]-1) # not to overload your computer
  registerDoParallel(cl)
  Importance_Table <- foreach (t=1:10, .combine=rbind) %dopar% {
    seed <- sample.int(10)
    set.seed(seed)  
    seeds <- vector(mode = "list", length = 50)
    for(i in 1:50){
      seeds[[i]] <- sample.int(1000, 12)
    }
    
    ## For the last model:
    seeds[[50]] <- sample.int(1000, 1)
  
    myData$Stage <- as.factor(myData$Stage)
    
    training <- myData
    
    train_control <- caret::trainControl(method="cv", number=10, verboseIter = TRUE, seeds = seeds) 
    
    model <- caret::train(training[,2:ncol(training)],
                          as.factor(training[,1]), 
                          method = "rf", type="classification",
                          metric= "Accuracy", maximize= TRUE, 
                          trControl = train_control, importance = TRUE) 
    
    
    Adding_columns <- NULL
    varImp2 <- caret::varImp(model, scale = TRUE)
    Adding_columns <- t(varImp2$importance)
    rownames(Adding_columns) <- paste0(rownames(Adding_columns),".", t)
    Importance_Table <- rbind(Importance_Table, Adding_columns[1, ,drop = F])
  }
  stopCluster(cl) #stop cluster   
  Importance_Table_csv <- FileCreate(DirPath = paste0('../11.Classifier/RF/FeatureSel_Important/'), 
                                     Prefix = paste0('Importance_Table_',sub_type), Suffix = 'csv')
  Importance_Table_Mean <- t(apply(Importance_Table, MARGIN = 2, function(x) mean(x, na.rm=TRUE)))
  rownames(Importance_Table_Mean) <- "all_Mean"
  Importance_list <- rbind(Importance_list, Importance_Table_Mean)
  Importance_Table_2 <- as.data.frame(Importance_Table)
  Importance_Table_2 <- rbind(Mean=Importance_Table_Mean, Importance_Table_2)
  Importance_Table_3 <- as.data.frame(t(Importance_Table_2))
  Importance_Table_3$Rank <- rank(-Importance_Table_3$Mean)
  write.csv(x = Importance_Table_3, file = Importance_Table_csv)
}
```

### Train model
```{r, message = F, warning = F}

for (sub_type in c('fungi', 'bac', 'all')) {
# for (sub_type in c('bac', 'all')) {
  Importance_df2 <- ImportTable(file = paste0('../11.Classifier/RF/FeatureSel_Important/',
                                              sub_type, '/2022-01-19-Combine_importantScore_rank_all-v1.0.csv'), 
                                sep = ',')
  sel_all_list <- rownames(Importance_df2)[Importance_df2$Rank3 == 'Sel']
  sel_imp_df <- cn_combin_df[, c('Stage', sel_all_list)]
  message(sub_type, " ", date())
  auc_mean <- 0
  auc_seed <- 1
  cores=detectCores()
  cl <- makeCluster(cores[1]-1) # not to overload your computer
  registerDoParallel(cl)
  auc_df <- foreach (k=1:100, .combine=rbind) %dopar% {
    auc_tmp <- NULL
    for (c in cohort_lst) {
          train_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort != c], ]
          valda_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort == c], ]
          set.seed(k)
          # set.seed(123)
          rf <- randomForest::randomForest(Stage ~ ., data = train_set)
          pred <- stats::predict(rf, newdata=valda_set[-1], type = 'prob')[,2]
          pred_result <- ROCR::prediction(pred, valda_set[,1])
          auc_value <- round(ROCR::performance(pred_result, measure = "auc")@y.values[[1]]*100, 2)
          # message(c, "\t", auc_value, "%")
          auc_tmp <- c(auc_tmp, auc_value)
        }
        tmp <- matrix(c(auc_tmp, mean(auc_tmp)),nrow = 1)
        # if (mean(auc_tmp) > auc_mean) {
        #   message(sub_type, "\t", k, "\t", round(mean(auc_tmp), 2), "%")
        # }
        auc_seed <- ifelse(mean(auc_tmp) > auc_mean, k, auc_seed)
        auc_mean <- ifelse(mean(auc_tmp) > auc_mean, mean(auc_tmp), auc_mean)
        tmp
    }
    stopCluster(cl)
    rownames(auc_df) <- paste0("Seed_", 1:100)
    colnames(auc_df) <- c(cohort_lst, "Mean")
    auc_df_csv <- FileCreate(DirPath = paste0('../11.Classifier/RF/AUC_res/'), 
                           Prefix = paste0('AUC_matrix-', sub_type), Suffix = 'csv')
    write.csv(auc_df, auc_df_csv)
}

```




```{r, message = F, warning = F}
auc_comp <- matrix(data = NA, nrow = 0, ncol = 2)
for (sub_type in c('fungi', 'bac', 'all')) {
  auc_df <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2022-01-19-AUC_matrix-',
                                              sub_type, '-v1.0.csv'))
  auc_comp <- rbind(auc_comp, cbind(auc_df, Group = rep(sub_type, 100)))
}

wilcox.test(Mean ~ Group, data = auc_comp, subset = Group %in% c('all', 'bac'))
wilcox.test(Mean ~ Group, data = auc_comp, subset = Group %in% c('all', 'fungi'))
wilcox.test(Mean ~ Group, data = auc_comp, subset = Group %in% c('fungi', 'bac'))
```

## Plot Heatmap

```{r, message = F, warning = F}
Importance_df2 <- ImportTable(file = '../11.Classifier/RF/FeatureSel_Important/all/2022-01-19-Combine_importantScore_rank_all-v1.0.csv')
subset_imp <- Importance_df2[Importance_df2$Rank3 == 'Sel', ]
subset_imp2 <- subset_imp[, grep('_Rank', colnames(subset_imp))]

fungi_median <- ImportTable(file = '../07.FeatureSelection/01.SSTF/2021-07-30-walsh-median-all-v1.0.0.csv')
rownames(fungi_median) <- gsub("\\W+", '_', rownames(fungi_median))
rownames(fungi_median) <- gsub("^_", '', rownames(fungi_median))
fungi_median$EnrDep <- ifelse(fungi_median$`all-Median` >0, 'Enriched', 'Depleted')
fungi_median$Kingdoms <- 'fungi'

bac_median <- ImportTable(file = '../09.Bacteria/2021-08-06-wilcoxon-all-v1.0.0.csv')
rownames(bac_median) <- gsub("\\W+", '_', rownames(bac_median))
rownames(bac_median) <- gsub("^_", '', rownames(bac_median))
bac_median$EnrDep <- ifelse(bac_median$mFC >1, 'Enriched', 'Depleted')
bac_median$Kingdoms <- 'bac'

combine_EnrDep <- rbind(fungi_median[, c(5, 6), drop = F], bac_median[, c(5, 6), drop = F])
combine_EnrDep <- combine_EnrDep[rownames(subset_imp), , drop = F]

comb_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2022-01-19-AUC_matrix-all-v1.0.csv'))
bac_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2022-01-19-AUC_matrix-bac-v1.0.csv'))
fungi_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2022-01-19-AUC_matrix-fungi-v1.0.csv'))
summary_AUC <- rbind(Fun_AUC = fungi_AUC[which.max(fungi_AUC$Mean),],
                     rbind(Bac_AUC = bac_AUC[which.max(bac_AUC$Mean),], 
                           Comb_AUC = comb_AUC[which.max(comb_AUC$Mean),]))

summary_AUC <- rbind(summary_AUC, Ratio = (summary_AUC[3,]/summary_AUC[2,]-1)*100)

all_Feature_imp <- ImportTable(file ="../11.Classifier/RF/FeatureSel_Important/2022-01-19-Importance_Table_all-v1.0.csv")
all_Feature_imp1 <- all_Feature_imp[rownames(subset_imp), "Rank", drop = F]
all_Feature_imp2 <- cbind(all_Feature_imp1, combine_EnrDep)
all_Feature_imp2 <- all_Feature_imp2[order(all_Feature_imp2$Rank,all_Feature_imp2$EnrDep), ]


## plot enriched sub
enrich_mt <- subset_imp2[rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched'],]
enrich_col <- colorRamp2(c(0, 30), c('#bc0025', '#ffffcc'))

enrich_all_list <- all_Feature_imp2$Rank[all_Feature_imp2$EnrDep == 'Enriched']
names(enrich_all_list) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched']

enrich_all_list_label <- as.character(enrich_all_list)
names(enrich_all_list_label) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched']

anno_col_enrich <- enrich_col(enrich_all_list)
names(anno_col_enrich) <- enrich_all_list
enrich_label_col <- lapply(enrich_all_list, function(x){ifelse(x<=3, "white", "black")}) %>% unlist

enrich_rowname_col <- lapply(all_Feature_imp2$Kingdoms[all_Feature_imp2$EnrDep == 'Enriched'],
                             function(x){ifelse(x == 'fungi', "#ff410d", "#014586")})%>% unlist

enrich_row_anno <- rowAnnotation(all = anno_simple(enrich_all_list, 
                                                   pch = enrich_all_list_label,
                                                   col = anno_col_enrich,
                                                   gp = gpar(col = "black", lwd = 2),
                                                   simple_anno_size = unit(1, "cm"),
                                                   pt_gp = gpar(fontsize = 12, fontface = 'bold',
                                                                col = enrich_label_col),
                                                   pt_size = unit(5, "mm")), annotation_label = c(""))
h1 <- Heatmap(enrich_mt, col = enrich_col, name = "CRC-enriched\nSpecies",
              cluster_columns = F, cluster_rows = F, border = T, 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(enrich_mt[i, j], x, y, gp = gpar(fontsize = 12, fontface = 'bold', 
                                    col = ifelse(enrich_mt[i, j]<=3,'white', 'black')))},
              width = ncol(enrich_mt)*unit(10, "mm"), 
              height = nrow(enrich_mt)*unit(10, "mm"),
              right_annotation = enrich_row_anno,
              row_names_gp = gpar(col = enrich_rowname_col, fontface = 4))

h1




## plot depleted sub

deplet_mt <- subset_imp2[rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted'],]
deplet_col <- colorRamp2(c(0, 30), c('#243493', '#ffffcc'))

deplet_all_list <- all_Feature_imp2$Rank[all_Feature_imp2$EnrDep == 'Depleted']
names(deplet_all_list) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted']

deplet_all_list_label <- as.character(deplet_all_list)
names(deplet_all_list_label) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted']

anno_col_deplet <- deplet_col(deplet_all_list)
names(anno_col_deplet) <- deplet_all_list
deplet_label_col <- lapply(deplet_all_list, function(x){ifelse(x<=3, "white", "black")}) %>% unlist

deplet_rowname_col <- lapply(all_Feature_imp2$Kingdoms[all_Feature_imp2$EnrDep == 'Depleted'],
                             function(x){ifelse(x == 'fungi', "#ff410d", "#014586")})%>% unlist

deplet_row_anno <- rowAnnotation(all = anno_simple(deplet_all_list, 
                                                   pch = deplet_all_list_label,
                                                   col = anno_col_deplet,
                                                   gp = gpar(col = "black", lwd = 2),
                                                   simple_anno_size = unit(1, "cm"),
                                                   pt_gp = gpar(fontsize = 12, fontface = 'bold',
                                                                col = deplet_label_col),
                                                   pt_size = unit(5, "mm")))

h2 <- Heatmap(deplet_mt, col = deplet_col, name = "CRC-depleted\nSpecies",
              cluster_columns = F, cluster_rows = F, border = T, 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(deplet_mt[i, j], x, y, 
                          gp = gpar(fontsize = 12, fontface = 'bold', 
                                    col = ifelse(deplet_mt[i, j]<=3,'white', 'black')))},
              width = ncol(deplet_mt)*unit(10, "mm"), 
              height = nrow(deplet_mt)*unit(10, "mm"), 
              right_annotation = deplet_row_anno,
              row_names_gp = gpar(col = deplet_rowname_col, fontface = 4))
h2

AUC_col <- colorRamp2(c(65, 70,75,80,85,90), c('snow', 'khaki1', 'yellow', 'red', 'red3', 'red4'))

h3 <- Heatmap(summary_AUC[c(1:3), c(1:8), drop =F],
              cluster_columns = F, cluster_rows = F, 
              col = AUC_col, split = c("", " ", "  "), 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(round(summary_AUC[i, j],2), x, y, 
                          gp = gpar(fontsize = 8, fontface = 'bold', 
                                    col = ifelse(summary_AUC[i, j] > 80, "white",'black')))},
              height = unit(30, "mm"))

# h4 <- Heatmap(summary_AUC[2, c(1:8), drop =F],
#               cluster_columns = F, col = AUC_col, 
#               rect_gp = gpar(col = "black", lwd = 2),
#               cell_fun = function(j, i, x, y, width, height, fill){
#                 grid.text(round(summary_AUC[2, j],2), x, y, 
#                           gp = gpar(fontsize = 8, fontface = 'bold', 
#                                     col = ifelse(summary_AUC[2, j] > 80, "white",'black')))},
#               height = unit(10, "mm"))


Ratio_col <- colorRamp2(c(-1, 0, 5, 10), c('darkblue', 'white', 'violetred', '#810f7b'))

h5 <- Heatmap(summary_AUC[4, c(1:8), drop =F],
              cluster_columns = F, col = Ratio_col, 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(round(abs(summary_AUC[4, j]),2), x, y, 
                          gp = gpar(fontsize = 8, fontface = 'bold', 
                                    col = ifelse(abs(summary_AUC[4, j]) > 5, "white",'black')))},
              height = unit(10, "mm"))

# ht_list <- h1 %v% h2 %v% h3 %v% h4 %v% h5
ht_list <- h1 %v% h2 %v% h3  %v% h5

#ht_pdf <- FileCreate(DirPath = '../11.Classifier/RF/Plot', Prefix = 'heatmap_RF_AUC', Suffix = 'pdf')

#pdf(file = ht_pdf, width = 15, height = 15)
draw(ht_list)
#dev.off()

```




## Plot AUC

```{r, message = F, warning = F}

comb_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-all-v1.0.0.csv'))
bac_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-bac-v1.0.0.csv'))
fungi_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-fungi-v1.0.0.csv'))

best_seed_list <- list()
best_seed_list[['all']] <- which.max(comb_AUC$Mean)
best_seed_list[['bac']] <- which.max(bac_AUC$Mean)
best_seed_list[['fungi']] <- which.max(fungi_AUC$Mean)


roc_comp_df <- matrix(data = NA, nrow = 1, ncol = 2)
roc_comp_df <- roc_comp_df[-1,]
for (c in cohort_lst) {
  roc_list <- list()
  for (sub_type in c('fungi', 'bac', 'all')) {
    Importance_df2 <- ImportTable(file = paste0('../11.Classifier/RF/FeatureSel_Important/',
                                                sub_type, '/2021-09-29-Combine_importantScore_rank_all-v1.0.0.csv'), 
                                  sep = ',')
    sel_all_list <- rownames(Importance_df2)[Importance_df2$Rank3 == 'Sel']
    sel_imp_df <- cn_combin_core[, c('Stage', sel_all_list)]
    train_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort != c], ]
    valda_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort == c], ]
    set.seed(best_seed_list[[sub_type]])
    # set.seed(123)
    rf <- randomForest(Stage ~ ., data = train_set)
    pred <- predict(rf, newdata=valda_set[-1], type = 'prob')[,2]
    pred_result <- prediction(pred, valda_set[,1])
    auc_value <- round(performance(pred_result, measure = "auc")@y.values[[1]]*100, 2)
    
    roc_list[[sub_type]] <- roc(valda_set[,1], pred, legacy.axes = T, percent = T)
  }
  roc_pdf <- FileCreate(DirPath = "../11.Classifier/RF/Comparison/", Prefix = paste0('ROC-', c), Suffix = 'pdf')
  pdf(file = roc_pdf)
  par(pty = "s")
  plot(roc_list$all, col = '#ffd320', lwd = 4, legacy.axes = T, 
       xlab = "False Postive Percentage", ylab = "True Postive Percentage")
  # lines(roc_list$fungi, col = '#ff420e', lwd = 2, lty = 1)
  lines(roc_list$bac, col = '#004586', lwd = 2, lty = 1)
  dev.off()
  roc_comp <- roc.test(roc_list$all, roc_list$bac)
  roc_comp_df <- rbind(roc_comp_df, c(roc_comp$p.value, roc_comp$statistic))
  
}
roc_comp_df <- as.data.frame(roc_comp_df)
rownames(roc_comp_df) <- cohort_lst
colnames(roc_comp_df) <- c("p-value", "Z")

roc_comp_csv <- FileCreate(DirPath = "../11.Classifier/RF/Comparison", Prefix = "ROC_Comparison", Suffix = 'csv')
write.csv(roc_comp_df, roc_comp_csv)

#ff420e   fungi
#004586   bac
#ffd320   combine


```


## Supplementary Heatmap

### Bac Heatmap

```{r, message = F, warning = F}
Importance_df2 <- ImportTable(file = '../11.Classifier/RF/FeatureSel_Important/bac/2021-09-29-Combine_importantScore_rank_all-v1.0.0.csv')
subset_imp <- Importance_df2[Importance_df2$Rank3 == 'Sel', ]
subset_imp2 <- subset_imp[, grep('_Rank', colnames(subset_imp))]


bac_median <- ImportTable(file = '../09.Bacteria/2021-08-06-wilcoxon-all-v1.0.0.csv')
rownames(bac_median) <- gsub("\\W+", '_', rownames(bac_median))
rownames(bac_median) <- gsub("^_", '', rownames(bac_median))
bac_median$EnrDep <- ifelse(bac_median$mFC >1, 'Enriched', 'Depleted')
bac_median$Kingdoms <- 'bac'

subset_imp2 <- subset_imp2[rownames(subset_imp2) %in% rownames(bac_median), ]


combine_EnrDep <- bac_median[, c(5, 6), drop = F]
combine_EnrDep <- combine_EnrDep[rownames(subset_imp2), , drop = F]

comb_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-all-v1.0.0.csv'))
bac_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-bac-v1.0.0.csv'))
fungi_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-fungi-v1.0.0.csv'))
summary_AUC <- rbind(Fun_AUC = fungi_AUC[which.max(fungi_AUC$Mean),],
                     rbind(Bac_AUC = bac_AUC[which.max(bac_AUC$Mean),], 
                           Comb_AUC = comb_AUC[which.max(comb_AUC$Mean),]))


all_Feature_imp <- ImportTable(file ="../11.Classifier/RF/FeatureSel_Important/2021-09-30-Importance_Table_bac-v1.0.0.csv")
all_Feature_imp1 <- all_Feature_imp[rownames(subset_imp2), "Rank", drop = F]
all_Feature_imp2 <- cbind(all_Feature_imp1, combine_EnrDep)
all_Feature_imp2 <- all_Feature_imp2[order(all_Feature_imp2$Rank,all_Feature_imp2$EnrDep), ]


## plot enriched sub
enrich_mt <- subset_imp2[rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched'],]
enrich_col <- colorRamp2(c(0, 30), c('#bc0025', '#ffffcc'))

enrich_all_list <- all_Feature_imp2$Rank[all_Feature_imp2$EnrDep == 'Enriched']
names(enrich_all_list) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched']

enrich_all_list_label <- as.character(enrich_all_list)
names(enrich_all_list_label) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched']

anno_col_enrich <- enrich_col(enrich_all_list)
names(anno_col_enrich) <- enrich_all_list
enrich_label_col <- lapply(enrich_all_list, function(x){ifelse(x<=3, "white", "black")}) %>% unlist

enrich_rowname_col <- lapply(all_Feature_imp2$Kingdoms[all_Feature_imp2$EnrDep == 'Enriched'],
                             function(x){ifelse(x == 'fungi', "#ff410d", "#014586")})%>% unlist

enrich_row_anno <- rowAnnotation(all = anno_simple(enrich_all_list, 
                                                   pch = enrich_all_list_label,
                                                   col = anno_col_enrich,
                                                   gp = gpar(col = "black", lwd = 2),
                                                   simple_anno_size = unit(1, "cm"),
                                                   pt_gp = gpar(fontsize = 12, fontface = 'bold',
                                                                col = enrich_label_col),
                                                   pt_size = unit(5, "mm")), annotation_label = c(""))

colnames(enrich_mt) <- c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", "2019_ThomasA", "2019_WirbelJ", "2019_YachidaS", "2017_JunY", "2021_JunY")
h1 <- Heatmap(enrich_mt, col = enrich_col, name = "CRC-enriched\nSpecies",
              cluster_columns = F, cluster_rows = F, border = T, 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(enrich_mt[i, j], x, y, gp = gpar(fontsize = 12, fontface = 'bold', 
                                    col = ifelse(enrich_mt[i, j]<=3,'white', 'black')))},
              width = ncol(enrich_mt)*unit(10, "mm"), 
              height = nrow(enrich_mt)*unit(10, "mm"),
              right_annotation = enrich_row_anno,
              row_names_gp = gpar(col = enrich_rowname_col, fontface = 4))

h1




## plot depleted sub

deplet_mt <- subset_imp2[rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted'],]
deplet_col <- colorRamp2(c(0, 30), c('#243493', '#ffffcc'))

deplet_all_list <- all_Feature_imp2$Rank[all_Feature_imp2$EnrDep == 'Depleted']
names(deplet_all_list) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted']

deplet_all_list_label <- as.character(deplet_all_list)
names(deplet_all_list_label) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted']

anno_col_deplet <- deplet_col(deplet_all_list)
names(anno_col_deplet) <- deplet_all_list
deplet_label_col <- lapply(deplet_all_list, function(x){ifelse(x<=3, "white", "black")}) %>% unlist

deplet_rowname_col <- lapply(all_Feature_imp2$Kingdoms[all_Feature_imp2$EnrDep == 'Depleted'],
                             function(x){ifelse(x == 'fungi', "#ff410d", "#014586")})%>% unlist

deplet_row_anno <- rowAnnotation(all = anno_simple(deplet_all_list, 
                                                   pch = deplet_all_list_label,
                                                   col = anno_col_deplet,
                                                   gp = gpar(col = "black", lwd = 2),
                                                   simple_anno_size = unit(1, "cm"),
                                                   pt_gp = gpar(fontsize = 12, fontface = 'bold',
                                                                col = deplet_label_col),
                                                   pt_size = unit(5, "mm")))

colnames(deplet_mt) <- c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", "2019_ThomasA", "2019_WirbelJ", "2019_YachidaS", "2017_JunY", "2021_JunY")
h2 <- Heatmap(deplet_mt, col = deplet_col, name = "CRC-depleted\nSpecies",
              cluster_columns = F, cluster_rows = F, border = T, 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(deplet_mt[i, j], x, y, 
                          gp = gpar(fontsize = 12, fontface = 'bold', 
                                    col = ifelse(deplet_mt[i, j]<=3,'white', 'black')))},
              width = ncol(deplet_mt)*unit(10, "mm"), 
              height = nrow(deplet_mt)*unit(10, "mm"), 
              right_annotation = deplet_row_anno,
              row_names_gp = gpar(col = deplet_rowname_col, fontface = 4))
h2

h3 <- Heatmap(summary_AUC[, c(1:8), drop =F], col = AUC_col, 
              cluster_columns = F, cluster_rows = F, split = c("", " ", "  "),
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(round(summary_AUC[i, j],2), x, y, 
                          gp = gpar(fontsize = 8, fontface = 'bold', 
                                    col = ifelse(summary_AUC[i, j] > 80, "white",'black')))},
              height = unit(30, "mm"))

# h3 <- Heatmap(summary_AUC[2, c(1:8), drop =F],
#               cluster_columns = F, col = AUC_col, 
#               rect_gp = gpar(col = "black", lwd = 2),
#               cell_fun = function(j, i, x, y, width, height, fill){
#                 grid.text(round(summary_AUC[2, j],2), x, y, 
#                           gp = gpar(fontsize = 8, fontface = 'bold', 
#                                     col = ifelse(summary_AUC[2, j] > 80, "white",'black')))},
#               height = unit(10, "mm"))


ht_list <- h1 %v% h2 %v% h3

ht_pdf <- FileCreate(DirPath = '../11.Classifier/RF/Plot', Prefix = 'supplementary-Bac-heatmap_RF_AUC', Suffix = 'pdf')

pdf(file = ht_pdf, width = 15, height = 15)
draw(ht_list)
dev.off()

```

### Fungi Heatmap

```{r, message = F, warning = F}
Importance_df2 <- ImportTable(file = '../11.Classifier/RF/FeatureSel_Important/fungi/2021-09-29-Combine_importantScore_rank_all-v1.0.0.csv')
subset_imp <- Importance_df2[Importance_df2$Rank3 == 'Sel', ]
subset_imp2 <- subset_imp[, grep('_Rank', colnames(subset_imp))]


fungi_median <- ImportTable(file = '../07.FeatureSelection/01.SSTF/2021-07-30-walsh-median-all-v1.0.0.csv')
rownames(fungi_median) <- gsub("\\W+", '_', rownames(fungi_median))
rownames(fungi_median) <- gsub("^_", '', rownames(fungi_median))
fungi_median$EnrDep <- ifelse(fungi_median$`all-Median` >0, 'Enriched', 'Depleted')
fungi_median$Kingdoms <- 'fungi'

subset_imp2 <- subset_imp2[(rownames(subset_imp2) %in% rownames(fungi_median)), ]


combine_EnrDep <- fungi_median[, c(5, 6), drop = F]
combine_EnrDep <- combine_EnrDep[rownames(subset_imp2), , drop = F]

comb_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-all-v1.0.0.csv'))
bac_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-bac-v1.0.0.csv'))
fungi_AUC <- ImportTable(file = paste0('../11.Classifier/RF/AUC_res/2021-09-29-AUC_matrix-fungi-v1.0.0.csv'))
summary_AUC <- rbind(Fun_AUC = fungi_AUC[which.max(fungi_AUC$Mean),],
                     rbind(Bac_AUC = bac_AUC[which.max(bac_AUC$Mean),], 
                           Comb_AUC = comb_AUC[which.max(comb_AUC$Mean),]))

all_Feature_imp <- ImportTable(file ="../11.Classifier/RF/FeatureSel_Important/2021-09-30-Importance_Table_fungi-v1.0.0.csv")
all_Feature_imp1 <- all_Feature_imp[rownames(subset_imp2), "Rank", drop = F]
all_Feature_imp2 <- cbind(all_Feature_imp1, combine_EnrDep)
all_Feature_imp2 <- all_Feature_imp2[order(all_Feature_imp2$Rank,all_Feature_imp2$EnrDep), ]


## plot enriched sub
enrich_mt <- subset_imp2[rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched'],]
enrich_col <- colorRamp2(c(0, 30), c('#bc0025', '#ffffcc'))

enrich_all_list <- all_Feature_imp2$Rank[all_Feature_imp2$EnrDep == 'Enriched']
names(enrich_all_list) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched']

enrich_all_list_label <- as.character(enrich_all_list)
names(enrich_all_list_label) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Enriched']

anno_col_enrich <- enrich_col(enrich_all_list)
names(anno_col_enrich) <- enrich_all_list
enrich_label_col <- lapply(enrich_all_list, function(x){ifelse(x<=3, "white", "black")}) %>% unlist

enrich_rowname_col <- lapply(all_Feature_imp2$Kingdoms[all_Feature_imp2$EnrDep == 'Enriched'],
                             function(x){ifelse(x == 'fungi', "#ff410d", "#014586")})%>% unlist

enrich_row_anno <- rowAnnotation(all = anno_simple(enrich_all_list, 
                                                   pch = enrich_all_list_label,
                                                   col = anno_col_enrich,
                                                   gp = gpar(col = "black", lwd = 2),
                                                   simple_anno_size = unit(1, "cm"),
                                                   pt_gp = gpar(fontsize = 12, fontface = 'bold',
                                                                col = enrich_label_col),
                                                   pt_size = unit(5, "mm")), annotation_label = c(""))

colnames(enrich_mt) <- c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", "2019_ThomasA", "2019_WirbelJ", "2019_YachidaS", "2017_JunY", "2021_JunY")
h1 <- Heatmap(enrich_mt, col = enrich_col, name = "CRC-enriched\nSpecies",
              cluster_columns = F, cluster_rows = F, border = T, 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(enrich_mt[i, j], x, y, gp = gpar(fontsize = 12, fontface = 'bold', 
                                    col = ifelse(enrich_mt[i, j]<=3,'white', 'black')))},
              width = ncol(enrich_mt)*unit(10, "mm"), 
              height = nrow(enrich_mt)*unit(10, "mm"),
              right_annotation = enrich_row_anno,
              row_names_gp = gpar(col = enrich_rowname_col, fontface = 4))

h1




## plot depleted sub

deplet_mt <- subset_imp2[rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted'],]
deplet_col <- colorRamp2(c(0, 30), c('#243493', '#ffffcc'))

deplet_all_list <- all_Feature_imp2$Rank[all_Feature_imp2$EnrDep == 'Depleted']
names(deplet_all_list) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted']

deplet_all_list_label <- as.character(deplet_all_list)
names(deplet_all_list_label) <- rownames(all_Feature_imp2)[all_Feature_imp2$EnrDep == 'Depleted']

anno_col_deplet <- deplet_col(deplet_all_list)
names(anno_col_deplet) <- deplet_all_list
deplet_label_col <- lapply(deplet_all_list, function(x){ifelse(x<=3, "white", "black")}) %>% unlist

deplet_rowname_col <- lapply(all_Feature_imp2$Kingdoms[all_Feature_imp2$EnrDep == 'Depleted'],
                             function(x){ifelse(x == 'fungi', "#ff410d", "#014586")})%>% unlist

deplet_row_anno <- rowAnnotation(all = anno_simple(deplet_all_list, 
                                                   pch = deplet_all_list_label,
                                                   col = anno_col_deplet,
                                                   gp = gpar(col = "black", lwd = 2),
                                                   simple_anno_size = unit(1, "cm"),
                                                   pt_gp = gpar(fontsize = 12, fontface = 'bold',
                                                                col = deplet_label_col),
                                                   pt_size = unit(5, "mm")))

colnames(deplet_mt) <- c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", "2019_ThomasA", "2019_WirbelJ", "2019_YachidaS", "2017_JunY", "2021_JunY")
h2 <- Heatmap(deplet_mt, col = deplet_col, name = "CRC-depleted\nSpecies",
              cluster_columns = F, cluster_rows = F, border = T, 
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(deplet_mt[i, j], x, y, 
                          gp = gpar(fontsize = 12, fontface = 'bold', 
                                    col = ifelse(deplet_mt[i, j]<=3,'white', 'black')))},
              width = ncol(deplet_mt)*unit(10, "mm"), 
              height = nrow(deplet_mt)*unit(10, "mm"), 
              right_annotation = deplet_row_anno,
              row_names_gp = gpar(col = deplet_rowname_col, fontface = 4))
h2

h3 <- Heatmap(summary_AUC[, c(1:8), drop =F], col = AUC_col, 
              cluster_columns = F, cluster_rows = F, split = c("", " ", "  "),
              rect_gp = gpar(col = "black", lwd = 2),
              cell_fun = function(j, i, x, y, width, height, fill){
                grid.text(round(summary_AUC[i, j],2), x, y, 
                          gp = gpar(fontsize = 8, fontface = 'bold', 
                                    col = ifelse(summary_AUC[i, j] > 80, "white",'black')))},
              height = unit(30, "mm"))

ht_list <- h1 %v% h2 %v% h3

ht_pdf <- FileCreate(DirPath = '../11.Classifier/RF/Plot', Prefix = 'supplementary-Fungi-heatmap_RF_AUC', Suffix = 'pdf')

pdf(file = ht_pdf, width = 15, height = 15)
draw(ht_list)
dev.off()

```

<!--
## Single Fungi Classifiers (study-to-study transfer)

### Calculate the model (best seed)
* it will cost a lot of time
```{r , message = F, warning = F}

target_euk60 <- cn_combin_euk[, c('Stage','Aspergillus_rambellii'), drop = F] # median abundance > 60%
mydata <- target_euk60
cohort_lst <- unique(cn_meta_df$Cohort)

auc_sub_list <- list()
calculate_or_not <- F
# calculate_or_not <- T
seed_num <- 100
if (calculate_or_not) {
  My_function <- function(seed, train_set, valid_list, cohort_lst){
    set.seed(seed)
    auc_list <- rep(NA, length(cohort_lst))
    names(auc_list) <- cohort_lst
    for(single_feature in colnames(mydata)[2:ncol(mydata)]){
        set.seed(seed)
        rf <- randomForest::randomForest(as.formula(paste0("Stage ~ ", single_feature)), data = train_set)
        for (val_cohort in names(valid_list)) {
          pred <- predict(rf, newdata=valid_list[[val_cohort]][, single_feature, drop = F], type = 'prob')[,2]
          pred_result <- ROCR::prediction(pred, valid_list[[val_cohort]][,1])
          auc_list[val_cohort] <- round(ROCR::performance(pred_result, measure = "auc")@y.values[[1]]*100, 2)
        }
    }
    auc_list <- c(auc_list, seed = sel_seed, species)
    return(matrix(data = auc_list, nrow = 1, dimnames = list(c(paste0('Seed_', sel_seed)), 
                                                             colnames(mydata)[2:ncol(mydata)])))
  }
  for (sub_cohort in cohort_lst) {    
    train_set <- mydata[rownames(cn_meta_df)[cn_meta_df$Cohort == sub_cohort], ]
    valid_list <- list()
    for (val_cohort in cohort_lst) {
      if (val_cohort == sub_cohort) next
      valid_list[[val_cohort]] <- mydata[rownames(cn_meta_df)[cn_meta_df$Cohort == val_cohort], ]
    }
    cores=detectCores()
    cl <- makeCluster(cores[1]-1) #not to overload your computer
    registerDoParallel(cl)
    # parallel
    auc_res <- foreach(sel_seed = 1:seed_num, .combine=rbind) %dopar%{
        tmp_res <- My_function(seed = sel_seed, train_set = train_set, valid_list = valid_list, cohort_lst = cohort_lst)
    }
    stopCluster(cl) #stop cluster   
    roc_csv <- FileCreate(DirPath = "../11.Classifier/RF/single_AUC", Suffix = "csv",
                          Prefix = paste0("SingleFeature_AUC_Performance-", sub_cohort))
    write.csv(x = auc_res, file = roc_csv)
    auc_sub_list[[sub_cohort]] <- auc_res
  }
}else{
  for (sub_cohort in cohort_lst) {
    auc_sub_list[[sub_cohort]] <- ImportTable(file = paste0( '../11.Classifier/RF/single_AUC/2022-01-18-SingleFeature_AUC_Performance-',sub_cohort,'-v1.0.csv'))
  }
}



```

### select the best seed for each cohort
```{r , message = F, warning = F}
best_seed_df <- matrix(NA, nrow = 1, ncol = 4)
best_seed_df <- best_seed_df[-1, ]
for (sub_cohort in cohort_lst) {
  tmp_df <- auc_sub_list[[sub_cohort]]
  for (single_feature in colnames(mydata)[2:ncol(mydata)]) {
    best_seed_df <- rbind(best_seed_df, 
                          c(sub_cohort, single_feature,
                            which.max(tmp_df[, single_feature]),
                            max(tmp_df[, single_feature])))
  }
}

best_seed_df <- as.data.frame(best_seed_df)
colnames(best_seed_df) <- c("Cohort", "Feature", "Seed", "AUC")
best_seed_df$MeanAUC <- apply(best_seed_df, 1, function(x){
  round(mean(as.numeric(best_seed_df$AUC[best_seed_df$Feature == x[2]])),2)
})

best_seed_df$Kingdoms <- rep('fungi', nrow(best_seed_df))

# save 
best_seed_csv <- FileCreate(DirPath = "../11.Classifier/RF/single_AUC",
                            Prefix = "Summary-Best_Seed_AUC", Suffix = "csv")

write.csv(x = best_seed_df, file = best_seed_csv)

```

### calculate the best model

```{r , message = F, warning = F}
all_euk_list <- colnames(cn_combin_euk)[2: ncol(cn_combin_euk)]
prj_list <- list()
auc_df <- matrix(data = NA, nrow = 0, ncol = 3)
for (c in cohort_lst) {
  
  for (sel_single_feature in all_euk_list) {
    sel_imp_df <- cn_combin_euk[, c('Stage', sel_single_feature)]
    train_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort != c], ]
    valda_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort == c], ]
    set.seed(best_seed_df$Seed[best_seed_df$Cohort == c & best_seed_df$Feature == sel_single_feature])
    # set.seed(123)
    rf <- randomForest(Stage ~ ., data = train_set)
    pred <- predict(rf, newdata=valda_set[-1], type = 'prob')[,2]
    pred_result <- prediction(pred, valda_set[,1])
    auc_value <- round(performance(pred_result, measure = "auc")@y.values[[1]]*100, 2)
    prj_list[[c]][[sel_single_feature]][['rf']] <- rf
    prj_list[[c]][[sel_single_feature]][['roc']] <- roc(valda_set[,1], pred, legacy.axes = T, percent = T)
    auc_df <- rbind(auc_df, c(cohort = c, fungi = sel_single_feature, auc = auc_value))
  }
}
auc_df <- as.data.frame(auc_df)
auc_df$auc <- as.numeric(auc_df$auc)
```

### Plot the mean of AUC

```{r , message = F, warning = F}
# mean of AUC
# mean
mean_auc_df <- aggregate(auc ~ fungi, auc_df, mean)
mean_auc_df <- mean_auc_df[order(mean_auc_df$auc, decreasing = T), ]
rownames(mean_auc_df) <- mean_auc_df$fungi
comb_auc_df <- cbind(auc_df, mean_auc = mean_auc_df[auc_df$fungi, 'auc'])
mean_auc_df2 <- mean_auc_df
mean_auc_df2$max <- unlist(lapply(mean_auc_df2$fungi, function(x){
  max(auc_df[auc_df$fungi == x, ]$auc)
}))
  
my_color <- c(ochre_palettes$namatjira_qual)
names(my_color) <-  c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", 
                      "2019_ThomasAM", "2019_WirbelJ", "2019_YachidaS", 
                      "2017_YuJ", "2021_YuJ")

comb_auc_df$cohort <- ifelse(comb_auc_df$cohort == "2014_ZellerG", "2014_ZellerG", 
                        ifelse(comb_auc_df$cohort == "2015_FengQ", "2015_FengQ", 
                               ifelse(comb_auc_df$cohort == "2016_VogtmannE", "2016_VogtmannE", 
                                      ifelse(comb_auc_df$cohort == "2019_Thom", "2019_ThomasAM", 
                                             ifelse(comb_auc_df$cohort == "2019_WirbelJ", "2019_WirbelJ", 
                                                    ifelse(comb_auc_df$cohort == "2019_Yachida", "2019_YachidaS", 
                                                           ifelse(comb_auc_df$cohort == "2021_JunY_1", "2017_YuJ","2021_YuJ")))))))

euk_sel_mod_list <- gsub("\\W+", '_', euk_sel_list)
euk_sel_mod_list <- gsub("^_", '', euk_sel_mod_list)

mean_auc_df3 <- mean_auc_df2[mean_auc_df2$fungi %in% euk_sel_mod_list,]


comb_auc_df$label <- ifelse(comb_auc_df$fungi %in% euk_sel_mod_list, 'core', 'other')
g_point <- ggplot() + 
  geom_point(data = comb_auc_df, 
             shape = 20, 
             aes(x = auc, y = mean_auc,
                 size = label, col = cohort, 
                 alpha = label)) +
  scale_color_manual(values = my_color)+
  scale_alpha_manual(values = c(core = 1, other = 0.2)) +
  scale_size_manual(values = c(core = 3, other = 2)) +
  geom_text_repel(data = mean_auc_df3[mean_auc_df3$auc > 55,], aes(y = auc, x= max,label = fungi),
                  hjust = -0.1, direction = 'x') +
  theme_bw()+
  # geom_hline(data = mean_auc_df3, aes(yintercept = auc), linetype = 2)
  geom_line(data = comb_auc_df[comb_auc_df$fungi %in% euk_sel_mod_list & comb_auc_df$mean_auc > 55, ],
            aes(x = auc, y = mean_auc, group = fungi), linetype = 2)
#ff420e   fungi
#004586   bac
#ffd320   combine

scatter_pdf <- FileCreate(DirPath = '../11.Classifier/RF/single_scatter', Prefix = 'AUC_scatter_plot_mean', Suffix = 'pdf')

pdf(file = scatter_pdf, width = 7, height = 7)
plot(g_point + theme(legend.position = "none"))
plot(g_point)
dev.off()
```

### Plot the median of AUC

```{r , message = F, warning = F}

# median AUC

# median
median_auc_df <- aggregate(auc ~ fungi, auc_df, median)
median_auc_df <- median_auc_df[order(median_auc_df$auc, decreasing = T), ]
rownames(median_auc_df) <- median_auc_df$fungi
comb_auc_df_ <- cbind(auc_df, median_auc = median_auc_df[auc_df$fungi, 'auc'])
median_auc_df2 <- median_auc_df
median_auc_df2$max <- unlist(lapply(median_auc_df2$fungi, function(x){
  max(auc_df[auc_df$fungi == x, ]$auc)
}))
  
my_color <- c(ochre_palettes$namatjira_qual)
names(my_color) <-  c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", 
                      "2019_ThomasAM", "2019_WirbelJ", "2019_YachidaS", 
                      "2017_YuJ", "2021_YuJ")

comb_auc_df_$cohort <- ifelse(comb_auc_df_$cohort == "2014_ZellerG", "2014_ZellerG", 
                        ifelse(comb_auc_df_$cohort == "2015_FengQ", "2015_FengQ", 
                               ifelse(comb_auc_df_$cohort == "2016_VogtmannE", "2016_VogtmannE", 
                                      ifelse(comb_auc_df_$cohort == "2019_Thom", "2019_ThomasAM", 
                                             ifelse(comb_auc_df_$cohort == "2019_WirbelJ", "2019_WirbelJ", 
                                                    ifelse(comb_auc_df_$cohort == "2019_Yachida", "2019_YachidaS", 
                                                           ifelse(comb_auc_df_$cohort == "2021_JunY_1", "2017_YuJ","2021_YuJ")))))))

euk_sel_mod_list <- gsub("\\W+", '_', euk_sel_list)
euk_sel_mod_list <- gsub("^_", '', euk_sel_mod_list)

median_auc_df3 <- median_auc_df2[median_auc_df2$fungi %in% euk_sel_mod_list,]


comb_auc_df_$label <- ifelse(comb_auc_df_$fungi %in% euk_sel_mod_list, 'core', 'other')
g_point_median <- ggplot() + 
  geom_point(data = comb_auc_df_, 
             shape = 20, 
             aes(x = auc, y = median_auc,
                 size = label, col = cohort, 
                 alpha = label)) +
  scale_color_manual(values = my_color)+
  scale_alpha_manual(values = c(core = 1, other = 0.2)) +
  scale_size_manual(values = c(core = 3, other = 2)) +
  geom_text(data = median_auc_df3[median_auc_df3$auc > 55,], aes(y = auc, x= max,label = fungi),
                  hjust = -0.1) +
  theme_bw()+
  # geom_hline(data = median_auc_df3, aes(yintercept = auc), linetype = 2)
  geom_line(data = comb_auc_df_[comb_auc_df_$fungi %in% euk_sel_mod_list & comb_auc_df_$median_auc > 55, ],
            aes(x = auc, y = median_auc, group = fungi), linetype = 2)
#ff420e   fungi
#004586   bac
#ffd320   combine

scatter_pdf <- FileCreate(DirPath = '../11.Classifier/RF/single_scatter', Prefix = 'AUC_scatter_plot_median', Suffix = 'pdf')

pdf(file = scatter_pdf, width = 7, height = 7)
plot(g_point_median + theme(legend.position = "none"))
plot(g_point_median)
dev.off()

```

### Draw the selected median AUC in ROC Plot 

```{r, message =F, warning = F}
my_color <- c(ochre_palettes$namatjira_qual)
names(my_color) <-  c("2014_ZellerG", "2015_FengQ", "2016_VogtmannE", 
                      "2019_Thom", "2019_WirbelJ", "2019_Yachida", 
                      "2021_JunY_1", "2021_JunY_2")
median_sel_fungi <- unique(comb_auc_df_$fungi[comb_auc_df_$fungi %in% euk_sel_mod_list & comb_auc_df_$median_auc > 55])

auc_df <- matrix(data = NA, nrow = 0, ncol = 3)
for (sel_single_feature in median_sel_fungi) {
    roc_list <- list()
    for (c in cohort_lst) {  
        sel_imp_df <- cn_combin_euk[, c('Stage', sel_single_feature)]
        train_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort != c], ]
        valda_set <- sel_imp_df[rownames(cn_meta_df)[cn_meta_df$Cohort == c], ]
        set.seed(best_seed_df$Seed[best_seed_df$Cohort == c & best_seed_df$Feature == sel_single_feature])
        # set.seed(123)
        rf <- randomForest(Stage ~ ., data = train_set)
        pred <- predict(rf, newdata=valda_set[-1], type = 'prob')[,2]
        pred_result <- prediction(pred, valda_set[,1])
        auc_value <- round(performance(pred_result, measure = "auc")@y.values[[1]]*100, 2)
        roc_list[[c]] <- roc(valda_set[,1], pred, legacy.axes = T, percent = T)
    }
    roc_pdf <- FileCreate(DirPath = "../11.Classifier/RF/single_Comparison/", 
                          Prefix = paste0('ROC-', sel_single_feature), 
                          Suffix = 'pdf')
    pdf(file = roc_pdf)
    par(pty = "s")
    plot(roc_list[[cohort_lst[1]]], col = my_color[cohort_lst[1]], lwd = 2, legacy.axes = T, 
         xlab = "False Postive Percentage", ylab = "True Postive Percentage")
    for(c in cohort_lst[2:length(cohort_lst)]){
        lines(roc_list[[c]], col = my_color[c], lwd = 2)
    }
    dev.off()   
}

```
-->








