---
title: "Selected Same Trend Features"
author: "ifanlyn@outlook.com"
date: "7/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages and subroutines {.tabset}

### Packages

```{r packages, message=FALSE, warning=FALSE}
# require(Seurat)
require(DT) # for datatable
# require(SeuratWrappers) # for fastMNN
# require(harmony) # for RunHarmony
require(reshape2) # for dcast
```

### Import table

```{r ImportTable }
ImportTable <- function(file, header = T, row.names = 1, sep = ',', check.names = FALSE, ...){
  data <- read.csv(file = file, header = header, row.names = row.names, sep = sep, check.names = check.names, ...)
  return(data)
}
```

### show_table

```{r show_table}
show_table <- function(df, rownames = T, filter="top", options = list(pageLength = 10, scrollX=T), ...){
  if (ncol(df) > 50) {
    df <- df[, 1:50]
    message('Due to the column dim is > 50, it only shows the top 50 columns.')
  }
  
  datatable(df, rownames = rownames, 
            filter=filter, options = options, ... )

}
```

### MultMedFC

MultMedFC: Multiple Median Fold Change, compared each samples between 2 Stages (Control vs CRC)

```{r}
MultMedFC <- function(data1, data2, FeatureAsRow = T){
  if (FeatureAsRow) {
    data1 <- t(data1)
    data2 <- t(data2)
  }
  data1 <- as.data.frame(data1)
  data2 <- as.data.frame(data2)
  if (! all(colnames(data1) == colnames(data2))) {
    message("rowname of data1 and data2 is not fullly same!! plz check it !!")
    return()
  }
  
  sf_df <- matrix(NA, ncol = 4); sf_df <- sf_df[-1, ]
  for (sf in colnames(data1)) {
    sf_FC <- NULL
    if (sum(data1[[sf]] == 0) == nrow(data1) & sum(data2[[sf]] == 0) == nrow(data2)) {
      sf_df <- rbind(sf_df, c(0, 0, 0, 1))
    } else {
      min_value <- min(c(data1[[sf]][data1[[sf]] != 0], data2[[sf]][data2[[sf]] != 0]))/10
      data1[[sf]][data1[[sf]] == 0] <- min_value
      data2[[sf]][data2[[sf]] == 0] <- min_value
      for (i1 in 1:nrow(data1)) {
        for (i2 in 1:nrow(data2)) {
          sf_FC <- c(sf_FC, data1[[sf]][i1]/data2[[sf]][i2])
        } 
      }
      sf_FC <- log2(sf_FC)
      sf_wil <- wilcox.test(sf_FC, conf.int = TRUE, mu = 0)
      sf_df <- rbind(sf_df, c(as.numeric(sf_wil$estimate), as.numeric(sf_wil$conf.int), as.numeric(sf_wil$p.value)))
    }
  }
  sf_df <- as.data.frame(sf_df); rownames(sf_df) <- colnames(data1); colnames(sf_df) <- c('Median', 'low-CI', 'high-CI', 'p-value')
  return(sf_df)
}
```

### FileCreate

```{r}
FileCreate <- function(DirPath = "./",Prefix = "ExampleTest", Suffix = "pdf", version = "0.0.0"){
  date=as.character(Sys.Date())
  DirPath = gsub("/$","",DirPath)
  Suffix = gsub('^[.]',"",Suffix)
  if(! dir.exists(DirPath)){
    dir.create(DirPath,recursive =TRUE)
  }
  if (version == "0.0.0") {
    version=100
    while(TRUE){
      version_=paste(unlist(strsplit(as.character(version),"")),collapse=".")
      DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
      if(! file.exists(DirPathName)){
        return(DirPathName)
        break
      }
      version = version + 1
    }
  }else{
    DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
    if(! dir.exists(DirPathName)){
      return(DirPathName)
    }
    return(DirPathName)
  }
}
  
```

## Import data {.tabset}

### Taxonomy name

```{r taxonomy name, warning = FALSE}
tax_name <- ImportTable(file = '../00.ProcessData/2021-04-12-allTaxonomySplitLevel-v1.0.1.tsv', sep = '\t')
tax_name$new_name <- gsub(pattern = 's__', replacement = '', x = tax_name$Specie) %>% 
  gsub('_', ' ', .) # due to Feature names of CreateSeuratObject cannot have underscores, replace with blanket space.
show_table(tax_name)
```

### Taxonomy matrix

```{r taxonomy matrix}
raw_df <- ImportTable(file = '../00.ProcessData/2021-04-12-RawData-All-filter1802-v1.0.0.tsv', sep = '\t')
raw_df <- raw_df[grep('^k__Bacteria', rownames(raw_df)),]
tax_df <- apply(raw_df, 2, function(x){x/sum(x)}*100)
tax_df <- as.data.frame(tax_df[rowSums(tax_df!=0) > (0.5* ncol(tax_df)), ])
# tax_df <- ImportTable(file = '../00.ProcessData/2021-04-12-RawData-Fungi-filter1802-v1.0.0.tsv', sep = '\t')
rownames(tax_df) <- tax_name[rownames(tax_df), "new_name"] # use the short species names
show_table(tax_df)
```

### Meta information

```{r meta information}
meta_df <- ImportTable(file = '../00.ProcessData/metaInfo-subgroup-v4.1.csv')
meta_df <- meta_df[colnames(tax_df),]
show_table(meta_df)
```

## Pre-process: selected same trend features

SSTF: selected same trend features, require more than 3/4 cohorts (at less 6 cohorts) perform the same trend <br> MultMedFC: Multiple Median Fold Change, compared each samples between 2 Stages (Control vs CRC) \#\#\# Calculate the trend in different cohorts

```{r, warning = FALSE}
Trend_List <- list()
calculateOrRead <- F
if (calculateOrRead) {
  for (coh in unique(meta_df$Cohort)) {
    message(coh, ' begins at ', date())
    data1 <- tax_df[, rownames(meta_df)[meta_df$Cohort == coh & meta_df$Stage == 'CRC']]
    data2 <- tax_df[, rownames(meta_df)[meta_df$Cohort == coh & meta_df$Stage == 'CTRL']]
    sub_trend <- MultMedFC(data1 = data1, data2 = data2, FeatureAsRow = T)
    Trend_List[[coh]] <- sub_trend
    message(coh, ' has finished at ', date())
    median_file <- FileCreate(DirPath = '../01.SSTF/Cohort', Prefix = paste0('walsh-median-', coh), Suffix = 'csv')
    write.csv(x = Trend_List[[coh]], file = median_file)
  }
  # Combine the cohorts together
  Trend_Comb <- Trend_List[[unique(meta_df$Cohort)[1]]]
  colnames(Trend_Comb) <- paste0(unique(meta_df$Cohort)[1], '-', colnames(Trend_Comb))
  for (coh in unique(meta_df$Cohort)[-1]) {
    sub_trend <- Trend_List[[coh]]
    colnames(sub_trend) <- paste0(coh, '-', colnames(sub_trend))
    Trend_Comb <- cbind(Trend_Comb, sub_trend)
  }

  
  median_file <- FileCreate(DirPath = '../01.SSTF', Prefix = 'walsh-median-ALL', Suffix = 'csv')
  write.csv(x = Trend_Comb, file = median_file)
  Trend_Med_log2 <- Trend_Comb[,grep(pattern = 'Median|p-value', x = colnames(Trend_Comb))]
  # Trend_Med_log2$Average <- apply(Trend_Med_log2[, seq(1,15,2)], 1, mean)
  Trend_Med_log2$N.count <- rowSums(Trend_Med_log2[, seq(1,15,2)] < 0)
  Trend_Med_log2$P.count <- rowSums(Trend_Med_log2[, seq(1,15,2)] > 0)
  Trend_Med_log2$Missing <- rowSums(Trend_Med_log2[, seq(1,15,2)] == 0)
  Trend_Med_log2$pvalue <- rowSums(Trend_Med_log2[, seq(2,16,2)] < 0.05)
  Trend_Med_log2 <- Trend_Med_log2[, c(17:20,1:16)]
  median_file <- FileCreate(DirPath = '../01.SSTF', Prefix = 'Summary-walsh-median', Suffix = 'csv')
  write.csv(x = Trend_Med_log2, file = median_file)
  
}else{
  for (coh in unique(meta_df$Cohort)) {
    Trend_List[[coh]] <- read.csv(paste0('../01.SSTF/Cohort/2021-07-07-walsh-median-', coh, '-v1.0.0.csv'), header = T, row.names = 1, check.names = F)
  }
  Trend_Comb <- read.csv('../01.SSTF/2021-07-07-walsh-median-ALL-v1.0.0.csv', header = T, row.names = 1, check.names = F)
  Trend_Med_log2 <- read.csv('../01.SSTF/2021-07-07-Summary-walsh-median-v1.0.0.csv', header = T, row.names = 1, check.names = F)
}

show_table(Trend_Med_log2) %>%
  formatSignif(columns = colnames(Trend_Med_log2[, c(5:20)]),
               digits = 3, interval = 1) %>%
  formatRound(columns = colnames(Trend_Med_log2[, c(1:4)]),
              digits = 0, interval = 1) %>%
  formatStyle(columns = colnames(Trend_Med_log2[, c(1:4)]),
              fontWeight = 'bold')
```
