---
title: "Comparison Features after Batch Effect"
author: "ifanlyn@outlook.com"
date: "7/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 20, fig.height=7)
```

## Loading packages and subroutines {.tabset}

### Packages

```{r packages, message=FALSE, warning=FALSE}
require(ggplot2)
require(DT) # for datatable
require(dplyr) # for mutate.
require(ggpubr) # for stat_compare_means
# require(ComplexHeatmap) # for Heatmap
require(circlize)  # for colorRamp2
require(ochRe)  # for ochre_palettes
```

### addZeroMatrix

```{r}
addZeroMatrix <-function(data, min.value = 0, shrink = 0.1, seedNum =123){
  set.seed(seedNum)
  if (min(data) <= min.value) {
    new.min <- min(data[data > min.value])*shrink
    min.len <- sum(data <= min.value)
    data[data <= min.value]  <- rnorm(n = min.len, mean = new.min, sd = new.min*0.1)
  }
  return(data)
}
```

### wilcoxFC

```{r}
wilcoxFC <- function(data, group, FCrank, seednum = 123){
  set.seed(seednum)
  data <- t(addZeroMatrix(data, min.value = 0))
  df1 <- as.matrix(data[names(group)[group == FCrank[1]], ])
  df2 <- as.matrix(data[names(group)[group == FCrank[2]], ])
  logCI_df <- matrix(NA, nrow = ncol(data), ncol = 3)
  
  for (i in 1:ncol(data)) {
    df1_ <- df1[, i, drop=FALSE]
    df2_ <- as.matrix(t(1/df2[, i, drop=FALSE]))
    fc_df <- df1_ %*% df2_
    fc_num <- log2(as.numeric(fc_df))
    wil_res <- wilcox.test(fc_num, conf.int = T)
    logCI_df[i, ] <- as.numeric(c(wil_res$estimate, wil_res$conf.int))
  }
  logCI_df <- as.data.frame(logCI_df); dimnames(logCI_df) <- list(colnames(data), c('observed value', 'low CI', 'high CI'))
  logCI_df <- logCI_df[order(sign(logCI_df[,1]), abs(logCI_df[,1]), decreasing = T),]
  # logCI_df <- logCI_df[rev(rownames(logCI_df)), ]
  return(logCI_df)
}
```

### wilcoxBaplot

```{r}
wilcoxBarplot <- function(data_list, group_list, FCrank, seednum = 123){
  ci_all <- matrix(NA, nrow = 1, ncol = 4); ci_all <- ci_all[-1, ]
  for (i in names(data_list)) {
    set.seed(seednum)
    ci_df <- wilcoxFC(data =  data_list[[i]], group = group_list[[i]], FCrank = FCrank, seednum = seednum)
    ci_df <- cbind(Cohort = rep(i, nrow(ci_df)), cbind(Fungi = rownames(ci_df), ci_df))
    ci_all <- rbind(ci_all, ci_df)
  }
  return(ci_all)
}
```

### show_table

```{r show_table}
show_table <- function(df, rownames = T, filter="top", options = list(pageLength = 10, scrollX=T), ...){
  if (ncol(df) > 50) {
    df <- df[, 1:50]
    message('Due to the column dim is > 50, it only shows the ')
  }
  datatable(df, rownames = rownames, 
            filter=filter, options = options, ... )

}
```

### FileCreate

```{r}
FileCreate <- function(DirPath = "./",Prefix = "ExampleTest", Suffix = "pdf", version = "0.0.0"){
  date=as.character(Sys.Date())
  DirPath = gsub("/$","",DirPath)
  Suffix = gsub('^[.]',"",Suffix)
  if(! dir.exists(DirPath)){
    dir.create(DirPath,recursive =TRUE)
  }
  if (version == "0.0.0") {
    version=100
    while(TRUE){
      version_=paste(unlist(strsplit(as.character(version),"")),collapse=".")
      DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
      if(! file.exists(DirPathName)){
        return(DirPathName)
        break
      }
      version = version + 1
    }
  }else{
    DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
    if(! dir.exists(DirPathName)){
      return(DirPathName)
    }
    return(DirPathName)
  }
}
  
```

### Import table

```{r ImportTable}
ImportTable <- function(file, header = T, row.names = 1, sep = ',', check.names = FALSE, ...){
  data <- read.csv(file = file, header = header, row.names = row.names, sep = sep, check.names = check.names, ...)
  return(data)
}
```

### matrix_wilcox

matrix_wilcox: subroutine to comparison each row in matrix

```{r subroutine: wilcoxon test from matrix}
matrix_wilcox <- function(data, group, ad_pvalue = 'BH'){
  data <- as.matrix(data)
  group1 <- names(group)[group == unique(group)[1]]
  group2 <- names(group)[group == unique(group)[2]]
  p_values <- list()
  for (i in 1:nrow(data)) {
    if (sum(data[i, ]) == 0) {
      p_values[[rownames(data)[i]]] = 1
    }else{
      p_values[[rownames(data)[i]]] = wilcox.test(data[i, group1], data[i, group2], 
                                                  alternative = "two.sided", exact = FALSE)$p.value
    }
  }
  p_values = data.frame(p_values = sapply(p_values, c))
  p_values$adj_pvalue = p.adjust(p_values$p_values, method = ad_pvalue, n = nrow(p_values))
  return(p_values)
}
```

## Import data {.tabset}

### Profile of the trend

In previously analysis, we compared the trend of fungi in various cohorts.

```{r trend}

trend_df <- ImportTable(file = '../01.SSTF/2021-07-07-Summary-walsh-median-v1.0.0.csv')

trend_df <- trend_df[, -grep(pattern = "pvalue|p-value", x = colnames(trend_df))]# ignore the pvalue columns

trend_df <- trend_df[order(trend_df$P.count, decreasing = T), ]

show_table(trend_df) %>% 
  formatSignif(columns = colnames(trend_df)[3:11],
               digits = 3, interval = 1)


```

### Taxonomy name

```{r taxonomy name, warning = FALSE}
tax_name <- ImportTable(file = '../00.ProcessData/2021-04-12-allTaxonomySplitLevel-v1.0.1.tsv', sep = '\t')
tax_name$new_name <- gsub(pattern = 's__', replacement = '', x = tax_name$Specie) %>% 
  gsub('_', ' ', .) # due to Feature names of CreateSeuratObject cannot have underscores, replace with blanket space.
show_table(tax_name)
```

### Taxonomy data

Import the raw taxonomy data

```{r taxonomy data }
tax_df <- ImportTable(file = '../00.ProcessData/2021-04-12-ReAbun-Fungi-filter1802-v1.0.0.tsv', sep = '\t')
rownames(tax_df) <- tax_name[rownames(tax_df), "new_name"] # use the short species names

show_table(tax_df) %>%
  formatSignif(columns = colnames(tax_df)[1:50],
               digits = 3, interval = 1)
```

### Meta information

```{r meta information}
meta_df <- ImportTable(file = '../00.ProcessData/metaInfo-subgroup-v4.1.csv')
meta_df <- meta_df[colnames(tax_df), ]
show_table(meta_df)
```

## Data filtering {.tabset}

filter the features step by step:

### Filtered the excursive trend features

**Five features** plays enriched trend in the CRC group, and **80 characters** are depleted in CRC. (at less 7 cohort play the same trend would be included)

```{r selected same trend featre}
sel_tre <- trend_df[trend_df$N.count >= 7 | trend_df$P.count >= 7,]

show_table(sel_tre) %>%
  formatSignif(columns = colnames(sel_tre)[3:11],
               digits = 3, interval = 1)
```

### Wilcoxon test (p value)

Comparison the features in different stages from each cohorts. **Eighty-one** fungi play a significant difference ( p-value \< 0.05 ) in at least half cohorts ( above or equal to 4 ).

-   Only one fungi ( ***Aspergillus rambellii*** ) plays a significant difference in 7 cohorts.

-   Five fungi ( *Rhizophagus irregularis, Leucoagaricus sp. SymC.cos, Cyberlindnera fabianii, Paracoccidioides lutzii, Phialocephala subalpina* ) play a significant difference in 6 cohorts.

-   Twenty-two and fifty-three fungi play a difference in 5 and 4 cohorts.

```{r wilcoxon}
calculate_or_not <- F
wil_res <- NULL
p_value_cutoff <- 0.05
adj_p_cutoff <- 0.1
counts_cutoff <- 4
if(calculate_or_not){
	for (i in unique(meta_df$Cohort)) {
	  sub_meta <- meta_df[meta_df$Cohort == i & meta_df$Stage != 'adenoma', ]
	  sub_data <- tax_df[, rownames(sub_meta)]
	  sub_group <- sub_meta$Stage
	  names(sub_group) <- rownames(sub_meta)
	  sub_wil <- matrix_wilcox(data = sub_data, group = sub_group)
	  colnames(sub_wil) <- paste0(i, '-', colnames(sub_wil))
	  # message(i, ' has finished at ', date())
	  if (is.null(wil_res)) {
		  wil_res <- sub_wil
	  }else{
		  wil_res <- cbind(wil_res, sub_wil)
	  }
	}
	summarized_wilcox_df <- data.frame(
	  pvalue_counts = rowSums(wil_res[,grep('p_values', colnames(wil_res))] < p_value_cutoff),
	  adj_p_counts = rowSums(wil_res[,grep('adj_pvalue', colnames(wil_res))] < adj_p_cutoff)
	  )
	summarized_wilcox_df <- as.data.frame(cbind(summarized_wilcox_df, wil_res))
	summarized_wilcox_df <- summarized_wilcox_df[order(summarized_wilcox_df$pvalue_counts, decreasing = T), ]

	summ_wilcox_csv <- FileCreate(DirPath = '../01.wilcox',
								  Prefix = 'summarized_wilcoxon-test',Suffix = 'csv')
	write.csv(x = summarized_wilcox_df, file = summ_wilcox_csv)

	sel_wilcox_df <- summarized_wilcox_df[summarized_wilcox_df$pvalue_counts >= counts_cutoff, ]

	sel_wilcox_csv <- FileCreate(DirPath = '../01.wilcox',
								 Prefix = 'selected-summarized_wilcoxon-test',Suffix = 'csv')
	write.csv(x = sel_wilcox_df, file = sel_wilcox_csv)
}else{
  summarized_wilcox_df <- ImportTable(file = '../01.wilcox/2021-07-13-summarized_wilcoxon-test-v1.0.0.csv')
  sel_wilcox_df <- ImportTable(file = '../01.wilcox/2021-07-13-selected-summarized_wilcoxon-test-v1.0.0.csv')
}
	

show_table(sel_wilcox_df) %>%
  formatSignif(columns = colnames(summarized_wilcox_df)[3:18],
               digits = 3, interval = 1)


```

## Overlap between SSTF & Wilcox filtering {.tabset .tabset-fade}

The relative abundance has been transformed by [arcsin-3thSquareRoot](https://rpubs.com/ozjimbob/asqrt) in the beginning.

### Code

```{r}
overlap_candidates <- intersect(rownames(sel_wilcox_df), rownames(sel_tre))
plot_or_not <- F

ol_df <- asin((t(tax_df[overlap_candidates,    
                        rownames(meta_df)[meta_df$Stage != 'adenoma']])*.01)^(1/3))   # ol_df means: overlap between 2 filters

ol_df <- cbind(meta_df[meta_df$Stage != 'adenoma',c("Cohort", "Stage")], ol_df)
ol_df$Stage <- factor(x = ol_df$Stage, levels = c('CTRL', 'CRC'))

if (plot_or_not) {
  g_list <- list()
  wil_box_pdf <- FileCreate(DirPath = '../02.DataFilter', Prefix = 'Boxplot-SSTF-Wilcox',
                          Suffix = 'pdf')
  pdf(file = wil_box_pdf, width = 20, height = 7)
  for (i in overlap_candidates) {
    agd1 = group_by(ol_df, Cohort, Stage) %>% 
      summarise(mn = mean(.data[[i]]),sd = sd(.data[[i]]), 
                 .groups = 'keep') %>%
      mutate(up = mn+sd, lw = mn-sd)
    agd1$Stage <- factor(x = agd1$Stage, levels = c('CTRL', 'CRC'))
     
    g <- ggplot() +
      geom_boxplot(data = ol_df, mapping = aes(x = Cohort, y = .data[[i]], color = Stage),
                   position = position_dodge(0.8), outlier.shape = NA, width = 0.5)+
      geom_jitter(data = ol_df, mapping = aes(x = Cohort, y = .data[[i]], color = Stage),
                   position = position_jitterdodge(0.2), alpha = 0.1) +
      # geom_errorbar(aes(ymin = lw, ymax = up), width = 0.2, position = position_dodge(0.8))+
      geom_point(data = agd1, mapping = aes(x = Cohort, y = mn, group = Stage),
                  position = position_dodge(0.8), color = 'black', 
                  shape = 10, size = 4) +
      scale_color_manual(values = c("#00AFBB", "red"))  + theme_bw() +
      theme(legend.position = "top",
            text = element_text(size = 15),
            plot.title = element_text(color=ifelse(sel_tre[i, ]$P.count > 4, "#993333", "#1D2D5F"),
                                      size = 20, hjust = 0.5),
            plot.subtitle = element_text(color=ifelse(sel_tre[i, ]$P.count > 4, "#993333", "#1D2D5F"),
                                         size = 12, hjust = 1)) +
      stat_compare_means(data = ol_df, 
                         mapping = aes(x = Cohort, y = .data[[i]], group = Stage),
                         method = 'wilcox.test')+
      ylab(paste0('acsin 3th square root of ', i)) + 
      ggtitle(label = i, subtitle = ifelse(sel_tre[i, ]$P.count > 4, 
                                           paste0(sel_tre[i, ]$P.count, " cohorts has same trend in CRC entiched.\n", 
                                                  sel_wilcox_df[i, ]$pvalue_counts, " cohorts play significant different."),
                                           paste0(sel_tre[i, ]$N.count, " cohorts has same trend in CRC depleted.\n", 
                                                  sel_wilcox_df[i, ]$pvalue_counts, " cohorts play significant different.")))
    plot(g)
    g_list[[i]] <- g
  }
dev.off()
plot_rds <- FileCreate(DirPath = '../02.DataFilter/', Prefix = 'Boxplot-SSTF-Wilcox',
                          Suffix = 'rds')
saveRDS(g_list, plot_rds)
}else{
  g_list <- readRDS(file = '../02.DataFilter/2021-07-13-Boxplot-SSTF-Wilcox-v1.0.0.rds')
}
```

### Aspergillus rambellii

```{r}
plot(g_list$`Aspergillus rambellii`)
```

### Leucoagaricus sp. SymC.cos

```{r}
plot(g_list$`Leucoagaricus sp. SymC.cos`)
```

### Rhizophagus irregularis

```{r}
plot(g_list$`Rhizophagus irregularis`)
```

### Piromyces sp. E2

```{r}
plot(g_list$`Piromyces sp. E2`)
```

### Hanseniaspora opuntiae

```{r}
plot(g_list$`Hanseniaspora opuntiae`)
```

### Torulaspora delbrueckii

```{r}
plot(g_list$`Torulaspora delbrueckii`)
```

### Coccidioides immitis

```{r}
plot(g_list$`Coccidioides immitis`)
```

### Aspergillus kawachii

```{r}
plot(g_list$`Aspergillus kawachii`)
```

### Trichoderma atroviride

```{r}
plot(g_list$`Trichoderma atroviride`)
```

### Batrachochytrium salamandrivorans

```{r}
plot(g_list$`Batrachochytrium salamandrivorans`)
```

### Mucor circinelloides

```{r}
plot(g_list$`Mucor circinelloides`)
```

### Edhazardia aedis

```{r}
plot(g_list$`Edhazardia aedis`)
```

### Encephalitozoon hellem

```{r}
plot(g_list$`Encephalitozoon hellem`)
```

### Tilletia controversa

```{r}
plot(g_list$`Tilletia controversa`)
```

### Kwoniella pini

```{r}
plot(g_list$`Kwoniella pini`)
```

### Komagataella pastoris

```{r}
plot(g_list$`Komagataella pastoris`)
```

### Brettanomyces bruxellensis

```{r}
plot(g_list$`Brettanomyces bruxellensis`)
```

### Debaryomyces hansenii

```{r}
plot(g_list$`Debaryomyces hansenii`)
```

### Hyphopichia burtonii

```{r}
plot(g_list$`Hyphopichia burtonii`)
```

### Eremothecium sinecaudum

```{r}
plot(g_list$`Eremothecium sinecaudum`)
```

### Torulaspora globosa

```{r}
plot(g_list$`Torulaspora globosa`)
```

### Tuber magnatum

```{r}
plot(g_list$`Tuber magnatum`)
```

### Pseudocercospora fijiensis

```{r}
plot(g_list$`Pseudocercospora fijiensis`)
```

### Emmonsia sp. CAC-2015a

```{r}
plot(g_list$`Emmonsia sp. CAC-2015a`)
```

### Rutstroemia sp. NJR-2017a BVV2

```{r}
plot(g_list$`Rutstroemia sp. NJR-2017a BVV2`)
```

### Rutstroemia sp. NJR-2017a WRK4

```{r}
plot(g_list$`Rutstroemia sp. NJR-2017a WRK4`)
```

### Thielaviopsis punctulata

```{r}
plot(g_list$`Thielaviopsis punctulata`)
```

### Colletotrichum fioriniae

```{r}
plot(g_list$`Colletotrichum fioriniae`)
```

##  {.unnumbered}

## Error Bar plot of Fold Change {.tabset}

-   Due to 0 couldn't be the denominator, the all the zero would instead by the value: `rnorm(n, min_value, min_sd)`,

    -   `n` means the counts of zero in the OTU matrix.

    -   `min_value` means the min value excluded the 0.

    -   `min_sd` equal `min_value /10`

-   In order to reduce the bias of Fold change, the Fold Change was calculate by:

    $$
    Fold\ Change_i = median(\displaystyle \sum^{n_{crc}}_{j\ =\ 1}\displaystyle \sum^{n_{ctrl}}_{k\ =\ 1}{\frac{ReAbund_{j\ i}}{ReAbund_{k\ i}}})
    $$

    -   `i` means the fungi names

    -   `n_CRC / n_CTRL` means the counts of CRC/ control samples.

-   Totally **28** fungi leave after twice filters, and **6** of them are also general Fold change difference in all samples (`|log2(FC)| > 0.5`).

### code

```{r}
calculate_or_not <- F

if (calculate_or_not) {
  wilc_err_res <- wilcoxBarplot(data_list = hm_mt_list[['RelAbundance']], group_list = hm_mt_list[['group']], FCrank = c('CRC', 'CTRL'))
  all_group <- meta_df$Stage[meta_df$Stage != 'adenoma']; names(all_group) <- rownames(meta_df)[meta_df$Stage != 'adenoma']
  all_wilc_res <- wilcoxFC(data = tax_df[overlap_candidates, names(all_group)], group = all_group, FCrank = c('CRC', 'CTRL'))
  wilc_err_res_csv <- FileCreate(DirPath = '../02.DataFilter/', Prefix = '28_sel_cand-wilcox-CI-cohort', Suffix = 'csv')
  write.csv(x = wilc_err_res, file = wilc_err_res_csv)
  
  all_err_res_csv <- FileCreate(DirPath = '../02.DataFilter/', Prefix = '28_sel_cand-wilcox-CI-all', Suffix = 'csv')
  write.csv(x = all_wilc_res, file = all_err_res_csv)
}else{
  wilc_err_res <- ImportTable(file = '../02.DataFilter/2021-07-14-28_sel_cand-wilcox-CI-cohort-v1.0.0.csv')
  all_wilc_res <- ImportTable(file = '../02.DataFilter/2021-07-14-28_sel_cand-wilcox-CI-all-v1.0.0.csv')
}


errbar_mt <- all_wilc_res
errbar_mt$Name <- factor(x = rownames(errbar_mt), levels = rev(rownames(errbar_mt)))
errbar_mt$col <- ifelse(errbar_mt$`observed value` > 0, 'red', 'blue')


cohort_col <-ochre_palettes$parliament; names(cohort_col) <- unique(meta_df$Cohort)

all_plot <- ggplot(data = errbar_mt, mapping = aes(x = Name, y = `observed value`, color = col))+
  geom_bar(stat="identity", fill = 'white', width = 0.7) +
  scale_color_manual(values = c(red = 'red', blue = 'blue', cohort_col)) + theme_bw()+
  scale_shape_manual(values = 0:7)+
  geom_errorbar(aes(ymin=`low CI`, ymax=`high CI`), width=.2) +
  geom_point(data = wilc_err_res, mapping = aes(x = Fungi, y = `observed value`, color = Cohort, shape = Cohort), alpha = 0.5) + ylab('Log2(FC)') +
  coord_flip()

sel_errbar_mt <- errbar_mt[abs(errbar_mt$`observed value`) > 0.5, ]
sel_errbar_mt$Name <- factor(x = rownames(sel_errbar_mt), levels = rev(rownames(sel_errbar_mt)))

sel_wilc_err_res <- wilc_err_res[wilc_err_res$Fungi %in% rownames(sel_errbar_mt), ]

sel_plot <- ggplot(data = sel_errbar_mt, mapping = aes(x = Name, y = `observed value`, color = col))+
  geom_bar(stat="identity", fill = 'white', width = 0.7) +
  scale_color_manual(values = c(red = 'red', blue = 'blue', cohort_col)) + theme_bw()+
  scale_shape_manual(values = 0:7)+
  geom_errorbar(aes(ymin=`low CI`, ymax=`high CI`), width=.2) +
  # geom_errorbar(data = sel_wilc_err_res, mapping = aes(x = Fungi, ymin=`low CI`, ymax=`high CI`, color = Cohort), width=.2) +
  geom_point(data = sel_wilc_err_res, mapping = aes(x = Fungi, y = `observed value`, color = Cohort, shape = Cohort), alpha = 0.5)+ ylab('Log2(FC)') +
  coord_flip()

if (calculate_or_not) {
  all_plot_pdf <- FileCreate(DirPath = '../02.DataFilter/', Prefix = 'ErrBar-FC-all_28-wilcox', Suffix = 'pdf')
  pdf(file = all_plot_pdf, width = 8, height = 10)
  plot(all_plot)
  dev.off()
  
  sel_plot_pdf <- FileCreate(DirPath = '../02.DataFilter/', Prefix = 'ErrBar-FC-sel_6_FC_0.5-wilcox', Suffix = 'pdf')
  pdf(file = sel_plot_pdf, width = 8, height = 2.5)
  plot(sel_plot)
  dev.off()
}
```

### all 28 candidates

```{r, fig.height= 10, fig.width=8}
plot(all_plot)
```

### selected 6 candidates

`|log2(FC)| > 0.5`

```{r, fig.height= 2.5, fig.width=7}
plot(sel_plot+ theme(legend.position = "none"))
```
