---
title: "BatchEffect-Seurat"
author: "ifanlyn@outlook.com"
date: "7/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages and subroutines {.tabset}

### Packages

```{r packages, message=FALSE, warning=FALSE}
require(Seurat)
require(DT) # for datatable
require(SeuratWrappers) # for fastMNN
require(harmony) # for RunHarmony
require(reshape2) # for dcast
```

### Import table
```{r ImportTable }
ImportTable <- function(file, header = T, row.names = 1, sep = ',', check.names = FALSE, ...){
  data <- read.csv(file = file, header = header, row.names = row.names, sep = sep, check.names = check.names, ...)
  return(data)
}
```

### show_table
```{r show_table}
show_table <- function(df, rownames = T, filter="top", options = list(pageLength = 10, scrollX=T), ...){
  if (ncol(df) > 50) {
    df <- df[, 1:50]
    message('Due to the column dim is > 50, it only shows the ')
  }
  datatable(df, rownames = rownames, 
            filter=filter, options = options, ... )

}
```

### MultMedFC 
MultMedFC: Multiple Median Fold Change, compared each samples between 2 Stages (Control vs CRC)
```{r}
MultMedFC <- function(data1, data2, FeatureAsRow = Tsf){
  if (FeatureAsRow) {
    data1 <- t(data1)
    data2 <- t(data2)
  }
  data1 <- as.data.frame(data1)
  data2 <- as.data.frame(data2)
  if (! all(colnames(data1) == colnames(data2))) {
    message("rowname of data1 and data2 is not fullly same!! plz check it !!")
    return()
  }
  
  sf_df <- matrix(NA, ncol = 3); sf_df <- sf_df[-1, ]
  for (sf in colnames(data1)) {
    sf_FC <- NULL
    if (sum(data1[[sf]] == 0) == nrow(data1) & sum(data2[[sf]] == 0) == nrow(data2)) {
      sf_df <- rbind(sf_df, c(1,1,1))
    }else{
      min_value <- min(c(data1[[sf]][data1[[sf]] != 0], data2[[sf]][data2[[sf]] != 0]))/10
      data1[[sf]][data1[[sf]] == 0] <- min_value
      data2[[sf]][data2[[sf]] == 0] <- min_value
      for (i1 in 1:nrow(data1)) {
        for (i2 in 1:nrow(data2)) {
          sf_FC <- c(sf_FC, data1[[sf]][i1]/data2[[sf]][i2])
        } 
      }
      sf_wil <- wilcox.test(sf_FC, conf.int = TRUE)
      sf_df <- rbind(sf_df, c(as.numeric(sf_wil$estimate), as.numeric(sf_wil$conf.int)))
    }
  }
  sf_df <- as.data.frame(sf_df); rownames(sf_df) <- colnames(data1); colnames(sf_df) <- c("Median", 'low-CI', 'high-CI')
  return(sf_df)
}
```

### FileCreate

```{r}
FileCreate <- function(DirPath = "./",Prefix = "ExampleTest", Suffix = "pdf", version = "0.0.0"){
  date=as.character(Sys.Date())
  DirPath = gsub("/$","",DirPath)
  Suffix = gsub('^[.]',"",Suffix)
  if(! dir.exists(DirPath)){
    dir.create(DirPath,recursive =TRUE)
  }
  if (version == "0.0.0") {
    version=100
    while(TRUE){
      version_=paste(unlist(strsplit(as.character(version),"")),collapse=".")
      DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
      if(! file.exists(DirPathName)){
        return(DirPathName)
        break
      }
      version = version + 1
    }
  }else{
    DirPathName = paste0(DirPath,"/",date,'-',Prefix,'-v',version_,".",Suffix)
    if(! dir.exists(DirPathName)){
      return(DirPathName)
    }
    return(DirPathName)
  }
}
  
```

## Import data {.tabset}


### Selected matrix
+ leave the features with p-value less than 0.05 in all cohorts,
+ select the feature with same trend in all cohorts

```{r selected matrix}
summarized_df <- ImportTable(file = '../01.SSTF/2021-07-06-Summary-walsh-median-v1.0.0.csv')
show_table(summarized_df)
sel_df <- summarized_df[summarized_df$pvalue == 8 & (summarized_df$P.count == 8 | summarized_df$N.count == 8), ]
sel2_df <- summarized_df[summarized_df$pvalue >= 7 & (summarized_df$P.count >= 7 | summarized_df$N.count >= 7), ]
sel3_df <- summarized_df[summarized_df$pvalue >= 6 & (summarized_df$P.count >= 6 | summarized_df$N.count >= 6), ]

```
### Taxonomy name 
```{r taxonomy name, warning = FALSE}
tax_name <- ImportTable(file = '../00.ProcessData/2021-04-12-allTaxonomySplitLevel-v1.0.1.tsv', sep = '\t')
tax_name$new_name <- gsub(pattern = 's__', replacement = '', x = tax_name$Specie) %>% 
  gsub('_', ' ', .) # due to Feature names of CreateSeuratObject cannot have underscores, replace with blanket space.
show_table(tax_name)
```

### Taxonomy matrix

```{r taxonomy matrix}
raw_tax_df <- ImportTable(file = '../00.ProcessData/2021-04-12-ReAbun-Fungi-filter1802-v1.0.0.tsv', sep = '\t')
# tax_df <- ImportTable(file = '../00.ProcessData/2021-04-12-RawData-Fungi-filter1802-v1.0.0.tsv', sep = '\t')
rownames(raw_tax_df) <- tax_name[rownames(raw_tax_df), "new_name"] # use the short species names
show_table(raw_tax_df)
tax_df <- raw_tax_df[rownames(sel2_df),]
show_table(tax_df)

```



### Meta information 
```{r meta information}
meta_df <- ImportTable(file = '../00.ProcessData/metaInfo-subgroup-v4.1.csv')
meta_df <- meta_df[colnames(tax_df),]
show_table(meta_df)
```


## Remove the batch effect by Seurat 

### Create seurat object
CreateSeuratObject: Create a Seurat object [link](https://www.rdocumentation.org/packages/Seurat/versions/3.0.1/topics/CreateSeuratObject)
```{r create seurat object}
s_obj <- CreateSeuratObject(counts = tax_df, meta.data = meta_df)

# split the object by cohorts
s_list <- SplitObject(s_obj, split.by = 'Cohort')
```

### Remove batch effect {.tabset}

#### RPCA {.tabset}
Fast integration using reciprocal PCA (RPCA) [link](https://satijalab.org/seurat/articles/integration_rpca.html)

##### Pre-process
```{r RPCA, message=FALSE, warning=FALSE, fig.width = 20, fig.height = 7}

for (i in 1:length(s_list)) {
    s_list[[i]] <- NormalizeData(s_list[[i]], verbose = FALSE)
    s_list[[i]] <- FindVariableFeatures(s_list[[i]], selection.method = "vst", nfeatures = 2000, verbose = FALSE)
}

s_anchors <- FindIntegrationAnchors(object.list = s_list, k.anchor = 20)

s_integrated <- IntegrateData(anchorset = s_anchors, k.weight = 10) %>%
  ScaleData(. , verbose = FALSE) %>% 
  RunPCA(. , npcs = 30, verbose = FALSE) %>%
  RunUMAP(. , reduction = "pca", dims = 1:30, verbose = FALSE)

prca_obj <- FindNeighbors(s_integrated, dims = 1:30) %>%
  FindClusters()

prca_table <- dcast(as.data.frame(table(prca_obj@active.ident,prca_obj@meta.data[["Stage"]])), Var1 ~ Var2)
rownames(prca_table) <- prca_table$Var1; prca_table <- prca_table[,-1]
```
##### Table 
```{r, message=FALSE, warning=FALSE, fig.width = 20, fig.height = 7}
show_table(prca_table)
```

##### Figure
```{r, message=FALSE, warning=FALSE, fig.width = 20, fig.height = 7}
DimPlot(prca_obj, group.by = c("Cohort", "Stage", 'ident'), ncol = 3)
```

#### fastMNN {.tabset}
fastMNN: Batch effects in single-cell RNA-sequencing data are corrected by matching mutual nearest neighbors [link](https://htmlpreview.github.io/?https://github.com/satijalab/seurat.wrappers/blob/master/docs/fast_mnn.html) <br>
Nature Biotechnology, 2018, DOI: [10.1038/nbt.4091](https://www.nature.com/articles/nbt.4091)

##### Pre-process

```{r fastMNN, message=FALSE, warning=FALSE, fig.width = 20, fig.height = 7}
fMNN_obj <- NormalizeData(s_obj) %>% 
  FindVariableFeatures() %>%
  SplitObject(. ,  split.by = 'Cohort') %>%
  RunFastMNN()

fMNN_obj2 <- RunUMAP(fMNN_obj, reduction = 'mnn', dims = 1:30) %>%
  FindNeighbors(. , reduction = 'mnn', dims = 1:30) %>%
  FindClusters()

fMNN_table <- dcast(as.data.frame(table(fMNN_obj2@active.ident,fMNN_obj2@meta.data[["Stage"]])), Var1 ~ Var2)
rownames(fMNN_table) <- fMNN_table$Var1; fMNN_table <- fMNN_table[,-1]

```

##### Table
```{r, message=FALSE, warning=FALSE, fig.width = 20, fig.height = 7}
show_table(fMNN_table)
```

##### Figure
```{r, message=FALSE, warning=FALSE, fig.width = 20, fig.height = 7}
DimPlot(fMNN_obj2, group.by = c("Cohort", "Stage", 'ident'), ncol = 3)
```

#### Harmony {.tabset}
Fast, sensitive, and flexible integration of single cell data with Harmony. [link](https://htmlpreview.github.io/?https://github.com/satijalab/seurat.wrappers/blob/master/docs/harmony.html) <br>

Nature Methods, 2019, DOI: [10.1038/s41592-019-0619-0](https://www.nature.com/articles/s41592-019-0619-0)

##### Pre-process
```{r Harmony, message=FALSE, warning=FALSE, fig.width = 20, fig.height = 7}
hmy_obj <- NormalizeData(s_obj) %>%
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA(. , verbose = FALSE)

hmy_obj2 <- RunHarmony(object = hmy_obj, group.by.vars = 'Cohort') %>%
  RunUMAP(., reduction = "harmony", dims = 1:30) %>%
  FindNeighbors(., reduction = "harmony", dims = 1:30) %>% 
  FindClusters()

hmy_table <- dcast(as.data.frame(table(hmy_obj2@active.ident,hmy_obj2@meta.data[["Stage"]])), Var1 ~ Var2)
rownames(hmy_table) <- hmy_table$Var1; hmy_table <- hmy_table[,-1]


```

##### Table
```{r, message=FALSE, warning=FALSE, fig.width = 20, fig.height = 7}
show_table(hmy_table)
```

##### Figure
```{r, message=FALSE, warning=FALSE, fig.width = 20, fig.height = 7}
DimPlot(hmy_obj2, group.by = c("Cohort", "Stage", 'ident'), ncol = 3)
```

<!--#### test 
```{r}
library(Seurat)
library(SeuratData)
library(SeuratWrappers)
data("panc8")
panc8 <- NormalizeData(panc8)
panc8 <- FindVariableFeatures(panc8)
panc8 <- RunFastMNN(object.list = SplitObject(panc8, split.by = "replicate")[c("celseq", "celseq2", "fluidigmc1", "smartseq2")])
panc8 <- RunUMAP(panc8, reduction = "mnn", dims = 1:30)
panc8 <- FindNeighbors(panc8, reduction = "mnn", dims = 1:30)
panc8 <- FindClusters(panc8)
DimPlot(panc8, group.by = c("replicate", "ident", "celltype"), ncol = 3)

```
-->




